### 2023-09-06 Continuity coding final.Rmd
initial commit of paper coding files

---
title: "SEER Medicaid"
author: "Kim Johnson"
date: 
output: html_document
---

# Open libraries
```{r}
# install.packages("pacman") 
pacman::p_load(readr, dplyr, tidyverse, openxlsx) # installs if not installed and opens library
```
# A Import and create dataframes and variables needed for continuity variable classifications

# Steps A1 and A2
- Import mcaid file made by co-author DB that has enrollment data coded for each month 2006 through 2013
- Import SEER0to39 file (created from medicaid.cancer.text file and limited to those diagnosed between ages 0 to 39 years)
```{r}
# A1
mcaid <- read_csv("/Volumes/kijohnson/Active/mcseer_dsb/archive/mcseer_dsb.csv") # created by coauthor DB

mcaid <- mcaid %>%
  select(1:97) # keep just mcaid variables that give monthly enrollment classification (0 = not enrolled, 1 = enrolled, NA) and patient id

# A2
seer0to39<-read_csv("/Volumes/kijohnson/Active/seer0to39.csv") # this is 0 to 39 cancer patient data, created by KJ in paper analysis code
```

# Steps A3 
- Limit SEER0to39 file with exclusions that are the same as the analytic dataset used for the paper analysis
- Merge variables from SEER and mcaid data 
- Create variables  needed for continuity variable construction (dxmo, dthmo, earlylate)
```{r}
# need this dataset for merge for exclusions becasue the mcaid file was not all unique ids
seer0to39_formerge <- seer0to39 %>% # created by KJ step A2 of protocol
select(patient_id, Year_of_diagnosis, Month_of_diagnosis_recode, Month_of_diagnosis, SEER_DateofDeath_Year, SEER_DateofDeath_Month, Firstmalignantprimary_indicator, Sequence_number, In_Medicaid_Flag) %>% 
  filter(Firstmalignantprimary_indicator == 1, In_Medicaid_Flag == 1) %>%
  filter(Sequence_number %in% c("00","01")) # to get unique ids
```

# Step A4
```{r}
mcaid0to39<- left_join(seer0to39_formerge, mcaid,  by = "patient_id")  # left with unique SEER ids 0 to 39
```

# Step A5a
- scale diagnosis and death month variables
```{r}
# Add 0, 12, 36, 48, 60, 72, and 84 to dx mo and death mo for diagnosis years and death years 2006 to 2013, respectively to get dxmo and deathmo on 1 to 96 (or higher for dthmo) 
mcaid0to39 <- mcaid0to39 %>%
  mutate(dxmo = case_when(Year_of_diagnosis == 2006 ~ Month_of_diagnosis_recode + 0,
                          Year_of_diagnosis == 2007 ~ Month_of_diagnosis_recode + 12,
                          Year_of_diagnosis == 2008 ~ Month_of_diagnosis_recode + 24,
                          Year_of_diagnosis == 2009 ~ Month_of_diagnosis_recode + 36,
                          Year_of_diagnosis == 2010 ~ Month_of_diagnosis_recode + 48,
                          Year_of_diagnosis == 2011 ~ Month_of_diagnosis_recode + 60,
                          Year_of_diagnosis == 2012 ~ Month_of_diagnosis_recode + 72,
                          Year_of_diagnosis == 2013 ~ Month_of_diagnosis_recode + 84))

mcaid0to39$SEER_DateofDeath_Month <- as.numeric(mcaid0to39$SEER_DateofDeath_Month)

mcaid0to39 <- mcaid0to39 %>%
  mutate(dthmo = case_when(SEER_DateofDeath_Year == 2006 ~ SEER_DateofDeath_Month + 0,
                          SEER_DateofDeath_Year  == 2007 ~ SEER_DateofDeath_Month + 12,
                          SEER_DateofDeath_Year  == 2008 ~ SEER_DateofDeath_Month + 24,
                          SEER_DateofDeath_Year  == 2009 ~ SEER_DateofDeath_Month + 36,
                          SEER_DateofDeath_Year  == 2010 ~ SEER_DateofDeath_Month + 48,
                          SEER_DateofDeath_Year  == 2011 ~ SEER_DateofDeath_Month + 60,
                          SEER_DateofDeath_Year  == 2012 ~ SEER_DateofDeath_Month + 72,
                          SEER_DateofDeath_Year  == 2013 ~ SEER_DateofDeath_Month + 84,
                          SEER_DateofDeath_Year  == 2014 ~ SEER_DateofDeath_Month + 96,
                          SEER_DateofDeath_Year  == 2015 ~ SEER_DateofDeath_Month + 108,
                          SEER_DateofDeath_Year  == 2016 ~ SEER_DateofDeath_Month + 120,
                          SEER_DateofDeath_Year  == 2017 ~ SEER_DateofDeath_Month + 132,
                          SEER_DateofDeath_Year  == 2018 ~ SEER_DateofDeath_Month + 144)) 
```

# Step A5b
- define earlylate
```{r}
mcaid0to39 <- mcaid0to39 %>%
  mutate(earlylate = if_else(dxmo<7| dxmo>90, 1, 0))

# check
table(mcaid0to39$earlylate, useNA = "always")
```

# Step A5c
- define status of enrollment during diagnosis month variable
```{r}
mcaid0to39 <- mcaid0to39 %>%
  rowwise() |>
  mutate(status = rowSums(dplyr::pick(starts_with("mcaid_"))[, dxmo])) |> # this code finds the value in the mcaid variable matching the diagnosis number and puts that value in a new variable called status
  ungroup() %>%
# recode NAs as 9
  mutate(status = case_when(is.na(status) ~ 9,
                              !is.na(status) ~ status))

# check
table(mcaid0to39$status, useNA = "always")
```

# Step A5d
- define vitalstat_post 6 to indicate whether they died within 6 months of diagnosis
```{r}
mcaid0to39 <- mcaid0to39 %>%
  mutate(dthmo_dxmo_diff = dthmo - dxmo) %>%
  mutate(vitalstat_post6 = if_else(dthmo_dxmo_diff <6, 1, 0, missing = 0)) # variable for selection of cases alive or who died but have 6 months of follow-up >0, missing = 0 codes missing as 0

# check
table(mcaid0to39$vitalstat_post6, useNA = "always")
```

# Validation: check status coding in random sample of 10 for each status 0, 1, 9
```{r}
set.seed(10)
status0 <- mcaid0to39 %>% filter(status == 0)
check_dx0 <- status0[sample(nrow(status0), 10),] # all validate
write.xlsx(check_dx0, file="/Users/kimjohnson/Desktop/Validation/status0.xlsx", showNA=TRUE, overwrite = TRUE)

set.seed(10)
status1 <- mcaid0to39 %>% filter(status == 1)
check_dx1 <- status1[sample(nrow(status1), 10),] # all validate
write.xlsx(check_dx1, file="/Users/kimjohnson/Desktop/Validation/status1.xlsx", showNA=TRUE, overwrite = TRUE)

set.seed(10)
status9 <- mcaid0to39 %>% filter(status == 9)
check_dx9 <- status9[sample(nrow(status9), 10),] # all validate
write.xlsx(check_dx9, file="/Users/kimjohnson/Desktop/Validation/status9.xlsx", showNA=TRUE, overwrite = TRUE)
```

# B: Continuity coding

# Create T1_pre to T6_post variables for creation of pre_mo6, post_mo6, any_zeros_pre and any_zeros_post variables in B
```{r}
# make T1 to T6 pre and post variables to use below to identify values for each mcaid variable in the 6 months pre and post dx
mcaid0to39 <- mcaid0to39 %>% 
  mutate(T1_pre = dxmo-6) %>%
  mutate(T2_pre = dxmo-5) %>%
  mutate(T3_pre = dxmo-4) %>%
  mutate(T4_pre = dxmo-3) %>%
  mutate(T5_pre = dxmo-2) %>%
  mutate(T6_pre = dxmo-1) %>%
  mutate(T1_post = dxmo+1) %>%
  mutate(T2_post = dxmo+2) %>%
  mutate(T3_post = dxmo+3) %>%
  mutate(T4_post = dxmo+4) %>%
  mutate(T5_post = dxmo+5) %>%
  mutate(T6_post = dxmo+6)

# assign all NAs for mcaid variables a value of 999 because rowwise/rowSums functions won't work with NAs
mcaid0to39df <- mcaid0to39 %>%
  select(patient_id, mcaid_1:mcaid_96) %>%
  mutate_all(~replace(., is.na(.), 999))

# merge back in with mcaid0to39
mcaid0to39 <- left_join(mcaid0to39df, mcaid0to39 %>% select(-c(mcaid_1:mcaid_96)), by = "patient_id")
```

# B. 
# B1. Non-special cases (not diagnosed in 1-6 or 91-96 and did not die within 6 month of dx)
- define pre_mo6 and post_mo6 period variables by summing values of mcaid variables 6 months pre and 6 months post dx
- define any_zero_pre and any_zero_post
```{r}
mcaid0to39_df1 <- mcaid0to39 %>%
  filter(vitalstat_post6 == 0 & earlylate == 0) %>% 
  select(patient_id, mcaid_1:mcaid_96, status, T1_pre, T2_pre, T3_pre, T4_pre, T5_pre, T6_pre,
         T1_post, T2_post, T3_post, T4_post, T5_post, T6_post, dxmo, dthmo) %>%
  rowwise() |>
  mutate(pre_mo6 = rowSums(dplyr::pick(starts_with("mcaid_"))[, T1_pre:T6_pre])) |> 
  mutate(post_mo6 = rowSums(dplyr::pick(starts_with("mcaid_"))[, T1_post:T6_post])) |> 
  ungroup() %>%

# anything greater than 6 assign an NA (this is because there were 999s)
  mutate(pre_mo6 = case_when(pre_mo6 > 6 ~ NA,
                               pre_mo6 <=6 ~ pre_mo6)) %>%
  mutate(post_mo6 = case_when(post_mo6 > 6 ~ NA,
                               post_mo6 <=6 ~ post_mo6))

# Defining variables that are used to determine whether there are any zeros pre and post diagnosis 6 month periods
# This gets values for each mcaid variable 6 mnths pre and post
mcaid0to39_df2 <- mcaid0to39_df1 %>%
  rowwise() |>
  mutate(T1_pre_val = rowSums(dplyr::pick(starts_with("mcaid_"))[, T1_pre])) |> 
  mutate(T2_pre_val = rowSums(dplyr::pick(starts_with("mcaid_"))[, T2_pre])) |>
  mutate(T3_pre_val = rowSums(dplyr::pick(starts_with("mcaid_"))[, T3_pre])) |>
  mutate(T4_pre_val = rowSums(dplyr::pick(starts_with("mcaid_"))[, T4_pre])) |>
  mutate(T5_pre_val = rowSums(dplyr::pick(starts_with("mcaid_"))[, T5_pre])) |>
  mutate(T6_pre_val = rowSums(dplyr::pick(starts_with("mcaid_"))[, T6_pre])) |>

  mutate(T1_post_val = rowSums(dplyr::pick(starts_with("mcaid_"))[, T1_post])) |> 
  mutate(T2_post_val = rowSums(dplyr::pick(starts_with("mcaid_"))[, T2_post])) |>
  mutate(T3_post_val = rowSums(dplyr::pick(starts_with("mcaid_"))[, T3_post])) |>
  mutate(T4_post_val = rowSums(dplyr::pick(starts_with("mcaid_"))[, T4_post])) |>
  mutate(T5_post_val = rowSums(dplyr::pick(starts_with("mcaid_"))[, T5_post])) |>
  mutate(T6_post_val = rowSums(dplyr::pick(starts_with("mcaid_"))[, T6_post])) |>
  ungroup() %>%

  # this determines whether there are any zeros pre_mo and post mo
  mutate(any_zero_pre = if_else((T1_pre_val == 0 | T2_pre_val == 0 | T3_pre_val == 0 |
                                  T4_pre_val == 0 | T5_pre_val == 0 | T6_pre_val == 0), 1, 0 )) %>%
  mutate(any_zero_post = if_else((T1_post_val == 0 | T2_post_val == 0 | T3_post_val == 0 |
                                  T4_post_val == 0 | T5_post_val == 0 | T6_post_val == 0), 1, 0 ))
```

# Validate pre_mo6, post_mo6, and any_zero_pre/any_zero_post in a sample of 10
```{r}
set.seed(10)
check_df2a <- mcaid0to39_df2[sample(nrow(mcaid0to39_df2), 10),] # all variables validate
write.xlsx(check_df2a, file="/Users/kimjohnson/Desktop/Validation/df2a_vars.xlsx", showNA=TRUE, overwrite = TRUE)
```

# Now we can define continuity for this group (vitalstat_post6 == 0 & earlylate == 0)
```{r}
mcaid0to39_df2 <- mcaid0to39_df2 %>%

# continuouus
mutate(mcontb = case_when(pre_mo6 == 6 & post_mo6 ==6 & status == 1 ~ 0,

# discontinuous
                          status ==0 | any_zero_pre == 1 | any_zero_post == 1 ~ 1)) %>% 
# other
mutate(mcontb = case_when (is.na(mcontb) ~ 2, # code rest as other
                            !is.na(mcontb) ~ mcontb)) # size is 43832
```

# Validate mcontb for this set in a sample of 10 each for mcontb = 0, 1, 2
```{r}
set.seed(10)
check_df2_0 <- mcaid0to39_df2 %>% filter(mcontb ==0)
check_df2_0 <- check_df2_0[sample(nrow(check_df2_0), 10),] # validates
write.xlsx(check_df2_0, file="/Users/kimjohnson/Desktop/Validation/df2_mcontb0.xlsx", showNA=TRUE, overwrite = TRUE)

set.seed(10)
check_df2_1 <- mcaid0to39_df2 %>% filter(mcontb == 1)
check_df2_1 <- check_df2_1[sample(nrow(check_df2_1), 10),] # validates
write.xlsx(check_df2_1, file="/Users/kimjohnson/Desktop/Validation/df2_mcontb1.xlsx", showNA=TRUE, overwrite = TRUE)

set.seed(10)
check_df2_2 <- mcaid0to39_df2 %>% filter(mcontb == 2)
check_df2_2 <- check_df2_2[sample(nrow(check_df2_2), 10),] # validates
write.xlsx(check_df2_2, file="/Users/kimjohnson/Desktop/Validation/df2_mcontb2.xlsx", showNA=TRUE, overwrite = TRUE)
```

# B. Special cases
# B2. died within 6 months dx date not earlylate 
```{r}
mcaid0to39_df3 <- mcaid0to39 %>%
  filter(vitalstat_post6 == 1 & earlylate != 1) %>% # selecting these cases
  rowwise() |>
  # getting their enrollment values prior to dx
  mutate(T1_pre_val = rowSums(dplyr::pick(starts_with("mcaid_"))[, T1_pre])) |> 
  mutate(T2_pre_val = rowSums(dplyr::pick(starts_with("mcaid_"))[, T2_pre])) |>
  mutate(T3_pre_val = rowSums(dplyr::pick(starts_with("mcaid_"))[, T3_pre])) |>
  mutate(T4_pre_val = rowSums(dplyr::pick(starts_with("mcaid_"))[, T4_pre])) |>
  mutate(T5_pre_val = rowSums(dplyr::pick(starts_with("mcaid_"))[, T5_pre])) |>
  mutate(T6_pre_val = rowSums(dplyr::pick(starts_with("mcaid_"))[, T6_pre])) |>

  ungroup() %>%

  # defining any zeros pre_mo
  mutate(any_zero_pre = if_else((T1_pre_val == 0 | T2_pre_val == 0 | T3_pre_val == 0 |
                                  T4_pre_val == 0 | T5_pre_val == 0 | T6_pre_val == 0), 1, 0 )) %>%
  select(patient_id, mcaid_1:mcaid_96, status, status, T1_pre, T2_pre, T3_pre, T4_pre, T5_pre, T6_pre, T1_post, T2_post, T3_post, T4_post, T5_post, T6_post, T1_pre_val, T2_pre_val, T3_pre_val, T4_pre_val, T5_pre_val, T6_pre_val, earlylate, dxmo, dthmo, dthmo_dxmo_diff, any_zero_pre) %>%
  rowwise() |>

  # getting sum of number of months enrolled pre
  mutate(pre_mo6 = rowSums(dplyr::pick(starts_with("mcaid_"))[, T1_pre:T6_pre])) |> 
  ungroup() %>%

  # anything greater than 6 assign an NA (because of addition of 999s)
  mutate(pre_mo6 = case_when(pre_mo6 > 6 ~ NA,
                               pre_mo6 <=6 ~ pre_mo6)) %>%  
  rowwise() |>

  mutate(sum_T1_post_todth = rowSums(dplyr::pick(starts_with("M"))[, T1_post:dthmo])) |> # this finds the mcaid variable for the month after the death month and sums through the mcaid variable for the death month
  ungroup() %>% 
  mutate(sum_T1_post_todth = case_when(sum_T1_post_todth <= 5 ~ sum_T1_post_todth, # assign back to NA because of 999's
                                       sum_T1_post_todth >5 ~ NA)) %>%
  # continuous
  mutate(mcontb = case_when(pre_mo6 == 6 & status == 1 & (sum_T1_post_todth >=dthmo-dxmo|dxmo == dthmo) ~ 0,

  # discontinuous
                            any_zero_pre == 1 | status ==0 | sum_T1_post_todth < dthmo-dxmo ~ 1)) %>%

  # other
  mutate(mcontb = case_when (is.na(mcontb) ~ 2, # code rest as other
                           !is.na(mcontb) ~ mcontb)) 
```

# validate
```{r}
set.seed(10)
check_df3_0 <- mcaid0to39_df3 %>% filter(mcontb ==0)
check_df3_0 <- check_df3_0[sample(nrow(check_df3_0), 10),] # validates
write.xlsx(check_df3_0, file="/Users/kimjohnson/Desktop/Validation/df3_mcontb0.xlsx", showNA=TRUE, overwrite = TRUE)

set.seed(10)
check_df3_1 <- mcaid0to39_df3 %>% filter(mcontb == 1)
check_df3_1 <- check_df3_1[sample(nrow(check_df3_1), 10),] # validates
write.xlsx(check_df3_1, file="/Users/kimjohnson/Desktop/Validation/df3_mcontb1.xlsx", showNA=TRUE, overwrite = TRUE)

set.seed(10)
check_df3_2 <- mcaid0to39_df3 %>% filter(mcontb == 2)
check_df3_2 <- check_df3_2[sample(nrow(check_df3_2), 10),] # validates
write.xlsx(check_df3_2, file="/Users/kimjohnson/Desktop/Validation/df3_mcontb2.xlsx", showNA=TRUE, overwrite = TRUE)
```

# B3. Did not die or did not die within 6 months of dxmo and dxmo is in 1 to 6 
```{r}
mcaid0to39_df4 <- mcaid0to39 %>%
  filter(earlylate == 1 & dxmo %in% c(1:6) & vitalstat_post6 ==0)  %>%
  rowwise() |>
  mutate(post_mo6 = rowSums(dplyr::pick(starts_with("mcaid_"))[, T1_post:T6_post])) |> 
  ungroup() %>%

  # anything greater than 6 assign an NA (because of addition of 999s)
  mutate(post_mo6 = case_when(post_mo6 > 6 ~ NA,
                               post_mo6 <=6 ~ post_mo6)) %>%  
  select(patient_id, mcaid_1:mcaid_13, dxmo, dthmo, status, post_mo6,  vitalstat_post6) %>%

  # discontinuous
  mutate(mcontb = case_when(vitalstat_post6 ==0 & dxmo == 1 & (status == 0 | post_mo6 <6) ~ 1,
                            vitalstat_post6 ==0 & dxmo == 2 & (mcaid_1 == 0 | status == 0) |post_mo6 <6 ~ 1,
                            vitalstat_post6 ==0 & dxmo == 3 & (mcaid_1 == 0 | mcaid_2 == 0 | status == 0) | post_mo6 <6 ~ 1,
                            vitalstat_post6 ==0 & dxmo == 4 & (mcaid_1 == 0 | mcaid_2 == 0 | mcaid_3 == 0 |status == 0) |post_mo6 <6 ~ 1,
                            vitalstat_post6 ==0 & dxmo == 5 & (mcaid_1 == 0 | mcaid_2 ==0 | mcaid_3 == 0 | mcaid_4 == 0 |status == 0) |post_mo6 <6 ~ 1,
                            vitalstat_post6 ==0 & dxmo == 6 & (mcaid_1 == 0 | mcaid_2 ==0 | mcaid_3 == 0 | mcaid_4 == 0 | mcaid_5 == 0 |status == 0) |post_mo6 <6 ~ 1)) %>% 
  # other
                             mutate(mcontb = case_when(is.na(mcontb) ~ 2,
                            !is.na(mcontb) ~ mcontb))
```

# Validate
```{r}
set.seed(10)
check_df4_1 <- mcaid0to39_df4 %>% filter(mcontb == 1)
check_df4_1 <- check_df4_1[sample(nrow(check_df4_1), 10),] # validates
write.xlsx(check_df4_1, file="/Users/kimjohnson/Desktop/Validation/df4_mcontb1.xlsx", showNA=TRUE, overwrite = TRUE) # Validated

set.seed(10)
check_df4_2 <- mcaid0to39_df4 %>% filter(mcontb == 2)
check_df4_2 <- check_df4_2[sample(nrow(check_df4_2), 10),] # validates
write.xlsx(check_df4_2, file="/Users/kimjohnson/Desktop/Validation/df4_mcontb2.xlsx", showNA=TRUE, overwrite = TRUE) # Validated
```

# B4. Died within 6 months of dxmo and dxmo is in 1 to 6
```{r}
mcaid0to39_df4b <- mcaid0to39 %>%
  filter(earlylate == 1 & dxmo %in% c(1:6) & vitalstat_post6 ==1)  %>%
  rowwise() |>
  mutate(post_mo6 = rowSums(dplyr::pick(starts_with("mcaid_"))[, T1_post:T6_post])) |> 
  ungroup() %>%

  # anything greater than 6 assign an NA (because of addition of 999s)
  mutate(post_mo6 = case_when(post_mo6 > 6 ~ NA,
                               post_mo6 <=6 ~ post_mo6)) %>%  
  select(patient_id, mcaid_1:mcaid_13, dxmo, dthmo, status, vitalstat_post6, T1_post:T6_post) %>%

  rowwise() |>
  mutate(sum_T1_post_todth = rowSums(dplyr::pick(starts_with("M"))[, T1_post:dthmo])) |>  # this finds the mcaid variable for the month after the diagnosis month and sums through the mcaid variable for the death month
  ungroup() %>%
  mutate(sum_T1_post_todth = case_when(sum_T1_post_todth <= 5 ~ sum_T1_post_todth, # assign back to NA because of 999's
                                       sum_T1_post_todth >5 ~ NA)) %>%

  # discontinuous
  mutate(mcontb = case_when(dxmo == 1 & (status ==0 | (sum_T1_post_todth < dthmo-dxmo)) ~ 1, 
                                      dxmo == 2 & (mcaid_1 == 0 | status == 0 | (sum_T1_post_todth < dthmo-dxmo)) ~ 1,
                                      dxmo == 3 & (mcaid_1 == 0 | mcaid_2 == 0 | status == 0 | (sum_T1_post_todth < dthmo-dxmo)) ~ 1,
                                      dxmo == 4 & (mcaid_1 == 0 | mcaid_2 == 0 | mcaid_3 == 0 | status == 0 |  (sum_T1_post_todth < dthmo-dxmo))~ 1,
                                      dxmo == 5 & (mcaid_1 == 0 | mcaid_2 == 0 | mcaid_3 ==0 | mcaid_4 ==0 | status ==0 |  (sum_T1_post_todth < dthmo-dxmo)) ~ 1,
                                      dxmo == 6 & (mcaid_1 == 0 | mcaid_2 ==0 | mcaid_3 == 0 | mcaid_4 ==0 | mcaid_5 == 0 | status ==0 | (sum_T1_post_todth < dthmo-dxmo)) ~ 1)) %>%

  # other                      
  mutate(mcontb = case_when(is.na(mcontb) ~ 2,
                           !is.na(mcontb) ~ mcontb)) 
```

# Validate
```{r}
set.seed(10)
check_df4_1b <- mcaid0to39_df4b %>% filter(mcontb == 1)
check_df4_1b <- check_df4_1b[sample(nrow(check_df4_1b), 10),] # validates
write.xlsx(check_df4_1b, file="/Users/kimjohnson/Desktop/Validation/df4b_mcontb1.xlsx", showNA=TRUE, overwrite = TRUE) # Validated

set.seed(10)
check_df4_2b <- mcaid0to39_df4b %>% filter(mcontb == 2)
check_df4_2b <- check_df4_2b[sample(nrow(check_df4_2b), 10),] # validates
write.xlsx(check_df4_2b, file="/Users/kimjohnson/Desktop/Validation/df4b_mcontb2.xlsx", showNA=TRUE, overwrite = TRUE) # Validated
```

# B5. Diagnosed in months 91:96 and didn't die in months 91 to 96 or died after month 96
```{r}
mcaid0to39_df5<- mcaid0to39 %>%
  filter((vitalstat_post6 ==0 | dthmo >96) & earlylate == 1 & dxmo %in% c(91:96)) %>%
  rowwise() |>
  mutate(pre_mo6 = rowSums(dplyr::pick(starts_with("mcaid_"))[, T1_pre:T6_pre])) |> 
  ungroup() %>%
  # anything greater than 6 assign an NA (because of addition of 999s)
  mutate(pre_mo6 = case_when(pre_mo6 > 6 ~ NA,
                               pre_mo6 <=6 ~ pre_mo6)) %>%

  select(patient_id, mcaid_82:mcaid_96, dxmo, status, pre_mo6) %>% # if any of mcaid up to dx month  = 0 then discontinuous, need to do this for each dx month occurring 91 to 96

  # discontinuous
  mutate(mcontb = case_when(dxmo == 96 & (status == 0 | pre_mo6 <6) ~ 1,
                            dxmo == 95 & (mcaid_96 == 0 | pre_mo6 <6) ~ 1,
                            dxmo == 94 & (mcaid_96 == 0 | mcaid_95 == 0 | status == 0 | pre_mo6<6) ~ 1,
                            dxmo == 93 & (mcaid_96 == 0 | mcaid_95 == 0 | mcaid_94 == 0 | status == 0 |pre_mo6<6)~ 1,
                            dxmo == 92 & (mcaid_96 == 0 | mcaid_95 == 0 | mcaid_94 == 0 | mcaid_93 == 0 |status == 0 | pre_mo6) <6 ~ 1,
                            dxmo == 91 & (mcaid_96 == 0 | mcaid_95 == 0 | mcaid_94 == 0 | mcaid_93 == 0 | mcaid_92 == 0 |status == 0 | pre_mo6<6) ~ 1)) %>%

  # other
  mutate(mcontb = case_when(is.na(mcontb) ~ 2,
                            !is.na(mcontb) ~ mcontb))

```

# Validate
```{r}
set.seed(10)
check_df5_1 <- mcaid0to39_df5 %>% filter(mcontb == 1)
check_df5_1 <- check_df5_1[sample(nrow(check_df5_1), 10),] # validates
write.xlsx(check_df5_1, file="/Users/kimjohnson/Desktop/Validation/df5_mcontb1.xlsx", showNA=TRUE, overwrite = TRUE) # Validated

set.seed(10)
check_df5_2 <- mcaid0to39_df5 %>% filter(mcontb == 2)
check_df5_2 <- check_df5_2[sample(nrow(check_df5_2), 10),] # validates
write.xlsx(check_df5_2, file="/Users/kimjohnson/Desktop/Validation/df5_mcontb2.xlsx", showNA=TRUE, overwrite = TRUE) # Validated
```

# B6. Diagnosed in months 91:96 and died in months 91 to 96
```{r}
mcaid0to39_df6<- mcaid0to39 %>%
  filter((vitalstat_post6 == 1 & dthmo <=96)  & earlylate == 1 & dxmo %in% c(91:96)) %>%
  rowwise() |>
  mutate(pre_mo6 = rowSums(dplyr::pick(starts_with("mcaid_"))[, T1_pre:T6_pre])) |> 
  ungroup() %>%
  # anything greater than 6 assign an NA (because of addition of 999s)
  mutate(pre_mo6 = case_when(pre_mo6 > 6 ~ NA,
                               pre_mo6 <=6 ~ pre_mo6)) %>%

  select(patient_id, mcaid_82:mcaid_96, dxmo, status, pre_mo6, dthmo) %>%

  mutate(mcontb = case_when(status == 0 |pre_mo6 < 6 ~ 1,
                            status == 1 & pre_mo6 == 6 & (dxmo == dthmo) ~ 0)) %>%
# code NAs

  mutate(mcontb = case_when(!is.na(mcontb) ~ mcontb,
                            # discontinuous
                            dthmo == 91 & (pre_mo6 <6 | status == 0) & is.na(mcontb) ~ 1,
                            dthmo == 92 & (pre_mo6 <6 | status == 0 | mcaid_92 == 0) &  is.na(mcontb) ~ 1,
                            dthmo == 93 & (pre_mo6 <6 | status == 0 | mcaid_92 == 0| mcaid_93 ==0) &  is.na(mcontb) ~ 1,
                            dthmo == 94 & (pre_mo6 <6 | status == 0 | mcaid_92 == 0| mcaid_93 ==0 | mcaid_94 ==0) &  is.na(mcontb) ~ 1,
                            dthmo == 95 & (pre_mo6 <6 | status == 0 | mcaid_92 == 0| mcaid_93 ==0 | mcaid_94 ==0 | mcaid_95 == 0) &  is.na(mcontb) ~ 1,
                            dthmo == 96 & (pre_mo6 <6 | status == 0 | mcaid_92 == 0| mcaid_93 ==0 | mcaid_94 ==0 | mcaid_95 == 0 | mcaid_96 == 0) &  is.na(mcontb) ~ 1,

                            # continuous
                            dthmo == 91 & (pre_mo6 == 6 & status == 1) & is.na(mcontb) ~ 0,
                            dthmo == 92 & (pre_mo6 == 6 & status == 1 & mcaid_92 == 1) &  is.na(mcontb) ~ 0,
                            dthmo == 93 & (pre_mo6 == 6 & status == 1 & mcaid_92 == 1| mcaid_93 == 1) &  is.na(mcontb) ~ 0,
                            dthmo == 94 & (pre_mo6 == 6 & status == 1 & mcaid_92 == 1| mcaid_93 == 1 & mcaid_94 ==1) &  is.na(mcontb) ~ 0,
                            dthmo == 95 & (pre_mo6 == 6 & status == 1 & mcaid_92 == 1| mcaid_93 == 1 & mcaid_94 ==1 & mcaid_95 == 1) &  is.na(mcontb) ~ 0,
                            dthmo == 96 & (pre_mo6 == 6 & status == 1 & mcaid_92 == 1| mcaid_93 == 1 & mcaid_94 ==1 & mcaid_95 == 1 | mcaid_96 == 1) &  is.na(mcontb) ~ 0)) %>%

                            # other
                            mutate(mcontb = case_when(is.na(mcontb) ~2,
                                                      !is.na(mcontb) ~ mcontb))
```

# Validate
```{r}
set.seed(10)
check_df6_0 <- mcaid0to39_df6 %>% filter(mcontb == 0)
check_df6_0 <- check_df6_0[sample(nrow(check_df6_0), 10),] # validates
write.xlsx(check_df6_0, file="/Users/kimjohnson/Desktop/Validation/df6_mcontb0.xlsx", showNA=TRUE, overwrite = TRUE) # Validated

set.seed(10)
check_df6_1 <- mcaid0to39_df6 %>% filter(mcontb == 1)
check_df6_1 <- check_df6_1[sample(nrow(check_df6_1), 10),] # validates
write.xlsx(check_df6_1, file="/Users/kimjohnson/Desktop/Validation/df6_mcontb1.xlsx", showNA=TRUE, overwrite = TRUE) # Validated

set.seed(10)
check_df6_2 <- mcaid0to39_df6 %>% filter(mcontb == 2)
check_df6_2 <- check_df6_2 [sample(nrow(check_df6_2 ), 10),] # validates
write.xlsx(check_df6_2, file="/Users/kimjohnson/Desktop/Validation/df6_mcontb2.xlsx", showNA=TRUE, overwrite = TRUE) # Validated
```

# need to merge it all back together after validation above
```{r}
# prepare dfs for merging mcaid0to39_df2, df3, df4, df5, df6

mcaid0to39_df2 <- mcaid0to39_df2 %>% 
  select(patient_id, mcontb)

mcaid0to39_df3 <- mcaid0to39_df3 %>% 
  select(patient_id, mcontb)

mcaid0to39_df4 <- mcaid0to39_df4 %>% 
  select(patient_id, mcontb)

mcaid0to39_df4b <- mcaid0to39_df4b %>% 
  select(patient_id, mcontb)

mcaid0to39_df5 <- mcaid0to39_df5 %>% 
  select(patient_id, mcontb)

mcaid0to39_df6 <- mcaid0to39_df6 %>% 
  select(patient_id, mcontb)

mcaid0to39 <- rbind(mcaid0to39_df2, mcaid0to39_df3, mcaid0to39_df4, mcaid0to39_df4b, mcaid0to39_df5, mcaid0to39_df6)

# export for use in analysis 
write.xlsx(mcaid0to39, file="/Volumes/kijohnson/Active/mcaid0to39KJ.xlsx", showNA=TRUE, overwrite = TRUE)
```

### 2023-09-06 Manuscript Analysis code 0 to 19.Rmd
initial commit of paper coding files

---
title: "SEER Medicaid"
author: "Kim Johnson"
date: '2022-06-06'
output: html_document
---

# Open libraries
```{r}
# install.packages("pacman")
pacman::p_load(readr, dplyr, tidyverse, lubridate, survival, survminer, hablar, table1, patchwork, openxlsx, DiagrammeR, lmtest, forestplot, MASS, nnet, broom, forcats, sandwich, mlogit, sjstats, odds.n.ends, labelled, readxl)
```
# Functions used
```{r}
# function to calculate clustered SEs (from: https://stackoverflow.com/questions/73136246/clustered-standard-errors-stars-and-summary-statistics-in-modelsummary-for-mul)
mlogit.clust <- function(model,data,variable) {
  beta <- c(t(coef(model)))
  vcov <- vcov(model)
  k <- length(beta)
  n <- nrow(data)
  max_lev <- length(model$lev)
  xmat <- model.matrix(model)
  # u is deviance residuals times model.matrix
  u <- lapply(2:max_lev, function(x)
    residuals(model, type = "response")[, x] * xmat)
  u <- do.call(cbind, u)
  m <- dim(table(data[,variable]))
  u.clust <- matrix(NA, nrow = m, ncol = k)
  fc <- factor(data[,variable])
  for (i in 1:k) {
    u.clust[, i] <- tapply(u[, i], fc, sum)
  }
  cl.vcov <- vcov %*% ((m / (m - 1)) * t(u.clust) %*% (u.clust)) %*% vcov
  return(cl.vcov = cl.vcov)
}
```
# Import files sent by SEER Medicaid (IMS), get data into right columns, and assign column names
```{r, eval =FALSE}
# # Medicaid enrollment--imports fixed width columns and gives them names
# enroll <- read_fwf('medicaid.enroll.txt', 
#                fwf_cols(patient_id = 15, STATE_CD = 2, YEAR = 4, DOB_MATCH_SCORE_HI = 1, SEX_MATCH_SCORE_HI = 1,
#                        ID_MATCH_SCORE_HI=1, MATCH_SCORE_HI=2, MEDICAID_ELIG_CD1=1, MEDICAID_ELIG_CD2=1,
#                       MEDICAID_ELIG_CD3=1, MEDICAID_ELIG_CD4=1, MEDICAID_ELIG_CD5=1, MEDICAID_ELIG_CD6=1,
#                       MEDICAID_ELIG_CD7=1, MEDICAID_ELIG_CD8=1, MEDICAID_ELIG_CD9=1, MEDICAID_ELIG_CD10=1,
#                         MEDICAID_ELIG_CD11=1, MEDICAID_ELIG_CD12=1), 
#                skip = 0)

# # SEER file (2 is the second dataseet received that includes insurance variables)
# 
# seer<-read_fwf('medicaid.cancer 2.txt', na=".", fwf_positions(
# c(1, 16,18,20,21,22,24,25,28,31,33,35,39,41,45,46,50,51,55,56,57,58,59,62,64,66,67,69, 71, 72,73,74,75,76,77,78,79,80,81,82,83,84,86,88,90,92,94,95,96,97,98,99,100,103,106,109, 111,114,117,120,123,126,129,132,135,138,141,144,147,150,153,156, 159,161,163,165,167,168,174,180,186,188,189,190,192,193,194,195,196,197,199, 200,201,203,205,210,213,216,217,219,224,227,228,229,230,231,233,235,236,238,241,242, 243,248, 253, 254, 255,256, 258,263,264,265,266,267,268,269,270,271,273,279,282,285,288,291,293,295,297,299, 300,301,302,303,304,305,306,308,310,316,347,353,359,365,366,367,368,369,378,380,382,384,388,401,402,403,404,405, 406,408,412,414, 419, 423,425,427,429, 431,433,437,438,439,440,441,442,443,445, 447, 448, 452, 456, 460, 464,465,466,467,468,469,470, 471,472,473,474,475,476,477, 479, 480,485,490,495,500,505, 510,511,512,513,514,515,519,520, 521,524,525,526,530,532),

# c(15,17,19,20,21,23,24,27,30,32,34,38,40,44,45,49,50,54,55,56,57,58,61,63,65,66,68,70,71, 72,73,74,75,76,77,78,79,80,81,82,83,84,86,88,90,93,94,95,96,97,98, 99,102,105,108,110,113,116,119,122,125,128,131,134,137, 140, 143,146,149,152,155,158,160,162,164,166,167,173,179, 185,187,188,189,191, 192,193,194,195,195, 198,199, 200,202,204,209,212,215,216,218,220, 226,227,228,229,230,232,234,235,237,240,241,242,247, 252, 253, 254,255,257, 259, 263,264,265,266,267,268,269,270, 272,273,281,284,287,290,292,294,295,298,299, 300, 301,302,303,304,305,307, 309,311,316,352,358, 364,365,366,367,368,369, 379,381,383,387,391,401,402,403,404,405, 407,411,413,417, 422,424,426,428,430, 432,436,437,438,439,440,441,442,444,446, 447, 451, 455,459,463,464, 465, 466,467,468,469,470, 471,472,473,474,475,476,477,479, 484,489,494,499,504,509,510,511,512,513,514,515,519,520, 523,524,525,529,531,534)))

# colnames(seer)<- c("patient_id", "SEER_registry", "SEERregistrywithCAandGAaswholes", 
# "Louisiana20051stvs2ndhalfofyear", "Marital_
# _at_diagnosis", 
# "Race_ethnicity", "sex", "Agerecodewithsingleages_and_100", "agerecodewithsingle_ages_and_85", 
# "Sequence_number", "Month_of_diagnosis", "Year_of_diagnosis", 
# "Month_of_diagnosis_recode", "Primary_Site", "Laterality", "Histology_ICD_O_2", 
# "Behavior_code_ICD_O_2", "Histologic_Type_ICD_O_3", "Behavior_code_ICD_O_3", 
# "Grade", "Diagnostic_Confirmation", "Type_of_Reporting_Source", 
# "EOD_10_size_1988_2003", "EOD_10_extent_1988_2003", "EOD10Prostatepath_ext_1995_2003", 
# "EOD_10_nodes_1988_2003", "Regional_nodes_positive_1988", "Regional_nodes_examined_1988", 
# "Expanded_EOD_1_CP53_1973_1982", "Expanded_EOD_2_CP54_1973_1982", 
# "Expanded_EOD_3_CP55_1973_1982", "Expanded_EOD_4_CP56_1973_1982", 
# "Expanded_EOD_5_CP57_1973_1982", "Expanded_EOD_6_CP58_1973_1982", 
# "Expanded_EOD_7_CP59_1973_1982", "Expanded_EOD_8_CP60_1973_1982",
# "Expanded_EOD_9_CP61_1973_1982", "Expanded_EOD_10_CP62_1973_1982",
# "Expanded_EOD_11_CP63_1973_1982", "Expanded_EOD_12_CP64_1973_1982",
# "Expanded_EOD_13_CP65_1973_1982", "V2Digit_NS_EOD_part_1_1973_1982",
# "V2Digit_NS_EOD_part_2_1973_1982", "V2Digit_SS_EOD_part_1_1973_1982",
# "V2Digit_SS_EOD_part_2_1973_1982", "EOD_4_size_1983_1987", "EOD_4_extent_1983_1987",
# "EOD_4_nodes_1983_1987", "Coding_system_EOD_1973_2003", "Tumor_marker_1_1990_2003",
# "Tumor_marker_2_1990_2003", "Tumor_marker_3_1998_2003", "CS_tumor_size_2004_2015",
# "CS_extension_2004_2015", "CS_lymph_nodes_2004_2015", "CS_mets_at_dx_2004_2015",
# "CSsitespecificfactor12004varyin", "CSsitespecificfactor22004varyin",
# "CSsitespecificfactor32004varyin", "CSsitespecificfactor42004varyin",
# "CSsitespecificfactor52004varyin", "CSsitespecificfactor62004varyin",
# "CSsitespecificfactor72004varyin", "CSsitespecificfactor82004varyin",
# "CSsitespecificfactor92004varyin", "CSsitespecificfactor102004varyi",
# "CSsitespecificfactor112004varyi", "CSsitespecificfactor122004varyi",
# "CSsitespecificfactor132004varyi", "CSsitespecificfactor152004varyi",
# "CSsitespecificfactor162004varyi", "CSsitespecificfactor252004varyi",
# "Derived_AJCC_T_6th_ed_2004_2015", "Derived_AJCC_N_6th_ed_2004_2015",
# "Derived_AJCC_M_6th_ed_2004_2015", "DerivedAJCCStageGroup6thed20042",
# "SEERCombinedSummaryStage2000200", "CSversioninputoriginal2004_2015",
# "CS_version_derived_2004_2015", "CSversioninputcurrent_2004_2015",
# "RX_Summ_Surg_Prim_Site_1998", "RX_Summ_Scope_Reg_LN_Sur_2003",
# "RX_Summ_Surg_Oth_Reg_Dis_2003", "RXSummReg_LN_Examined_1998_2002",
# "RX_Summ_Systemic_Surg_Seq", "RX_Summ_Surg_Rad_Seq", "Reasonnocancer_directed_surgery",
# "Radiation_recode", "Chemotherapy_recode_yes_no_unk", "Sitespecificsurgery19731997vary",
# "Scopeofreglymphndsurg_1998_2002", "Surgeryofothregdissites19982002",
# "Record_number_recode", "Age_recode_with_1_year_olds", "Site_recode_ICD_O_3_WHO_2008",
# "ICCCsiterecode_ICD_O_3_WHO_2008", "ICCCsiterecextendedICDO3WHO2008",
# "Behavior_recode_for_analysis", "Histologyrecode_broad_groupings",
# "Histologyrecode_Brain_groupings", "TNM_7_CS_v0204_Schema", "Race_recode_White_Black_Other",
# "Race_recode_W_B_AI_API", "OriginrecodeNHIAHispanicNonHisp",
# "SEER_historic_stage_A_1973_2015", "AJCCstage_3rd_edition_1988_2003",
# "SEERmodifiedAJCCstage3rd1988200", "Firstmalignantprimary_indicator",
# "state", "county", "PRCDA_2017", "PRCDA_Region", "COD_to_site_recode",
# "COD_to_site_rec_KM", "Vital
# recodestudycutoffuse", "IHS_Link",
# "Summary_stage_2000_1998", "AYA_site_recode_WHO_2008", "Lymphomasubtype_recode_WHO_2008",
# "SEERcausespecificdeathclassific", "SEERothercauseofdeathclassifica",
# "CSTumor_Size_Ext_Eval_2004_2015", "CS_Reg_Node_Eval_2004_2015",
# "CS_Mets_Eval_2004_2015", "Primary_by_international_rules", "ER
# RecodeBreastCancer_1990",
# "PR
# RecodeBreastCancer_1990", "CS_Schema_AJCC_6th_Edition",
# "LymphvascularInvasion2004varyin", "Derived_AJCC_T_7th_ed_2010",
# "Derived_AJCC_N_7th_ed_2010", "Derived_AJCC_M_7th_ed_2010", "DerivedAJCCStageGroup7thed_2010",
# "BreastAdjustedAJCC6thT1988_2015", "BreastAdjustedAJCC6thN1988_2015",
# "BreastAdjustedAJCC6thM1988_2015", "BreastAdjustedAJCC6thStage19882",
# "Derived_HER2_Recode_2010", "Breast_Subtype_2010", "LymphomaAnnArborStage_1983_2015",
# "SEERCombinedMetsat_DX_bone_2010", "SEERCombinedMetsatDX_brain_2010",
# "SEERCombinedMetsatDX_liver_2010", "SEERCombinedMetsat_DX_lung_2010",
# "TvaluebasedonAJCC_3rd_1988_2003", "NvaluebasedonAJCC_3rd_1988_2003",
# "MvaluebasedonAJCC_3rd_1988_2003", "RadiationtoBrainorCNSRecode1988",
# "Census_Tract_1990", "Census_Tract_2000", "Census_Tract_2010",
# "Census_Coding_System", "Census_Tract_Certainty_1990", "Census_Tract_Certainty_2000",
# "Census_Tract_Certainty_2010", "Census_Tract_Poverty_Indicator",
# "Rural_Urban_Continuum_Code_1993", "Rural_Urban_Continuum_Code_2003",
# "Rural_Urban_Continuum_Code_2013", "Health_Service_Area", "HealthService_Area_NCI_Modified",
# "Derived_SS1977_flag", "Derived_SS2000_flag", "Firstcourseofreconstruct1998200",
# "Radiation", "RadiationtoBrainorCNS_1988_1997", "SEER_DateofDeath_Month",
# "SEER_DateofDeath_Year", "Month_of_last_follow_up_recode", "Year_of_last_follow_up_recode",
# "Year_of_birth", "Date_of_diagnosis_flag", "Date_therapy_started_flag",
# "Date_of_birth_flag", "Date_of_last_follow_up_flag", "Month_therapy_started",
# "Year_therapy_started", "Other_cancer_directed_therapy", "Derived_AJCC_flag",
# "Derived_SS1977", "Derived_SS2000", "SEER_Summary_stage_1977_9500",
# "SEER_Summary_stage_2000_0103", "NAPIIA_Derived_API_Race", "Primary_Payer_at_DX",
# "COD_ICD_code", "Cause_of_Death_ICD_8_or_9", "Cause_of_Death_ICD_10",
# "Recode_ICD_0_2_to_9", "Recode_ICD_0_2_to_10", "NHIA_Derived_Hisp_Origin",
# "Age_site_edit_override", "Sequencenumber_dx_conf_override",
# "Site_type_lat_seq_override", "Surgerydiagnostic_conf_override",
# "Site_type_edit_override", "Histology_edit_override", "Report_source_sequence_override",
# "Seq_ill_defined_site_override", "LeukLymphdxconfirmationoverride",
# "Site_behavior_override", "Site_EOD_dx_date_override", "Site_laterality_EOD_override",
# "Site_laterality_morph_override", "Insurance_Recode_2007", "Yost_ACS_2006_2010",
# "Yost_ACS_2010_2014", "Yost_ACS_2013_2017", "Yost_ACS_2006_2010_State_based",
# "Yost_ACS_2010_2014_State_based", "Yost_ACS_2013_2017_State_based",
# "Yost_ACS_2006_2010_quintile", "Yost_ACS_2010_2014_quintile",
# "Yost_ACS_2013_2017_quintile", "YostACS20062010quintileStatebas",
# "YostACS20102014quintileStatebas", "YostACS20132017quintileStatebas",
# "HPV_recode_2010", "In_Medicaid_Flag", "OncotypeDXBreastRecurrenceScore",
# "OncotypeDXRSgroupRS18RS1830RS30", "OncotypeDXreasonno_score_linked",
# "Oncotype_DX_year_of_test_report", "OncotypeDX_month_of_test_report",
# "OncotypeDXmonthssince_diagnosis")

# hit control shift c while highlighting above to uncomment/comment section
```

# Create dataframes for Medicaid enrollment and SEER for export that are limited to those diagnosed at 0 to 39 years 
# Exclude those over 39 years old and export data to make the datasets more manageeable on import
```{r}
# exclude those 40 and over from seer cancer filefirst
# seer_0to39<-seer %>%
#   filter(Agerecodewithsingleages_and_100<40)

# get ids from seer file to limit enrollment file to those ids

# ids<-seer_0to39$patient_id

# enroll_seerids<-enroll %>% # limit Medicaid enrollment file to SEER ids
 # filter(patient_id %in% ids)
```

# Export these data as csv files for quicker import
```{r}
# write.csv(seer_0to39,"seer0to39.csv", row.names = FALSE)

# write.csv(enroll_seerids,"enroll_seerids.csv", row.names = FALSE)
```

# Import files created with text commented out above for analysis
```{r}
seer0to39<-read_csv("/Volumes/kijohnson/Active/seer0to39.csv", show_col_types = FALSE) # this is 0 to 39 cancer patient data
enroll_seerids<-read_csv("/Volumes/kijohnson/Active/enroll_seerids.csv", show_col_types = FALSE) # this is 0 to 39 enrollee data
mcaid<- read_xlsx("/Volumes/kijohnson/Active/mcaid0to39KJ.xlsx") # created in continuity coding file
```

# Identify ids to exclude where patient has Medicaid files in multiple states (this is used below)
```{r}
df <- enroll_seerids %>% 
   find_duplicates(patient_id, YEAR) # this identifies patients in multiple states in the same year

x <- unique(df$patient_id) # 841 vector of unique ids who had enrollment in multiple states to exclude in exclusion section
```

# Reduce cancer dataset to variables needed for analysis
```{r}
# select variables
seer0to39<-seer0to39 %>%
  dplyr::select("patient_id", "SEER_registry",  "Race_ethnicity", "sex", "Agerecodewithsingleages_and_100",
"Sequence_number", "Month_of_diagnosis", "Year_of_diagnosis",
"Month_of_diagnosis_recode", "Primary_Site", "Laterality", "Histology_ICD_O_2",
"Behavior_code_ICD_O_2", "Histologic_Type_ICD_O_3", "Behavior_code_ICD_O_3",
"Grade", "Diagnostic_Confirmation", "Type_of_Reporting_Source", 
"Derived_AJCC_T_6th_ed_2004_2015", 
"Derived_AJCC_N_6th_ed_2004_2015",
"Derived_AJCC_M_6th_ed_2004_2015", 
"DerivedAJCCStageGroup6thed20042",
"SEERCombinedSummaryStage2000200", 
"Age_recode_with_1_year_olds", "Site_recode_ICD_O_3_WHO_2008",
"ICCCsiterecode_ICD_O_3_WHO_2008", "ICCCsiterecextendedICDO3WHO2008",
"Behavior_recode_for_analysis", "Histologyrecode_broad_groupings", "Race_recode_White_Black_Other",
"Race_recode_W_B_AI_API", "OriginrecodeNHIAHispanicNonHisp",
"SEER_historic_stage_A_1973_2015", "AJCCstage_3rd_edition_1988_2003",
"SEERmodifiedAJCCstage3rd1988200", "Firstmalignantprimary_indicator",
"state", "county", "PRCDA_2017", "PRCDA_Region", "COD_to_site_recode",
"COD_to_site_rec_KM", "Vitalstatusrecodestudycutoffuse", "IHS_Link",
"Summary_stage_2000_1998", "AYA_site_recode_WHO_2008", 
"SEERcausespecificdeathclassific", "SEERothercauseofdeathclassifica",
"Derived_AJCC_T_7th_ed_2010", "Derived_AJCC_N_7th_ed_2010", "Derived_AJCC_M_7th_ed_2010", "DerivedAJCCStageGroup7thed_2010", "Census_Tract_1990", "Census_Tract_2000", "Census_Tract_2010",
"Census_Coding_System", "Census_Tract_Certainty_1990", "Census_Tract_Certainty_2000",
"Census_Tract_Certainty_2010", "Census_Tract_Poverty_Indicator",
"Rural_Urban_Continuum_Code_1993", "Rural_Urban_Continuum_Code_2003",
"Rural_Urban_Continuum_Code_2013", "Health_Service_Area", "HealthService_Area_NCI_Modified",
"SEER_DateofDeath_Month", "SEER_DateofDeath_Year", "Month_of_last_follow_up_recode", "Year_of_last_follow_up_recode", "Year_of_birth", "Date_of_diagnosis_flag", "Date_therapy_started_flag",
"Date_of_birth_flag", "Date_of_last_follow_up_flag", "Month_therapy_started",
"Year_therapy_started",  "Derived_AJCC_flag",
"SEER_Summary_stage_1977_9500",
"SEER_Summary_stage_2000_0103", "NAPIIA_Derived_API_Race", "Primary_Payer_at_DX",
"COD_ICD_code", "Cause_of_Death_ICD_8_or_9", "Cause_of_Death_ICD_10",
"Recode_ICD_0_2_to_9", "Recode_ICD_0_2_to_10", "NHIA_Derived_Hisp_Origin",
"Insurance_Recode_2007", "Yost_ACS_2006_2010", "Yost_ACS_2010_2014", "Yost_ACS_2013_2017", "Yost_ACS_2006_2010_State_based", "Yost_ACS_2010_2014_State_based", "Yost_ACS_2013_2017_State_based",
"Yost_ACS_2006_2010_quintile", "Yost_ACS_2010_2014_quintile",
"Yost_ACS_2013_2017_quintile", "YostACS20062010quintileStatebas",
"YostACS20102014quintileStatebas", "YostACS20132017quintileStatebas",
"HPV_recode_2010", "In_Medicaid_Flag")
```

# Data management of variables 
```{r}
# recode demographic variables
# Age, race, hispanic, sex  cause of death, cause specfic death
seer0to39<-seer0to39 %>%
  mutate(age_cat=case_when(Age_recode_with_1_year_olds %in% c(0,1) ~ 0, # 0 to 4
                                Age_recode_with_1_year_olds == 2 ~1, # 5 to 9
                                Age_recode_with_1_year_olds == 3 ~2, # 10 to 14
                                Age_recode_with_1_year_olds == 4 ~3, # 15 to 19
                                Age_recode_with_1_year_olds %in% c(5,6,7,8) ~ 4)) %>% # 20 to 39

         mutate(race_cat=case_when(Race_recode_W_B_AI_API == 1 ~0, # White
                                Race_recode_W_B_AI_API == 2 ~1, # Black
                                Race_recode_W_B_AI_API == 3 ~2, # AIAN
                                Race_recode_W_B_AI_API == 4 ~3, # API
                                Race_recode_W_B_AI_API == 9 ~4)) %>% # Unknown

        mutate(hispanic = case_when(OriginrecodeNHIAHispanicNonHisp == 0 ~0, # Non-Hispanic
                                    OriginrecodeNHIAHispanicNonHisp == 1 ~1)) %>% # Hispanic

        mutate(race_ethnicity = case_when(race_cat == 0 & hispanic == 0 ~ 0, # Non-Hispanic White
                                race_cat == 1 & hispanic == 0 ~ 1, # Non-Hispanic Black
                                race_cat == 3 & hispanic == 0 ~ 2, # Non-Hispanic API
                                race_cat %in% c(2,4) & hispanic == 0 ~ 3, # Non-Hispanic Other
                                race_cat %in% c(0:4) & hispanic == 1 ~ 4)) %>% # Hispanic

        mutate(race_cat2 = case_when(race_cat == 0 ~ 0, # Non-Hispanic White
                                     race_cat == 1 ~ 1, # Non-Hispanic Black
                                     race_cat %in% c(2,3,4,5) ~ 2)) %>% # Non-Hispanic Other

        mutate(sex = case_when(sex == 1 ~ 0, # "Male"
                                sex == 2 ~ 1, # "Female"
                                sex == 9 ~ 2)) %>% # "Unknown"

        mutate(can_dead = case_when(SEERcausespecificdeathclassific == 0 ~0, # Alive or dead of other cause
                                SEERcausespecificdeathclassific == 1 ~1)) # Dead from cancer
# note 8 an 9 are missing or N/A

table(seer0to39$SEERcausespecificdeathclassific, useNA = "always")

# Factor variables
seer0to39$age_cat<-factor(seer0to39$age_cat, levels=c(0:4), labels=c("0-4", "5-9", "10-14", "15-19", "20-39"))
seer0to39$race_cat<-factor(seer0to39$race_cat, levels=c(0:4), labels=c("White", "Black", "AIAN", "API", "Unknown"))
seer0to39$hispanic<-factor(seer0to39$hispanic, levels=c(0:1), labels=c("Non-Hispanic", "Hispanic"))
seer0to39$race_ethnicity<-factor(seer0to39$race_ethnicity, levels=c(0:4), labels=c("Non-Hispanic White", "Non-Hispanic Black", "Non-Hispanic API", "Non-Hispanic Other",  "Hispanic")) 
seer0to39$sex <- factor(seer0to39$sex, levels = c(0:2), labels = c("Male", "Female", "Not Stated"))

# check variables
table(seer0to39$age_cat, useNA = "always")
table(seer0to39$race_cat, useNA = "always")
table(seer0to39$hispanic, useNA = "always")
table(seer0to39$race_ethnicity, useNA = "always")
table(seer0to39$can_dead, useNA = "always")
table(seer0to39$sex, useNA = "always")

# calculate follow-up time from month of dx, recode year of dx TO Month_of_last_follow_up_recode, Year_of_last_follow_up_recode. The difference between these variable dates would provide months of follow-up
summary(seer0to39$Month_of_diagnosis_recode)
summary(seer0to39$Year_of_diagnosis)
summary(as.numeric(seer0to39$Month_of_last_follow_up_recode))
seer0to39$Month_of_last_follow_up_recode<-as.numeric(seer0to39$Month_of_last_follow_up_recode) # make variable numeric
seer0to39$SEER_DateofDeath_Month<-as.numeric(seer0to39$SEER_DateofDeath_Month) # make variable numeric
summary(seer0to39$Month_of_last_follow_up_recode)
summary(seer0to39$Year_of_last_follow_up_recode)
summary(seer0to39$SEER_DateofDeath_Month)
summary(seer0to39$SEER_DateofDeath_Year)

# assign a date of diagnosis using 15 as the day because there is no day of dx so assume mean is 15
seer0to39$proxy_day_of_dx <-15
seer0to39$proxy_day_of_fu <-15
seer0to39$proxy_day_of_dth <-15

seer0to39<-seer0to39 %>%
  mutate(proxy_dx_date = make_date(Year_of_diagnosis, Month_of_diagnosis_recode, proxy_day_of_dx)) %>% # make date of dx
  mutate(proxy_fu_date = make_date(Year_of_last_follow_up_recode, Month_of_last_follow_up_recode, proxy_day_of_fu)) %>% # make date of follow-up
  mutate(proxy_dth_date = make_date(SEER_DateofDeath_Year, SEER_DateofDeath_Month, proxy_day_of_dth)) # make date of death

# calculate the difference in days between the estimated date of dx and estimated date of last follow-up for non-cases

# non-cases
seer0to39_noncases <- seer0to39 %>%
  filter(can_dead == 0|is.na(can_dead)) # used last follow-up for missing can_dead
seer0to39_noncases$survday<-as.numeric(difftime(seer0to39_noncases$proxy_fu_date, seer0to39_noncases$proxy_dx_date, units=c("days")))

# cases
seer0to39_cases <- seer0to39 %>%
  filter(can_dead == 1)
seer0to39_cases$survday <-as.numeric(difftime(seer0to39_cases$proxy_dth_date, seer0to39_cases$proxy_dx_date, units=c("days")))

summary(seer0to39_noncases$proxy_fu_date)
summary(seer0to39_cases$proxy_fu_date)

# bind these back together
seer0to39 <- rbind(seer0to39_cases, seer0to39_noncases)

# estimate follow-up months by dividing by the average number of days in a month
seer0to39$fumo<-seer0to39$survday/30.437

summary(seer0to39$fumo)

# cancer type coding
table(seer0to39$ICCCsiterecode_ICD_O_3_WHO_2008)

# ICCC3 (https://seer.cancer.gov/iccc/iccc-who2008.html)
seer0to39<-seer0to39 %>%
  mutate(can_type = case_when(
    ICCCsiterecode_ICD_O_3_WHO_2008 %in% c(11, 12,13,14,15)~0,
    ICCCsiterecode_ICD_O_3_WHO_2008 %in% c(21,22,23,24,25)~1,
    ICCCsiterecode_ICD_O_3_WHO_2008 %in% c(31,32,33,34,35,36)~2,
    ICCCsiterecode_ICD_O_3_WHO_2008 %in% c(41,42) ~3,
    ICCCsiterecode_ICD_O_3_WHO_2008 %in% c(50) ~4,
    ICCCsiterecode_ICD_O_3_WHO_2008 %in% c(61,62,63) ~5,
    ICCCsiterecode_ICD_O_3_WHO_2008 %in% c(71,72,73) ~6,
    ICCCsiterecode_ICD_O_3_WHO_2008 %in% c(81,82, 83,84,85) ~7,
    ICCCsiterecode_ICD_O_3_WHO_2008 %in% c(91,92,93,94,95)~8,
    ICCCsiterecode_ICD_O_3_WHO_2008 %in% c(101,102,103, 104,105) ~9,
    ICCCsiterecode_ICD_O_3_WHO_2008 %in% c(111,112,113,114,115,116) ~10,
    ICCCsiterecode_ICD_O_3_WHO_2008 %in% c(121,122)~11))

seer0to39$can_type<-factor(seer0to39$can_type, levels=c(0:11), labels=c(
  "I. Leukemias, myeloproliferative diseases, and myelodysplastic diseases",
  "II. Lymphomas and reticuloendothelial neoplasms",
  "III. CNS and miscellaneous intracranial and intraspinal neoplasms",
  "IV. Neuroblastoma and other peripheral nervous cell tumors",
  "V. Retinoblastoma",
  "VI. Renal tumors",
  "VII. Hepatic tumors",
  "VIII. Malignant bone tumors",
  "IX. Soft tissue and other extraosseous sarcomas",
  "X. Germ cell tumors, trophoblastic tumors, and neoplasms of gonads",
  "XI. Other malignant epithelial neoplasms and malignant melanomas",
  "XII. Other and unspecified malignant neoplasms"))

table(seer0to39$can_type, seer0to39$ICCCsiterecode_ICD_O_3_WHO_2008, useNA= "always")

table(seer0to39$AYA_site_recode_WHO_2008)

# use Yost index as a measure of poverty
table(seer0to39$Yost_ACS_2006_2010_quintile,  useNA = "always")

class(seer0to39$Yost_ACS_2006_2010_quintile)

seer0to39$Yost_ACS_2006_2010_quintile <-factor(seer0to39$Yost_ACS_2006_2010_quintile, levels = c(1:5), labels = c(
        "Group 1 (lowest SES)",
        "Group 2",
        "Group 3",
        "Group 4",
        "Group 5 (highest SES)"))

# classify stage
table(seer0to39$SEERCombinedSummaryStage2000200, useNA = "always")
seer0to39 <- seer0to39 %>%
  mutate(stage = case_when(SEERCombinedSummaryStage2000200 %in% c(0,1) ~ 0,
                           SEERCombinedSummaryStage2000200 %in% c(2,3,4,5) ~ 1,
                           SEERCombinedSummaryStage2000200 %in% c(7) ~ 2))

seer0to39$stage <- factor(seer0to39$stage, levels = c(0:2), labels = c("In situ/localized", "Regional", "Distant"))

table(seer0to39$stage, useNA ="always")

#
table(seer0to39$SEER_DateofDeath_Year, useNA = "always")

```

# Exclusions

# make separate dataframe for 0 to 19 
```{r}
seer0to19 <- seer0to39 %>%
  filter(age_cat %in% c("0-4", "5-9", "10-14", "15-19")) # 34,196

seer0to19$age_cat <- droplevels(seer0to19$age_cat)
```

# limit to first cancers
```{r}
seer0to19 <- seer0to19 %>%
  filter(Firstmalignantprimary_indicator == 1) # 32,396
```
# limit to sequence number 01 or 00
```{r}
seer0to19 <- seer0to19 %>%
  filter(Sequence_number %in% c("00", "01")) # 32,380
```

# Exclude those diagnosed before 2006 (a handful)
```{r}
seer0to19 <- seer0to19 %>%
  filter(Year_of_diagnosis >=2006) # 32,380
```

# Exclude multiple state Medicaid ids within a single year defined above
```{r}
seer0to19 <- seer0to19 %>% 
  filter(!patient_id %in% x) # 31,562
```

# Exclude those with Medicaid cancer registry match scores below  5
```{r}
match <- enroll_seerids %>%
  filter(MATCH_SCORE_HI <5)

# create list of unique ids to exclude
ids_2_exclude <- unique(match$patient_id)

seer0to19 <- seer0to19 %>%
  filter(!patient_id %in% ids_2_exclude) # 31,410
```

# Exclude those with missing data on variables used in the analysis
```{r}
seer0to19<-seer0to19 %>%
  filter(!is.na(fumo)) #31,410

seer0to19<-seer0to19 %>%
  filter(!is.na(can_dead)) # 31,247

seer0to19<-seer0to19 %>%
  filter(!is.na(In_Medicaid_Flag)) # 31,247

seer0to19<-seer0to19 %>%
  filter(!is.na(race_ethnicity)) # 31,247

seer0to19<-seer0to19 %>%
  filter(!is.na(age_cat)) # 31,247

seer0to19<-seer0to19 %>%
  filter(!is.na(Yost_ACS_2006_2010_quintile)) #30,654

seer0to19b <- seer0to19

missing_data_percent  <- 100- (30654/31410)*100
missing_data_percent  # <2.5%
```

# merge in mcont (medicaid enrollment data and recode 0 = Not linked, 1 = continuous, 2 = discontinuous, 3 = Other)
```{r}
seer0to19 <- left_join(seer0to19b, mcaid, by = "patient_id")

table(seer0to19$mcontb, useNA = "always")
table(mcaid$mcontb, useNA = "always")
table(seer0to19$In_Medicaid_Flag)

# recode a category for not in medicaid
seer0to19 <- seer0to19 %>%
  mutate(mcont66 = case_when(In_Medicaid_Flag == 0 ~ 0, # not linked
                             mcontb == 0 ~ 1, # continuous
                             mcontb == 1 ~ 2, # discontinuous
                             mcontb == 2 ~ 3)) # Other
table(seer0to19$mcont66, useNA = "always")

table(seer0to19$mcont66, seer0to19$In_Medicaid_Flag, useNA = "always")

# create factor variable for analyses
seer0to19$mcont66 <- factor(seer0to19$mcont66, levels = c(0:3), labels = c("Not enrolled", "Continuous", "Discontinuous", "Other"))
```

# check some variables
```{r}
table(seer0to19$can_type, seer0to19$SEERCombinedSummaryStage2000200, useNA = "always")

table(seer0to19$ICCCsiterecode_ICD_O_3_WHO_2008, seer0to19$can_type,  useNA= "always")
```

# FIGURE 1
```{r}
Figure1<-grViz("digraph flowchart {
      # node definitions with substituted label text
      node [fontname = Helvetica, shape = rectangle, fontsize=16]        
      tab1 [label = '@@1']
      tab2 [label = '@@2']
      tab3 [label = '@@3']
      tab4 [label = '@@4']
      tab5 [label = '@@5']
      tab6 [label = '@@6']
      tab7 [label = '@@7']
      # edge definitions with the node IDs
      tab1 -> tab2 -> tab3 -> tab4 -> tab5 -> tab6 -> tab7;
      }
      [1]: 'Records received for patients diagnosed with malignant cancers at ages 0 to 19 years from 2006-2013 n = 34,196'
      [2]: 'Excluding 1800 individuals whose cancer was not their first malignant primary was n = 32,396' 
      [3]: 'Excluding 16 individuals whose cancer sequence number was not in 0 or 1 n = 32,380'
      [4]: 'Excluding 818 individuals linked to multiple state Medicaid files n = 31,562'
      [5]: 'Excluding 152 individuals with Medicaid enrollment match scores <5 n = 31,410'
      [6]: 'Excluding 163 individuals with missing data on cancer specific cause of death n = 31,247'
      [7]: 'Excluding 593 individuals individuals with missing data on Yost ACS 2006 to 2010 Quintile n = 30,654'
      ")
Figure1
```

# TABLE 1
```{r}
# 0 to 19
label(seer0to19$age_cat)<-"Age category"
label(seer0to19$sex)<-"Sex"
label(seer0to19$race_ethnicity)<-"Race/ethnicity category"
label(seer0to19$fumo)<-"Follow-up months"
label(seer0to19$can_type)<-"Cancer type"
label(seer0to19$Yost_ACS_2006_2010_quintile) <- "Yost Index 2006 to 2010"
label(seer0to19$stage) <- "Stage"
label(seer0to19$mcont66) <- "Medicaid enrollment continuity"

# make can_dead a factor variable just for this table
seer0to19$can_dead2<-factor(seer0to19$can_dead, levels=c(0,1), labels= c("Alive", "Deceased"))
table1(~as.factor(seer0to19$SEERcausespecificdeathclassific)|mcont66, seer0to19)

label(seer0to19$can_dead2) <- "Cancer vital status"
seer0to19$In_Medicaid_Flag<-factor(seer0to19$In_Medicaid_Flag, levels=c(0,1), labels= c("Not Enrolled", "Enrolled"))
label(seer0to19$In_Medicaid_Flag) <- "Enrolled in Medicaid"

table(seer0to19$can_dead2)
table(seer0to19$In_Medicaid_Flag)

table1(~age_cat + sex  + race_ethnicity + Yost_ACS_2006_2010_quintile + In_Medicaid_Flag  + stage + fumo  + can_type + can_dead2|mcont66, droplevels = TRUE, seer0to19)

table1(~age_cat + sex  + race_ethnicity + Yost_ACS_2006_2010_quintile + In_Medicaid_Flag  + stage + fumo  + can_type + can_dead2|mcont66, droplevels = TRUE, seer0to19)

```

# TABLE 2: Stage analysis (unadjusted)
```{r}
# exclude leukemia cases from stage analysis and those with missing stage
seer0to19_stage <- seer0to19 %>% 
  filter(can_type != "I. Leukemias, myeloproliferative diseases, and myelodysplastic diseases") # this also excludes those with missing cancer type to yield 22,386

# the mlogit.cluster function does not like labels
seer0to19_stage <- seer0to19_stage %>% 
  remove_labels()

# exclude those with missing stage and redefine some variables so the function doesn't throw an error.
seer0to19_stage <- seer0to19_stage %>%
  filter(!is.na(stage)) %>% # this yields 21,502
  select(stage, In_Medicaid_Flag, SEER_registry, mcont66, race_cat2, age_cat, Yost_ACS_2006_2010_quintile, state, can_type) %>%
  mutate(stage2 = case_when(stage == "In situ/localized" ~ "0: In situ/localized",
                            stage == "Regional" ~ "1: Regional",
                            stage == "Distant" ~  "2: Distant"))

# function also likes it to only be a data.frame and nothing else
seer0to19_stage <- as.data.frame(seer0to19_stage)

# In_Medicaid_Flag unadjusted model
model1 <- multinom(stage2 ~ In_Medicaid_Flag, data=seer0to19_stage, trace = FALSE)
model1
summary(model1)

# extract coefficients with non-clustered SEs from the model and exponentiate for comparison 
exp(coef(model1))

# CI(95%)
exp(confint(model1, level=0.95))

# get coefficients, variance, clustered standard errors, and p values
b <- c(t(coef(model1)))
var <- mlogit.clust(model1, seer0to19_stage, "state") # get variance with function
se <- sqrt(diag(var))
p <- (1-pnorm(abs(b/se))) * 2

lowCI <- exp(b - 1.96*se)
highCI <-exp(b + 1.96*se)
df <- as.data.frame(exp(b))
df2<-as.data.frame(lowCI)
df3<-as.data.frame(highCI)
df4 <- cbind(df, df2, df3) # this is the result dataset

# relabel columns and dropt intercept rows
names(df4)[1:3]<-c("OR", "lowCI", "highCI")
df4

df4<-df4[-c(1,3),]
df4 <- tibble::rownames_to_column(df4, "Category") # Apply rownames_to_column

# Replace multiple strings at a time
rep_str = c('1: Regional:In_Medicaid_FlagEnrolled'='Regional','2: Distant:In_Medicaid_FlagEnrolled'='Distant')
rep_str2 = c('1: Regional:In_Medicaid_FlagEnrolled'='Enrolled','2: Distant:In_Medicaid_FlagEnrolled'='Enrolled')
df4$Stage<- str_replace_all(df4$Category, rep_str)
df4$Category<- str_replace_all(df4$Category, rep_str2) # this is the dataframe for summary results

# mcont66 model
model2 <- multinom(stage2 ~ mcont66, data=seer0to19_stage)
summary(model2)

# extract coefficients from the model and exponentiate
exp(coef(model2))

# CI(95%)
exp(confint(model2, level=0.95))

b <- c(t(coef(model2)))
var <- mlogit.clust(model2, seer0to19_stage, "state")
se <- sqrt(diag(var))

lowCI <- exp(b - 1.96*se)
highCI <-exp(b + 1.96*se)

df5 <- as.data.frame(exp(b))
df6<-as.data.frame(lowCI)
df7<-as.data.frame(highCI)
df8 <- cbind(df5, df6, df7) # this is the result dataset
# relabel columns and drop intercept rows

names(df8)[1:3]<-c("OR", "lowCI", "highCI")
df8

df8<-df8[-c(1,5),]
df8 <- tibble::rownames_to_column(df8, "Category") # Apply rownames_to_column
df8

# Replace multiple strings at a time
rep_str = c('1: Regional:mcont66Continuous'='Regional','1: Regional:mcont66Discontinuous'='Regional','1: Regional:mcont66Other'='Regional', '2: Distant:mcont66Continuous'='Distant','2: Distant:mcont66Discontinuous'='Distant','2: Distant:mcont66Other'='Distant')
df8$Stage<- str_replace_all(df8$Category, rep_str)
df8
rep_str2 = c('1: Regional:mcont66Continuous'='Continuous','1: Regional:mcont66Discontinuous'='Discontinuous','1: Regional:mcont66Other'='Other', '2: Distant:mcont66Continuous'='Continuous','2: Distant:mcont66Discontinuous'='Discontinuous','2: Distant:mcont66Other'='Other')
df8$Category<- str_replace_all(df8$Category, rep_str2)
df8 # this is the dataframe for summary results
```

# Adjusted models
```{r}
model1a <- multinom(stage2 ~ In_Medicaid_Flag + race_cat2 + age_cat + Yost_ACS_2006_2010_quintile , data=seer0to19_stage)
summary(model1a)
nrow(fitted(model1a)) # 21,507 is number of obs

b <- c(t(coef(model1a)))
var <- mlogit.clust(model1a, seer0to19_stage, "state")
se <- sqrt(diag(var))

lowCI <- exp(b - 1.96*se)
highCI <-exp(b + 1.96*se)
df9 <- as.data.frame(exp(b))
df10<-as.data.frame(lowCI)
df11<-as.data.frame(highCI)
df12 <- cbind(df9, df10, df11) # this is the result dataset

# relabel columns and drop intercept rows
names(df12)[1:3]<-c("OR", "cluster.lowCI", "cluster.highCI")
df12

df12<-df12[-c(1,3:11, 13:20),]
df12 <- tibble::rownames_to_column(df12, "Category") # Apply rownames_to_column
df12

# Replace multiple strings at a time
rep_str = c('1: Regional:In_Medicaid_FlagEnrolled'='Regional', '2: Distant:In_Medicaid_FlagEnrolled'='Distant')
df12$Stage<- str_replace_all(df12$Category, rep_str)

df12$Category<- "Enrolled"
df12 # this is the dataframe for summary results

# adjusted model
model2a <- multinom(stage2 ~mcont66 + race_cat2 + age_cat + Yost_ACS_2006_2010_quintile, data=seer0to19_stage)
summary(model2a)
nrow(fitted(model2a))

b <- c(t(coef(model2a)))
var <- mlogit.clust(model2a, seer0to19_stage, "state")
se <- sqrt(diag(var))

lowCI <- exp(b - 1.96*se)
highCI <-exp(b + 1.96*se)
df13 <- as.data.frame(exp(b))
df14<-as.data.frame(lowCI)
df15<-as.data.frame(highCI)
df16 <- cbind(df13, df14, df15) # this is the result dataset

# relabel columns and drop intercept rows
names(df16)[1:3]<-c("OR", "cluster.lowCI", "cluster.highCI")
df16

df16<-df16[-c(1,5:13, 17:24),]
df16 <- tibble::rownames_to_column(df16, "Category") # Apply rownames_to_column
df16

# Replace multiple strings at a time
rep_str = c('1: Regional:mcont66Continuous'='Regional','1: Regional:mcont66Discontinuous'='Regional','1: Regional:mcont66Other'='Regional', '2: Distant:mcont66Continuous'='Distant','2: Distant:mcont66Discontinuous'='Distant','2: Distant:mcont66Other'='Distant')
df16$Stage<- str_replace_all(df16$Category, rep_str)

rep_str2 = c('1: Regional:mcont66Continuous'='Continuous','1: Regional:mcont66Discontinuous'='Discontinuous','1: Regional:mcont66Other'='Other', '2: Distant:mcont66Continuous'='Continuous','2: Distant:mcont66Discontinuous'='Discontinuous','2: Distant:mcont66Other'='Other')
df16$Category<- str_replace_all(df16$Category, rep_str2)
df16 # this is the dataframe for summary results
```
# Table 2 with clustered standard errors used for confidence interval calculation
```{r}
Table2 <- rbind(df12, df16)
```
# Relevel to determine if there is a difference between continuous and discontinuous
```{r}
seer0to19_stage <- seer0to19_stage%>%
  mutate(mcont66.2 = case_when(mcont66 == "Continuous" ~ 0,
                               mcont66 == "Discontinuous" ~ 1,
                               mcont66 == "Other" ~ 2,
                               mcont66 == "Not enrolled" ~ 3))
seer0to19_stage$mcont66.2 <- factor(seer0to19_stage$mcont66.2, levels = c(0:3), labels = c("Continuous", "Discontinuous", "Other", "Not enrolled"))

# adjusted model
model2a <- multinom(stage2 ~ mcont66.2 + race_cat2 + age_cat + Yost_ACS_2006_2010_quintile, data=seer0to19_stage)
summary(model2a)
nrow(fitted(model2a))

b <- c(t(coef(model2a)))
var <- mlogit.clust(model2a, seer0to19_stage, "state")
se <- sqrt(diag(var))

lowCI <- exp(b - 1.96*se)
highCI <-exp(b + 1.96*se)
df17 <- as.data.frame(exp(b))
df18<-as.data.frame(lowCI)
df19<-as.data.frame(highCI)
df20 <- cbind(df17, df18, df19) # this is the result dataset

# relabel columns and drop intercept rows
names(df20)[1:3]<-c("OR", "cluster.lowCI", "cluster.highCI")
df20

df20<-df20[-c(1,5:13, 17:24),]
df20 <- tibble::rownames_to_column(df20, "Category") # Apply rownames_to_column
df20  
```

# SUPPLEMENTARY TABLE 1a by cancer type with clustered standard errors for confidence interval calculation (In_Medicaid_Flag)
```{r}
limited <- seer0to19_stage %>%
 filter(can_type %in% c("I. Leukemias, myeloproliferative diseases, and myelodysplastic diseases","II. Lymphomas and reticuloendothelial neoplasms", "III. CNS and miscellaneous intracranial and intraspinal neoplasms", "IV. Neuroblastoma and other peripheral nervous cell tumors", "V. Retinoblastoma", "VI. Renal tumors", "VII. Hepatic tumors", "VIII. Malignant bone tumors", "IX. Soft tissue and other extraosseous sarcomas", "X. Germ cell tumors, trophoblastic tumors, and neoplasms of gonads", "XI. Other malignant epithelial neoplasms and malignant melanomas"))

ct <- unique(limited$can_type)

# In_Medicaid_Flag
df1a <- data.frame(matrix(ncol=6, nrow=0)) # create the table
columns<-c("Category", "OR","cluster.lowCI", "cluster.highCI", "CancerType") # label the columns with names
colnames(df1a)<-columns

for (i in ct) {
  dfa<- limited %>% 
    filter(can_type == i)
  x<-multinom(stage2 ~ In_Medicaid_Flag + race_cat2 + age_cat + Yost_ACS_2006_2010_quintile, data=dfa)

b <- c(t(coef(x)))
var <- mlogit.clust(x, dfa, "state")
se <- sqrt(diag(var))

lowCI <- exp(b - 1.96*se)
highCI <-exp(b + 1.96*se)
b <- as.data.frame(exp(b))
lCI<-as.data.frame(lowCI)
hCI<-as.data.frame(highCI)
df2a <- cbind(b, lCI, hCI) # this is the result dataset
# relabel columns and dropt intercept rows

names(df2a)[1:3]<-c("OR", "cluster.lowCI", "cluster.highCI")
df2a<-df2a[-c(1,3:11, 13:20),]
df2a <- tibble::rownames_to_column(df2a, "Category") # Apply rownames_to_column
df2a  
df2a$CancerType<-i # add cancer type
df1a <-rbind(df2a, df1a)
}

SuppTable1a<- df1a[order(df1a$CancerType),] # In_Medicaid_Flag

# Mcont66
df3a <- data.frame(matrix(ncol=6, nrow=0)) #create the table
columns<-c("Category", "OR","cluster.lowCI", "cluster.highCI", "CancerType") #label the columns with names
colnames(df3a)<-columns

for (i in ct) {
  dfa<- limited %>% 
    filter(can_type == i)
  x<-multinom(stage2 ~ mcont66 + race_cat2 + age_cat + Yost_ACS_2006_2010_quintile, data=dfa)

b <- c(t(coef(x)))
var <- mlogit.clust(x, dfa, "state")
se <- sqrt(diag(var))

lowCI <- exp(b - 1.96*se)
highCI <-exp(b + 1.96*se)
b <- as.data.frame(exp(b))
lCI<-as.data.frame(lowCI)
hCI<-as.data.frame(highCI)
df4a <- cbind(b, lCI, hCI) # this is the result dataset
# relabel columns and dropt intercept rows

names(df4a)[1:3]<-c("OR", "cluster.lowCI", "cluster.highCI")
df4a <-df4a[-c(1,5:13, 17:24),]
df4a <- tibble::rownames_to_column(df4a, "Category") # Apply rownames_to_column
df4a  
df4a$CancerType<-i # add cancer type
df3a <-rbind(df4a, df3a)
}

SuppTable1b<- df3a[order(df3a$CancerType),] # mcont66
```

# SUPPLEMENTAEY TABLE 1:
```{r}
SuppTable1 <- rbind(SuppTable1a, SuppTable1b)

# Replace multiple strings at a time
rep_str = c('1: Regional:mcont66Continuous'='Regional','1: Regional:mcont66Discontinuous'='Regional','1: Regional:mcont66Other'='Regional', '2: Distant:mcont66Continuous'='Distant','2: Distant:mcont66Discontinuous'='Distant','2: Distant:mcont66Other'='Distant', '1: Regional:In_Medicaid_FlagEnrolled'='Regional','2: Distant:In_Medicaid_FlagEnrolled'='Distant')
SuppTable1$Stage<- str_replace_all(SuppTable1$Category, rep_str)

rep_str2 = c('1: Regional:mcont66Continuous'='Continuous','1: Regional:mcont66Discontinuous'='Discontinuous','1: Regional:mcont66Other'='Other', '2: Distant:mcont66Continuous'='Continuous','2: Distant:mcont66Discontinuous'='Discontinuous','2: Distant:mcont66Other'='Other', '1: Regional:In_Medicaid_FlagEnrolled'='Enrolled','2: Distant:In_Medicaid_FlagEnrolled'='Enrolled')
SuppTable1$Category<- str_replace_all(SuppTable1$Category, rep_str2)

SuppTable1
```

# Add table 2 to SuppTable1 (with clustered SEs) to plot on one plot
```{r}
Table2$CancerType<-"Overall"

SuppTable1 <- rbind(Table2, SuppTable1)
```

# FIGURE 2: Stage analysis by cancer type
```{r}
# graph ORs and 95% CIs
SuppTable1a <- SuppTable1 %>%
  filter(CancerType != "V. Retinoblastoma")

SuppTable1a <- SuppTable1a %>%
  mutate(Abbrev = case_when(CancerType == "Overall" ~ "Overall",
                            CancerType =="II. Lymphomas and reticuloendothelial neoplasms" ~ "LYMPH",
                            CancerType == "III. CNS and miscellaneous intracranial and intraspinal neoplasms" ~ "CNS",
                            CancerType == "IV. Neuroblastoma and other peripheral nervous cell tumors" ~ "NB",
                            CancerType == "V. Retinoblastoma" ~ "RB",
                            CancerType == "VI. Renal tumors" ~ "RT",
                            CancerType == "VII. Hepatic tumors" ~ "HT",
                            CancerType == "VIII. Malignant bone tumors" ~ "MBT",
                            CancerType == "IX. Soft tissue and other extraosseous sarcomas" ~ "STS", 
                            CancerType == "X. Germ cell tumors, trophoblastic tumors, and neoplasms of gonads" ~ "GCT",
                            CancerType == "XI. Other malignant epithelial neoplasms and malignant melanomas" ~ "OTHEP",
                            CancerType == "XII. Other and unspecified malignant neoplasms" ~ "OTHUN"))

SuppTable1a$Abbrev <- factor(SuppTable1a$Abbrev, levels = c("Overall", "LYMPH", "CNS","NB", "RB",
                                                  "RT", "HT", "MBT", "STS", "GCT",
                                                  "OTHEP", "OTHUN"))

SuppTable1a$Stage <- factor(SuppTable1a$Stage, levels = c("Regional", "Distant"))

SuppTable1a$Abbrev<- fct_rev(SuppTable1a$Abbrev)

Plot1 <- SuppTable1a %>%
  filter(Category == "Enrolled", Stage == "Regional") %>%
  ggplot() +
  geom_point(aes(y=OR, x = Abbrev, group = Category, color = Category)) + 
  geom_hline(yintercept =1, linetype=2) +
  theme(axis.text.y=element_blank()) +
  geom_errorbar(aes(y=OR, x=Abbrev, group = Category, color = Category, ymax = cluster.highCI, ymin = cluster.lowCI), width = 0.1) +
  theme_bw() +
  theme(axis.line = element_line(colour = "black"),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_blank(),
    panel.background = element_blank(),
    axis.text.x=element_text(size=11, face = "bold", color = "black"),
    axis.text.y=element_text(size=11, face = "bold", color = "black"),
    axis.title.x=element_text(size=12, face = "bold", color = "black"),
    axis.title.y=element_text(size=12, face = "bold", color = "black"))+
  coord_flip() +
  labs(color = NULL, title = "A", x = "Cancer type") +
  ylim(0,2) +
  theme(strip.placement = "inside", legend.position = "bottom",  strip.text = element_text(size = 9, face="bold", color="black"), strip.background = element_rect(fill="lightgray"), plot.title = element_text(face="bold"), legend.text=element_text(size=9, face="bold")) +
  scale_color_manual(breaks = c("Enrolled"), values=c("black"))

Plot1

Plot2 <- SuppTable1a %>%
  filter(Category == "Enrolled", Stage == "Distant") %>%
  ggplot() +
  geom_point(aes(y=OR, x = Abbrev, group = Category, color = Category)) + 
  geom_hline(yintercept =1, linetype=2) +
  theme(axis.text.y=element_blank()) +
  geom_errorbar(aes(y=OR, x=Abbrev, group = Category, color = Category, ymax = cluster.highCI, ymin = cluster.lowCI), width = 0.1) +
  theme_bw() +
  theme(axis.line = element_line(colour = "black"),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_blank(),
    panel.background = element_blank(),
    axis.text.x=element_text(size=11, face = "bold", color = "black"),
    axis.text.y=element_text(size=11, face = "bold", color = "black"),
    axis.title.x=element_text(size=12, face = "bold", color = "black"),
    axis.title.y=element_text(size=12, face = "bold", color = "black"))+
  coord_flip() +
  labs(color = NULL, title = "B", x = "Cancer type") +
  ylim(0,3.5) +
  theme(strip.placement = "inside", legend.position = "bottom",  strip.text = element_text(size = 9, face="bold", color="black"), strip.background = element_rect(fill="lightgray"), plot.title = element_text(face="bold"), legend.text=element_text(size=9, face="bold")) +
  scale_color_manual(breaks = c("Enrolled"), values=c("black"))

Plot2

```

# FIGURE 2:  Stage analysis by cancer type mcont66
```{r}
# graph ORs and 95% CIs
SuppTable1a$Category <- factor(SuppTable1a$Category, levels = c("Enrolled", "Discontinuous", "Other", "Continuous"))

Plot3 <- SuppTable1a %>%
  filter(Category != "Enrolled", Stage == "Regional") %>%
  ggplot() +
  geom_point(aes(y=OR, x = Abbrev, group = Category, color = Category), na.rm=TRUE, position=position_dodge(width=0.9)) + 
  geom_hline(yintercept =1, linetype=2) +
  geom_vline(xintercept =c(1.5, 2.5,3.5,4.5,5.5,6.5,7.5,8.5,9.5), linetype=2, color = "lightgray") +
  theme(axis.text.y=element_blank()) +
  geom_errorbar(position = position_dodge(width=0.9), aes(y=OR, x=Abbrev, group = Category, color = Category,  ymax = cluster.highCI, ymin = cluster.lowCI),  width = 0.1) +
  theme_bw() +
  theme(axis.line = element_line(colour = "black"),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_blank(),
    panel.background = element_blank(),
    axis.text.x=element_text(size=11, face = "bold", color = "black"),
    axis.text.y=element_text(size=11, face = "bold", color = "black"),
    axis.title.x=element_text(size=12, face = "bold", color = "black"),
    axis.title.y=element_text(size=12, face = "bold", color = "black"))+
  coord_flip() +
 # facet_grid(cols = vars(y.level)) +
  labs(color = NULL, title = "C", x = "Cancer type") +
  ylim(0,4.0) +
  theme(strip.placement = "inside", legend.position = "bottom",  strip.text = element_text(size = 9, face="bold", color="black"), strip.background = element_rect(fill="lightgray"), plot.title = element_text(face="bold"), legend.text=element_text(size=9, face="bold")) +
  scale_color_manual(breaks = c("Discontinuous", "Continuous", "Other"),
                        values=c("red", "cyan3", "blue"))

Plot3

Plot4 <- SuppTable1a %>%
  filter(Category != "Enrolled", Stage == "Distant") %>%
  ggplot() +
  geom_point(aes(y=OR, x = Abbrev, group = Category, color = Category), na.rm=TRUE, position=position_dodge(width=0.9)) + 
  geom_hline(yintercept =1, linetype=2) +
  geom_vline(xintercept =c(1.5, 2.5,3.5,4.5,5.5,6.5,7.5,8.5,9.5), linetype=2, color = "lightgray") +
  theme(axis.text.y=element_blank()) +
  geom_errorbar(position = position_dodge(width=0.9), aes(y=OR, x=Abbrev, group = Category, color = Category,  max = cluster.highCI, ymin = cluster.lowCI),  width = 0.1) +
  theme_bw() +
  theme(axis.line = element_line(colour = "black"),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_blank(),
    panel.background = element_blank(),
    axis.text.x=element_text(size=11, face = "bold", color = "black"),
    axis.text.y=element_text(size=11, face = "bold", color = "black"),
    axis.title.x=element_text(size=12, face = "bold", color = "black"),
    axis.title.y=element_text(size=12, face = "bold", color = "black"))+
  coord_flip() +
 # facet_grid(cols = vars(y.level)) +
  labs(color = NULL, title = "D", x = "Cancer type") +
  ylim(0,5.5) +
  theme(strip.placement = "inside", legend.position = "bottom",  strip.text = element_text(size = 9, face="bold", color="black"), strip.background = element_rect(fill="lightgray"), plot.title = element_text(face="bold"), legend.text=element_text(size=9, face="bold")) +
  scale_color_manual(breaks = c("Discontinuous", "Continuous", "Other"),
                        values=c("red", "cyan3", "blue"))

Plot4
```

```{r}
pdf("Figure 2.pdf", width=8.5, height=11, paper="USr")
Plot1 + Plot2 + Plot3 + Plot4 + plot_layout(ncol=2, nrow=2, guides='collect') &
  theme(legend.position='bottom')

dev.off()
```

# FIGURE 3 survival plots (# In_Medicaid_flag and mcont66) 
```{r}
# In_Medicaid_Flag
plot1 <- ggsurvplot(fit = survfit(Surv(fumo, can_dead) ~ In_Medicaid_Flag, data = seer0to19), 
    xlab = "Months from diagnosis", 
    ylab = "Survival probability", 
    fun="pct",  
    title = "A",
    font.title = c(12, "bold"),
    legend.title = "",
    legend = "bottom",
    font.legend=c(12, "bold"),
    legend.labs = c("Not enrolled", "Enrolled"),
    palette = c("red", "cyan3", "blue", "black"),
    font.x = c(12, "bold"),
    font.y = c(12, "bold"),
    xlim=c(0, 60),
    break.x.by = 12,
    ylim=c(50, 100),
    censor.size=0.5,
    font.tickslab = c(12, "bold"),
    pval=TRUE,
    pval.size = 5,
    pval.coord = c(2,55),
    conf.int = TRUE)

plot1

# mcont66
 plot2 <- ggsurvplot(fit = survfit(Surv(fumo, can_dead) ~ mcont66, data = seer0to19), 
    xlab = "Months from diagnosis", 
    ylab = "Survival probability", 
    fun="pct", 
    title = "B",
    font.title = c(12, "bold"),
    legend.title = "",
    legend = "bottom",
    font.legend=c(12, "bold"),
    legend.labs = c("Not enrolled", "Continuous", "Discontinuous", "Other"),
    palette = c("red", "cyan3", "blue", "black", "gray"),
    font.x = c(12, "bold"),
    font.y = c(12, "bold"),
    xlim=c(0, 60),
    break.x.by = 12,
    ylim=c(50, 100),
    censor.size=0.5,
    font.tickslab = c(12, "bold"),
    pval=TRUE,
    pval.size = 5,
    pval.coord = c(2,55),
    conf.int = TRUE)

plot2
pdf("Figure 3.pdf", width=8.5, height=11, paper="USr")
plot1$plot + plot2$plot + plot_layout(ncol=1, nrow=2, guides='keep') &
  theme(legend.position='bottom')

dev.off()
```
# SUPPLEMENTARY TABLE 2: Estimate survival time 0 to 19
```{r}
# In_Medicaid_Flag
SuppTable2a <- summary(survfit(Surv(fumo, can_dead) ~ In_Medicaid_Flag, data = seer0to19), times = 60) # estimates 60 month survival time
SuppTable2a <- as.data.frame(cbind(SuppTable2a$strata, round(SuppTable2a$surv*100, 2), round(SuppTable2a$lower*100, 2), round(SuppTable2a$upper*100,2))) # makes dataframe

SuppTable2a <- setNames(SuppTable2a, c("Category", "Survival Percentage", "Lower CI", "Upper CI")) # gives column names
SuppTable2a <- SuppTable2a %>%
  mutate(Category = case_when(Category ==1 ~ "Not enrolled",
                              Category ==2 ~ "Enrolled"))

#mcont66
SuppTable2b <- summary(survfit(Surv(fumo, can_dead) ~ mcont66, data = seer0to19), times = 60)
SuppTable2b <- as.data.frame(cbind(SuppTable2b$strata, round(SuppTable2b$surv*100, 2), round(SuppTable2b$lower*100, 2), round(SuppTable2b$upper*100,2)))

SuppTable2b <- setNames(SuppTable2b, c("Category", "Survival Percentage", "Lower CI", "Upper CI"))
SuppTable2b <- SuppTable2b %>%
  mutate(Category = case_when(Category ==1 ~ "Not enrolled",
                              Category ==2 ~ "Continuous",
                              Category == 3 ~ "Discontinuous",
                              Category == 4 ~ "Other"))

# Make Supplementary table 2
SuppTable2 <- rbind(SuppTable2a, SuppTable2b)
SuppTable2 <- unique(SuppTable2)
```

# SUPPLEMENTARY TABLE 3: Populate a table with survival by age group
```{r}
# In_Medicaid_Flag
# calculate survival at 60 mo
results <- data.frame(matrix(ncol=5, nrow=0)) #create the table
columns<-c("Medicaid_Category", "Survival Percentage",  "LowerCI", "UpperCI", "Age_cat") #label the columns with names
colnames(results)<-columns

for (i in age) {
  df<- seer0to19 %>% 
    filter(age_cat == i)
  x<-summary(survfit(Surv(fumo, can_dead) ~ In_Medicaid_Flag, df), times=60)
  results2 <- data.frame(matrix(ncol=4, nrow=0)) # create the table
  columns<-c("Medicaid_Category", "Survival Percentage",  "LowerCI", "UpperCI") #label the columns with names
  colnames(results2)<-columns
  results2[1:2,]<-c(x$strata, round(x$surv,3)*100, round(x$lower,3)*100 , round(x$upper,3)*100)
  results2$Group<- i
  results <-rbind(results, results2)
}

# sort by cancertype and add labels to Medicaid categories
SuppTable3a<- results[order(results$Group),]
SuppTable3a<- SuppTable3a %>%
  rename(Category = Medicaid_Category) %>%
  mutate(Category = case_when(Category == 1 ~ "Not enrolled",
                              Category == 2 ~ "Enrolled"))

# mcont66
# calculate survival at 60 mo
results <- data.frame(matrix(ncol=5, nrow=0)) #create the table
columns<-c("Medicaid_Category", "Survival Percentage",  "LowerCI", "UpperCI", "Age_cat") #label the columns with names
colnames(results)<-columns

for (i in age) {
  df<- seer0to19 %>% 
    filter(age_cat == i)
  x<-summary(survfit(Surv(fumo, can_dead) ~ mcont66, df), times=60)
  results2 <- data.frame(matrix(ncol=4, nrow=0)) # create the table
  columns<-c("Medicaid_Category", "Survival Percentage",  "LowerCI", "UpperCI") #label the columns with names
  colnames(results2)<-columns
  results2[1:4,]<-c(x$strata, round(x$surv,3)*100, round(x$lower,3)*100 , round(x$upper,3)*100)
  results2$Group<- i
  results <-rbind(results, results2)
}

# sort by group and add labels to Medicaid categories
SuppTable3b <- results[order(results$Group),]
SuppTable3b <- SuppTable3b%>%
  rename(Category = Medicaid_Category) %>%
  mutate(Category = case_when(Category == 1 ~ "Not enrolled",
                              Category == 2 ~ "Continuous",
                              Category == 3 ~ "Discontinuous",
                              Category == 4 ~ "Other"))

SuppTable3a$Group <- factor(SuppTable3a$Group, levels = c("0-4", "5-9", "10-14", "15-19"))
SuppTable3a <- SuppTable3a[order(SuppTable3a$Group),]

SuppTable3b$Group <- factor(SuppTable3b$Group, levels = c("0-4", "5-9", "10-14", "15-19"))
SuppTable3b <- SuppTable3b[order(SuppTable3b$Group),]

SuppTable3 <-rbind(SuppTable3a, SuppTable3b)
SuppTable3 <- unique(SuppTable3) # remove non-unique rows
```

# SUPPLEMENTARY TABLE 4: make a function to populate a table with survival by race
```{r}
race <- unique(seer0to19$race_ethnicity)
limited <- seer0to19 %>%
 filter(race_ethnicity %in% c("Non-Hispanic White","Non-Hispanic Black",  "Non-Hispanic API", "Hispanic", "Non-Hispanic Other"))

# In_Medicaid_Flag
# calculate survival at 60 mo
results <- data.frame(matrix(ncol=5, nrow=0)) # create the table
columns<-c("Medicaid_Category", "Survival Percentage",  "LowerCI", "UpperCI", "Race_ethnicity") #label the columns with names
colnames(results)<-columns

for (i in race) {
  df<- seer0to19 %>% 
    filter(race_ethnicity == i)
  x<-summary(survfit(Surv(fumo, can_dead) ~ In_Medicaid_Flag, df), times=60)
  results2 <- data.frame(matrix(ncol=4, nrow=0)) # create the table
  columns<-c("Medicaid_Category", "Survival Percentage",  "LowerCI", "UpperCI") # label the columns with names
  colnames(results2)<-columns
  results2[1:2,]<-c(x$strata, round(x$surv,3)*100, round(x$lower,3)*100 , round(x$upper,3)*100)
  results2$Group<- i
  results <-rbind(results, results2)
}

# sort by cancertype and add labels to Medicaid categories
SuppTable4a <- results[order(results$Group),]
SuppTable4a <- SuppTable4a %>%
  rename(Category = Medicaid_Category) %>%
  mutate(Category = case_when(Category == 1 ~ "Not enrolled",
                              Category == 2 ~ "Enrolled"))

# mcont66
# calculate survival at 60 mo
results <- data.frame(matrix(ncol=5, nrow=0)) #create the table
columns<-c("Medicaid_Category", "Survival Percentage",  "LowerCI", "UpperCI", "Race_ethnicity") #label the columns with names
colnames(results)<-columns

for (i in race) {
  df<- limited %>% 
    filter(race_ethnicity == i)
  x<-summary(survfit(Surv(fumo, can_dead) ~ mcont66, df), times=60)
  results2 <- data.frame(matrix(ncol=4, nrow=0)) # create the table
  columns<-c("Medicaid_Category", "Survival Percentage",  "LowerCI", "UpperCI") # label the columns with names
  colnames(results2)<-columns
  results2[1:4,]<-c(x$strata, round(x$surv,3)*100, round(x$lower,3)*100 , round(x$upper,3)*100)
  results2$Group<- i
  results <-rbind(results, results2)
}

# sort by group and add labels to Medicaid categories
SuppTable4b <- results[order(results$Group),]
SuppTable4b <- SuppTable4b %>%
  rename(Category = Medicaid_Category) %>%
  mutate(Category = case_when(Category == 1 ~ "Not enrolled",
                              Category == 2 ~ "Continuous",
                              Category == 3 ~ "Discontinuous",
                              Category == 4 ~ "Other"))

SuppTable4a$Group <- factor(SuppTable4a$Group, levels = c("Non-Hispanic White", "Non-Hispanic Black", "Non-Hispanic API", "Non-Hispanic Other", "Hispanic"))
SuppTable4a <- SuppTable4a[order(SuppTable4a$Group),]

SuppTable4b$Group <- factor(SuppTable4b$Group, levels = c("Non-Hispanic White", "Non-Hispanic Black", "Non-Hispanic API", "Non-Hispanic Other", "Hispanic"))
SuppTable4b <- SuppTable4b[order(SuppTable4b$Group),]

SuppTable4 <- rbind(SuppTable4a, SuppTable4b)
SuppTable4 <- unique(SuppTable4)
```

# SUPPLEMENTARY TABLE 5: make a function to populate a table with survival by tumor type
```{r}
limited <- seer0to19 %>%
 filter(can_type %in% c("I. Leukemias, myeloproliferative diseases, and myelodysplastic diseases","II. Lymphomas and reticuloendothelial neoplasms", "III. CNS and miscellaneous intracranial and intraspinal neoplasms", "IV. Neuroblastoma and other peripheral nervous cell tumors", "V. Retinoblastoma", "VI. Renal tumors", "VII. Hepatic tumors", "VIII. Malignant bone tumors", "IX. Soft tissue and other extraosseous sarcomas", "X. Germ cell tumors, trophoblastic tumors, and neoplasms of gonads", "XI. Other malignant epithelial neoplasms and malignant melanomas"))
ct<-unique(limited$can_type)

# In_Medicaid_Flag
# calculate survival at 60 mo
results <- data.frame(matrix(ncol=5, nrow=0)) # create the table
columns<-c("Medicaid_Category", "Survival Percentage",  "LowerCI", "UpperCI", "Cancer_type") #label the columns with names
colnames(results)<-columns

for (i in ct) {
  df<- limited %>% 
    filter(can_type == i)
  x<-summary(survfit(Surv(fumo, can_dead) ~ In_Medicaid_Flag, df), times=60)
  results2 <- data.frame(matrix(ncol=4, nrow=0)) # create the table
  columns<-c("Medicaid_Category", "Survival Percentage",  "LowerCI", "UpperCI") # label the columns with names
  colnames(results2)<-columns
  results2[1:2,]<-c(x$strata, round(x$surv,3)*100, round(x$lower,3)*100 , round(x$upper,3)*100)
  results2$Group<- i
  results <-rbind(results, results2)
}

# sort by cancertype and add labels to Medicaid categories
SuppTable5a <- results[order(results$Group),]
SuppTable5a <- SuppTable5a %>%
  rename(Category = Medicaid_Category) %>%
  mutate(Category = case_when(Category == 1 ~ "Not enrolled",
                              Category == 2 ~ "Enrolled"))

# mcont66
# calculate survival at 60 mo
results <- data.frame(matrix(ncol=5, nrow=0)) # create the table
columns<-c("Medicaid_Category", "Survival Percentage",  "LowerCI", "UpperCI", "Cancer_type") # label the columns with names
colnames(results)<-columns

for (i in ct) {
  df<- limited %>% 
    filter(can_type == i)
  x<-summary(survfit(Surv(fumo, can_dead) ~ mcont66, df), times=60)
  results2 <- data.frame(matrix(ncol=4, nrow=0)) # create the table
  columns<-c("Medicaid_Category", "Survival Percentage",  "LowerCI", "UpperCI") #label the columns with names
  colnames(results2)<-columns
  results2[1:4,]<-c(x$strata, round(x$surv,3)*100, round(x$lower,3)*100 , round(x$upper,3)*100)
  results2$Group<- i
  results <-rbind(results, results2)
}
# sort by group and add labels to Medicaid categories
SuppTable5b<- results[order(results$Group),]
SuppTable5b<- SuppTable5b %>%
  rename(Category = Medicaid_Category) %>%
  mutate(Category = case_when(Category == 1 ~ "Not enrolled",
                                       Category == 2 ~ "Continuous",
                                       Category == 3 ~ "Discontinuous",
                                       Category == 4 ~ "Other"))

SuppTable5a$Group <- factor(SuppTable5a$Group, levels = c("I. Leukemias, myeloproliferative diseases, and myelodysplastic diseases","II. Lymphomas and reticuloendothelial neoplasms", "III. CNS and miscellaneous intracranial and intraspinal neoplasms", "IV. Neuroblastoma and other peripheral nervous cell tumors", "V. Retinoblastoma", "VI. Renal tumors", "VII. Hepatic tumors", "VIII. Malignant bone tumors", "IX. Soft tissue and other extraosseous sarcomas", "X. Germ cell tumors, trophoblastic tumors, and neoplasms of gonads", "XI. Other malignant epithelial neoplasms and malignant melanomas"))

SuppTable5a <- SuppTable5a[order(SuppTable5a$Group),]

SuppTable5b$Group <- factor(SuppTable5b$Group, levels = c("I. Leukemias, myeloproliferative diseases, and myelodysplastic diseases","II. Lymphomas and reticuloendothelial neoplasms", "III. CNS and miscellaneous intracranial and intraspinal neoplasms", "IV. Neuroblastoma and other peripheral nervous cell tumors", "V. Retinoblastoma", "VI. Renal tumors", "VII. Hepatic tumors", "VIII. Malignant bone tumors", "IX. Soft tissue and other extraosseous sarcomas", "X. Germ cell tumors, trophoblastic tumors, and neoplasms of gonads", "XI. Other malignant epithelial neoplasms and malignant melanomas"))
SuppTable5b <- SuppTable5b[order(SuppTable5b$Group),]

SuppTable5 <- rbind(SuppTable5a, SuppTable5b)
SuppTable5 <- unique(SuppTable5)
```

# TABLE 3: Cox proportional hazards models (0 to 19) 
```{r}
# unadjusted model 0 TO 19
res.cox1 <- coxph(Surv(fumo, can_dead) ~ In_Medicaid_Flag + cluster(state), data = seer0to19) # Univariable model
res.cox1
summary(res.cox1)

# adjusted models 0 to 19
res.cox2 <- coxph(Surv(fumo, can_dead) ~ In_Medicaid_Flag + race_ethnicity + age_cat + Yost_ACS_2006_2010_quintile + cluster(state), data = seer0to19) # Model adjusted for race age and a measure of poverty
res.cox2
summary(res.cox2)

# adjust for stage
res.cox2a <- coxph(Surv(fumo, can_dead) ~ In_Medicaid_Flag + race_ethnicity + age_cat + Yost_ACS_2006_2010_quintile + stage + cluster(state), data = seer0to19) # Model adjusted for race age and a measure of poverty and stagee
res.cox2a
summary(res.cox2a)

# mcont66 unadjusted model 0 TO 19
res.cox3 <- coxph(Surv(fumo, can_dead) ~ mcont66 + cluster(state), data = seer0to19) # Univariable model
res.cox3
summary(res.cox3)

# mcont66 adjusted models 0 to 19
res.cox4 <- coxph(Surv(fumo, can_dead) ~ mcont66 + race_ethnicity +age_cat + Yost_ACS_2006_2010_quintile + cluster(state), data = seer0to19) # Model adjusted for race age and a measure of poverty
res.cox4
summary(res.cox4)

# adjust for stage
res.cox4b <- coxph(Surv(fumo, can_dead) ~ mcont66 + race_ethnicity +age_cat + Yost_ACS_2006_2010_quintile + stage + cluster(state), data = seer0to19) # Model adjusted for race age and a measure of poverty
res.cox4b
summary(res.cox4b)

a<-as.data.frame(exp(cbind(HR = coef(res.cox1), confint(res.cox1))))
   a<-a[c(1),] %>%
    rownames_to_column() %>%
    mutate(model="1. Unadjusted") %>%
    mutate(age_group = "0 to 19 years")

b<-as.data.frame(exp(cbind(HR = coef(res.cox2), confint(res.cox2))))
   b<-b[c(1),] %>%
    rownames_to_column() %>%
    mutate(model="2. Adjusted for race/ethnicity, age, Yost SES indicator") %>%
    mutate(age_group = "0 to 19 years")

c<-as.data.frame(exp(cbind(HR = coef(res.cox3), confint(res.cox3))))
   c<-c[c(1,2,3),] %>%
    rownames_to_column() %>%
    mutate(model="3. Unadjusted") %>%
    mutate(age_group = "0 to 19 years")

d<-as.data.frame(exp(cbind(HR = coef(res.cox4), confint(res.cox4))))
   d<-d[c(1,2,3),] %>%
    rownames_to_column() %>%
    mutate(model="4. Adjusted for race/ethnicity, age, Yost SES indicator") %>%
    mutate(age_group = "0 to 19 years")
Table3 <- rbind(a,b,c,d)  
```

# run a model with continuous as reference
```{r}
seer0to19 <- seer0to19%>%
  mutate(mcont66.2 = case_when(mcont66 == "Continuous" ~ 0,
                               mcont66 == "Discontinuous" ~ 1,
                               mcont66 == "Other" ~ 2,
                               mcont66 == "Not enrolled" ~ 3))

seer0to19$mcont66.2 <- factor(seer0to19$mcont66.2, levels = c(0:3), labels = c("Continuous", "Discontinuous", "Other", "Not enrolled"))

# mcont66.2 adjusted models 0 to 19
res.cox4.2 <- coxph(Surv(fumo, can_dead) ~ mcont66.2 + race_ethnicity +age_cat + Yost_ACS_2006_2010_quintile + cluster(state), data = seer0to19) # Model adjusted for race age and a measure of poverty
res.cox4.2
summary(res.cox4.2)
```

# Test of the ph assumption for adjusted models
```{r}
# adjusted models 0 to 19
Medflag <- cox.zph(res.cox2, terms = FALSE)
Medflag

# mcont66 adjusted models 0 to 19
Med4cat <- cox.zph(res.cox4, terms = FALSE)
Med4cat
# violated for both

plot(Medflag[1], resid =FALSE)
abline(h=0, col=2)

Medflag.surv <- survfit(Surv(fumo, can_dead) ~ In_Medicaid_Flag + race_ethnicity +age_cat + Yost_ACS_2006_2010_quintile + cluster(state), data = seer0to19) # get survival probabilities

plot(Medflag.surv[1], col=c("black", "red"), main="Medicaid_flag log log plot", fun="cloglog", xlab="Time", ylab="log(-log(S(t))") #  ph assumption violated

plot(Med4cat, resid = FALSE)

ggcoxdiagnostics(res.cox2, type = "schoenfeld")

ggcoxdiagnostics(res.cox4, type = "schoenfeld")  
```

# Run models in 12 month increments
```{r}
# Split a survival data set at specified times to form a counting process format
# For those with survival time of 0, add 0.5 mo because otherwise splitting doesn't work and it is more accurate

seer0to19 <- seer0to19 %>%
  mutate(fumo_half = if_else(fumo == 0, 0.5, fumo)) # need to add 0.5 or function doesn't work

seer0to19$patient_id <- as.numeric(rownames(seer0to19))

seer0to19.cp.format <- survSplit(data  = seer0to19,
                            cut   = c(12, 24, 36),       # cut at time 12,24,36
                            end   = "fumo_half",    # original survival time
                            event = "can_dead",   # event indicator
                            start = "start")    # will be created. zero by default

# Recoding
seer0to19.cp.format <- within(seer0to19.cp.format, {
    # Create new survival object
    SurvObj <- Surv(start, fumo_half, can_dead) 
    # Create interval indicator
    interval <- factor(start, levels = c(0,12, 24, 36), labels = c("First","Second", "Third", "Fourth"))
})

# Reordering
seer0to19.cp.format <- seer0to19.cp.format[with(seer0to19.cp.format, order(patient_id, fumo_half)),]  # sort by patient_id and then by fumo
summary(seer0to19$fumo_half)

```

# Run models based on intervals for Medicaid_flag
```{r}
# first interval
seer0to19first<-seer0to19.cp.format[which(seer0to19.cp.format$interval=="First"),]
# adjusted models 0 to 19
res.cox12 <- coxph(Surv(fumo, can_dead) ~ In_Medicaid_Flag + race_ethnicity +age_cat + 
                     Yost_ACS_2006_2010_quintile + cluster(state), data = seer0to19first) 
res.cox12
summary(res.cox12)

# second interval
seer0to19second<-seer0to19.cp.format[which(seer0to19.cp.format$interval=="Second"),]
# adjusted models 0 to 19
res.cox24 <- coxph(Surv(fumo, can_dead) ~ In_Medicaid_Flag + race_ethnicity +age_cat + 
                     Yost_ACS_2006_2010_quintile + cluster(state), data = seer0to19second) 
res.cox24
summary(res.cox24)

# third interval
seer0to19third<-seer0to19.cp.format[which(seer0to19.cp.format$interval=="Third"),]

# adjusted models 0 to 19
res.cox36 <- coxph(Surv(fumo, can_dead) ~ In_Medicaid_Flag + race_ethnicity +age_cat + 
                     Yost_ACS_2006_2010_quintile + cluster(state), data = seer0to19third) 
res.cox36
summary(res.cox36)

# fourth interval
seer0to19fourth<-seer0to19.cp.format[which(seer0to19.cp.format$interval=="Fourth"),]
# adjusted models 0 to 19
res.cox36plus <- coxph(Surv(fumo, can_dead) ~ In_Medicaid_Flag + race_ethnicity +age_cat + 
                     Yost_ACS_2006_2010_quintile + cluster(state), data = seer0to19fourth) 
res.cox36plus
summary(res.cox36plus)

# it seems the risk is highest later on in follow-up but this begs the question about which cancers are driving this?
```

# SUPPLEMENTARY TABLE 6a and Graph hazard ratios and confidence intervals across time for Medicaid_flag variable
```{r}
# need to get HRs in a dataframe
a<-as.data.frame(exp(cbind(HR = coef(res.cox12), confint(res.cox12))))
   a<-a[c(1),] %>%
    rownames_to_column() %>%
    mutate(model= 0) %>%
     mutate(Group = "Overall")

b<-as.data.frame(exp(cbind(HR = coef(res.cox24), confint(res.cox24))))
   b<-b[c(1),] %>%
    rownames_to_column() %>%
    mutate(model= 1) %>%
     mutate(Group = "Overall")

c<-as.data.frame(exp(cbind(HR = coef(res.cox36), confint(res.cox36))))
   c<-c[c(1),] %>%
    rownames_to_column() %>%
    mutate(model= 2) %>%
     mutate(Group = "Overall")

d <-as.data.frame(exp(cbind(HR = coef(res.cox36plus), confint(res.cox36plus))))
   d<-d[c(1),] %>%
    rownames_to_column() %>%
    mutate(model=3) %>%
     mutate(Group = "Overall")

SuppTable6a <- rbind(a,b,c,d)

SuppTable6a$model <- factor(SuppTable6a$model, levels = c(0:3), labels = c("0 to 12 months", "12 to 24 months","24 to 36 months", ">36 months"))

# graph HRs and 95% CIs
SuppTable6a %>%
  ggplot() +
  geom_point(aes(y=HR, x=model)) + 
  #scale_color_manual(values=c("black", "blue", "green", "red"), name="") + 
  geom_hline(yintercept =1, linetype=2) +
  theme(axis.text.y=element_blank()) +
  geom_errorbar(aes(y=HR, x=model, ymax = `2.5 %`, ymin = `97.5 %`), width = 0.1) +
  theme_bw() +
  theme(axis.line = element_line(colour = "black"),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_blank(),
    panel.background = element_blank(),
    axis.text.x=element_text(size=11, face = "bold", color = "black"),
    axis.text.y=element_text(size=11, face = "bold", color = "black"),
    axis.title.x=element_text(size=12, face = "bold", color = "black"),
    axis.title.y=element_text(size=12, face = "bold", color = "black"))+
  coord_flip() 
```

# SUPPLEMENTARY TABLE 6b: Which cancer is driving this? 
```{r}
# Split a survival data set at specified times to form a counting process format
# For those with survival time of 0, add 0.5 mo because otherwise splitting doesn't work and it is more accurate

cantype <- function(cantype, int, time){

df <- survSplit(data  = seer0to19 %>% 
                  filter(can_type == cantype),
                            cut   = c(12, 24, 36),       # cut at time 12,24,36
                            end   = "fumo_half",    # original survival time
                            event = "can_dead",   # event indicator
                            start = "start")    # will be created. zero by default

# Recoding
df <- within(df, {
    # Create new survival object
    SurvObj <- Surv(start, fumo_half, can_dead) 
    # Create interval indicator
    interval <- factor(start, levels = c(0,12, 24, 36), labels = c("First","Second", "Third", "Fourth"))
})

# Reordering
df <- df[with(df, order(patient_id, fumo_half)),]  # sort by patient_id and then by fumo

# run models based on intervals for Medicaid_flag
# first interval
df<-df[which(df$interval==int),]
mod<-coxph(Surv(fumo, can_dead) ~ In_Medicaid_Flag + race_ethnicity +age_cat + Yost_ACS_2006_2010_quintile + cluster(state), data = df) 
summary(mod)

# need to get HRs in a dataframe
a<-as.data.frame(exp(cbind(HR = coef(mod), confint(mod))))
  a<-a[c(1),] %>%
   rownames_to_column() %>%
    mutate(model= time,
          Group = cantype) 
}

# Run top three cancer types
a <-cantype("I. Leukemias, myeloproliferative diseases, and myelodysplastic diseases", "First", 0)
b <-cantype("I. Leukemias, myeloproliferative diseases, and myelodysplastic diseases", "Second", 1)
c <-cantype("I. Leukemias, myeloproliferative diseases, and myelodysplastic diseases", "Third", 2)
d <-cantype("I. Leukemias, myeloproliferative diseases, and myelodysplastic diseases", "Fourth", 3)

# Validate function for one cancer type
test <- survSplit(data  = seer0to19 %>% 
                  filter(can_type == "I. Leukemias, myeloproliferative diseases, and myelodysplastic diseases"),
                            cut   = c(12, 24, 36),       # cut at time 12,24,36
                            end   = "fumo_half",    # original survival time
                            event = "can_dead",   # event indicator
                            start = "start")    # will be created. zero by default

# Recoding
test<- within(test, {
    # Create new survival object
    SurvObj <- Surv(start, fumo_half, can_dead) 
    # Create interval indicator
    interval <- factor(start, levels = c(0,12, 24, 36), labels = c("First","Second", "Third", "Fourth"))
})

# Reordering
test <- test[with(test, order(patient_id, fumo_half)),]  # sort by patient_id and then by fumo

# run models based on intervals for Medicaid_flag
# first interval
test <- test[which(test$interval=="First"),]
mod<-coxph(Surv(fumo, can_dead) ~ In_Medicaid_Flag + race_ethnicity +age_cat + Yost_ACS_2006_2010_quintile + cluster(state), data = test) 
summary(mod)

# need to get HRs in a dataframe
a.2<-as.data.frame(exp(cbind(HR = coef(mod), confint(mod))))
  a.2<-a.2[c(1),] %>%
   rownames_to_column() %>%
    mutate(model= 0,
          Group = "I. Leukemias, myeloproliferative diseases, and myelodysplastic diseases")

# Not presented
# e <-cantype("II. Lymphomas and reticuloendothelial neoplasms", "First", 0) # converg. issues
# f <-cantype("II. Lymphomas and reticuloendothelial neoplasms", "Second", 1) # converg. issues
# g <-cantype("II. Lymphomas and reticuloendothelial neoplasms", "Third", 2) # converg. issues
# h <-cantype("II. Lymphomas and reticuloendothelial neoplasms", "Fourth", 3)

i <-cantype("III. CNS and miscellaneous intracranial and intraspinal neoplasms", "First", 0) 
j <-cantype("III. CNS and miscellaneous intracranial and intraspinal neoplasms", "Second", 1)
k <-cantype("III. CNS and miscellaneous intracranial and intraspinal neoplasms", "Third", 2) 
l <-cantype("III. CNS and miscellaneous intracranial and intraspinal neoplasms", "Fourth", 3)

SuppTable6b<-rbind(a,b,c,d,i,j,k,l)

SuppTable6b$model <- factor(SuppTable6b$model, levels = c(0:3), labels = c("0 to 12 months", "12 to 24 months","24 to 36 months", ">36 months"))

table(seer0to19$can_type)
```

# Group by cancer type and plot
```{r}
SuppTable6b %>%
  group_by(Group) %>%
  ggplot() +
  geom_point(aes(y=HR, x=model)) + 
  geom_hline(yintercept =1, linetype=2) +
  theme(axis.text.y=element_blank()) +
  geom_errorbar(aes(y=HR, x=model, ymax = `2.5 %`, ymin = `97.5 %`), width = 0.1) +
  theme_bw() +
  theme(axis.line = element_line(colour = "black"),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_blank(),
    panel.background = element_blank(),
    axis.text.x=element_text(size=11, face = "bold", color = "black"),
    axis.text.y=element_text(size=11, face = "bold", color = "black"),
    axis.title.x=element_text(size=12, face = "bold", color = "black"),
    axis.title.y=element_text(size=12, face = "bold", color = "black"))+
  coord_flip() + 
  facet_wrap(~Group, nrow=4, ncol=2, scales = "free") +
  labs( x = "Time interval") + 
  theme(strip.placement = "inside", legend.position = "bottom",  strip.text = element_text(size = 9, face="bold", color="black"), strip.background = element_rect(fill="lightgray"), plot.title = element_text(face="bold"), legend.text=element_text(size=12, face="bold"))
```

# run models based on intervals for mcont66
```{r}
# first interval
res.cox12m3 <- coxph(Surv(fumo, can_dead) ~ mcont66 + race_ethnicity +age_cat + 
                     Yost_ACS_2006_2010_quintile + cluster(state), data = seer0to19first) 
summary(res.cox12m3)

# second interval
seer0to19second<-seer0to19.cp.format[which(seer0to19.cp.format$interval=="Second"),]
# adjusted models 0 to 19
res.cox24m3 <- coxph(Surv(fumo, can_dead) ~ mcont66 + race_ethnicity +age_cat + 
                     Yost_ACS_2006_2010_quintile + cluster(state), data = seer0to19second) 
summary(res.cox24m3)

# third interval
seer0to19third<-seer0to19.cp.format[which(seer0to19.cp.format$interval=="Third"),]
# adjusted models 0 to 19
res.cox36m3 <- coxph(Surv(fumo, can_dead) ~ mcont66 + race_ethnicity +age_cat + 
                     Yost_ACS_2006_2010_quintile + cluster(state), data = seer0to19third) 
summary(res.cox36m3)

# fourth interval
seer0to19fourth<-seer0to19.cp.format[which(seer0to19.cp.format$interval=="Fourth"),]
# adjusted models 0 to 19
res.cox36plusm3 <- coxph(Surv(fumo, can_dead) ~ mcont66 + race_ethnicity +age_cat + 
                     Yost_ACS_2006_2010_quintile + cluster(state), data = seer0to19fourth) 
summary(res.cox36plusm3)
```

# SUPPLEMETARY TABLE 6c: Graph hazard ratios and confidence intervals across time for mcont66 variable
```{r}
# need to get HRs in a dataframe
a<-as.data.frame(exp(cbind(HR = coef(res.cox12m3), confint(res.cox12m3))))
   a<-a[c(1,2,3),] %>%
    rownames_to_column() %>%
    mutate(model= 0) %>%
    mutate(Group = "Overall")

b<-as.data.frame(exp(cbind(HR = coef(res.cox24m3), confint(res.cox24m3))))
   b<-b[c(1,2,3),] %>%
    rownames_to_column() %>%
    mutate(model= 1)  %>%
    mutate(Group ="Overall")

c<-as.data.frame(exp(cbind(HR = coef(res.cox36m3), confint(res.cox36m3))))
   c<-c[c(1,2,3),] %>%
    rownames_to_column() %>%
    mutate(model= 2)  %>%
    mutate(Group ="Overall")

d <-as.data.frame(exp(cbind(HR = coef(res.cox36plusm3), confint(res.cox36plusm3))))
   d<-d[c(1,2,3),] %>%
    rownames_to_column() %>%
    mutate(model=3)  %>%
    mutate(Group ="Overall")

SuppTable6c<- rbind(a,b,c,d)

SuppTable6c$model <- factor(SuppTable6c$model, levels = c(0:3), labels = c("0 to 12 months", "12 to 24 months","24 to 36 months", ">36 months"))

# graph HRs and 95% CIs
SuppTable6c %>%
  ggplot() +
  geom_point(aes(y=HR, x=model)) + 
  geom_hline(yintercept =1, linetype=2) +
  theme(axis.text.y=element_blank()) +
  geom_errorbar(aes(y=HR, x=model, ymax = `2.5 %`, ymin = `97.5 %`), width = 0.1) +
  theme_bw() +
  theme(axis.line = element_line(colour = "black"),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_blank(),
    panel.background = element_blank(),
    axis.text.x=element_text(size=11, face = "bold", color = "black"),
    axis.text.y=element_text(size=11, face = "bold", color = "black"),
    axis.title.x=element_text(size=12, face = "bold", color = "black"),
    axis.title.y=element_text(size=12, face = "bold", color = "black"))+
  coord_flip() +
 facet_wrap(~rowname, nrow=1, ncol=3) +
  labs(title = "Category", x = "time") + 
  theme(strip.placement = "inside", legend.position = "bottom",  strip.text = element_text(size = 9, face="bold", color="black"), strip.background = element_rect(fill="lightgray"), plot.title = element_text(face="bold"), legend.text=element_text(size=9, face="bold"))
```
# SUPPLEMENTARY TABLE 6: HRs by time interval
```{r}
SuppTable6 <- rbind(SuppTable6a, SuppTable6b, SuppTable6c)
```

# Effect modification by sex
```{r}
emsex1 <- coxph(Surv(fumo, can_dead) ~ In_Medicaid_Flag + sex + race_ethnicity +age_cat + Yost_ACS_2006_2010_quintile + cluster(state), data = seer0to19) # Model adjusted for race age and a measure of poverty
summary(emsex1)

emsex2 <- coxph(Surv(fumo, can_dead) ~ In_Medicaid_Flag + sex + race_ethnicity +age_cat + Yost_ACS_2006_2010_quintile + In_Medicaid_Flag*sex + cluster(state), data = seer0to19) # Model adjusted for race age and a measure of poverty
summary(emsex2)

lrtest(emsex1, emsex2) # not significant
```

# SUPPLEMENTARY TABLE 7a: Test effect modification by age
```{r}
# age_cat*Medicaid flag
res.cox2em <- coxph(Surv(fumo, can_dead) ~ In_Medicaid_Flag + race_ethnicity + age_cat + Yost_ACS_2006_2010_quintile + In_Medicaid_Flag*age_cat + cluster(state), data = seer0to19) # Model adjusted for race age and a measure of poverty
summary(res.cox2em)

lrtest(res.cox2em, res.cox2)

# significant effect modification. Run models by age group.

cox <-function(agegroup, var){
  x<<-seer0to19 %>%
    filter(age_cat == agegroup)

model<-coxph(Surv(fumo, can_dead) ~ var + race_ethnicity + Yost_ACS_2006_2010_quintile + cluster(state), data = x) # Model adjusted for race  and a measure of poverty

a<-as.data.frame(exp(cbind(HR = coef(model), confint(model))))
   a<-a[c(1),] %>%
    rownames_to_column() %>%
    mutate(model="Adjusted for race_ethnicty and Yost SES indicator") %>%
    mutate(Group = agegroup) 
}

cox2 <-function(agegroup, var){
  x<<-seer0to19 %>%
    filter(age_cat == agegroup)

model<-coxph(Surv(fumo, can_dead) ~ var + race_ethnicity + Yost_ACS_2006_2010_quintile + cluster(state), data = x) # Model adjusted for race  and a measure of poverty

a<-as.data.frame(exp(cbind(HR = coef(model), confint(model))))
   a<-a[c(1,2,3),] %>%
    rownames_to_column() %>%
    mutate(model="Adjusted for race_ethnicty and Yost SES indicator") %>%
    mutate(Group = agegroup) 
}

a<-cox("0-4", x$In_Medicaid_Flag)
b<-cox("5-9", x$In_Medicaid_Flag)
c<-cox("10-14", x$In_Medicaid_Flag)
d<-cox("15-19", x$In_Medicaid_Flag)

e<-cox2("0-4", x$mcont66)
f<-cox2("5-9", x$mcont66)
g<-cox2("10-14", x$mcont66)
h<-cox2("15-19",  x$mcont66)

SuppTable7a <- rbind(a,b,c,d,e,f,g,h)
```

# SUPPLEMENTARY TABLE 7b: Test effect modification by race
```{r}
# race_ethnicity*Medicaid flag
res.cox2emr1 <- coxph(Surv(fumo, can_dead) ~ In_Medicaid_Flag + race_ethnicity + age_cat + Yost_ACS_2006_2010_quintile + cluster(state), data = seer0to19) # Model adjusted for race age and Yost SES indicator
summary(res.cox2emr1)

res.cox2emr <- coxph(Surv(fumo, can_dead) ~ In_Medicaid_Flag + race_ethnicity + age_cat + Yost_ACS_2006_2010_quintile + In_Medicaid_Flag*race_ethnicity + cluster(state), data = seer0to19) # Model adjusted for race age and Yost SES indicator
summary(res.cox2emr)

lrtest(res.cox2emr, res.cox2emr1)

# Significant effect modification. Run models by race_ethnicity.
cox <-function(race_e, var){
  x<<-seer0to19 %>%
    filter(race_ethnicity == race_e)

model<-coxph(Surv(fumo, can_dead) ~ var  + age_cat + Yost_ACS_2006_2010_quintile + cluster(state), data = x) # Model adjusted for race  and a Yost SES indicator

a<-as.data.frame(exp(cbind(HR = coef(model), confint(model))))
   a<-a[c(1),] %>%
    rownames_to_column() %>%
    mutate(model="Adjusted for age cqtegory and Yost SES indicator") %>%
    mutate(Group = race_e)

}

cox2 <-function(race_e, var){
  x<<-seer0to19 %>%
    filter(race_ethnicity == race_e)

model<-coxph(Surv(fumo, can_dead) ~ var   + age_cat + Yost_ACS_2006_2010_quintile + cluster(state), data = x) # Model adjusted for race  and a measure of poverty

a<-as.data.frame(exp(cbind(HR = coef(model), confint(model))))
   a<-a[c(1,2,3),] %>%
    rownames_to_column() %>%
    mutate(model="Adjusted for race_ethnicty and Yost SES indicator") %>%
    mutate(Group = race_e) 
}

a<-cox("Non-Hispanic White", x$In_Medicaid_Flag) 
b<-cox("Non-Hispanic Black",  x$In_Medicaid_Flag)
c<-cox("Non-Hispanic API", x$In_Medicaid_Flag)
d<-cox("Non-Hispanic Other",  x$In_Medicaid_Flag)
e<-cox("Hispanic", x$In_Medicaid_Flag)

f<-cox2("Non-Hispanic White",  x$mcont66)
g<-cox2("Non-Hispanic Black",x$mcont66)
h<-cox2("Non-Hispanic API", x$mcont66)
i<-cox2("Non-Hispanic Other",  x$mcont66)
j<-cox2("Hispanic", x$mcont66)

SuppTable7b <- rbind(a,b,c,d,e,f,g,h,i,j)

```

# SUPPLEMENTARY TABLE 7c: Test effect modification by cancer type
```{r}
# can_type adjusted
res.cox5 <- coxph(Surv(fumo, can_dead) ~ In_Medicaid_Flag + race_ethnicity + age_cat + Yost_ACS_2006_2010_quintile + can_type + cluster(state), data = seer0to19) # Model adjusted for race age and Yost SES index
summary(res.cox5)

# can_type*Medicaid flag
res.cox5em <- coxph(Surv(fumo, can_dead) ~ In_Medicaid_Flag + race_ethnicity + age_cat + Yost_ACS_2006_2010_quintile + can_type + In_Medicaid_Flag*can_type + cluster(state), data = seer0to19) # Model adjusted for race age and Yost SES index
summary(res.cox5em)
lrtest(res.cox5, res.cox5em)

# Significant effect modification. Run models by cancer type.
cox <-function(can_type_e, var){
  x<<-seer0to19 %>%
    filter(can_type == can_type_e)

model<-coxph(Surv(fumo, can_dead) ~ var  + age_cat + race_ethnicity + Yost_ACS_2006_2010_quintile + cluster(state), data = x) # Model adjusted for race, age category, and a measure of poverty

a<-as.data.frame(exp(cbind(HR = coef(model), confint(model))))
   a<-a[c(1),] %>%
     rownames_to_column() %>%
    mutate(model="Adjusted for age cqtegory, race, and Yost SES indicator") %>%
    mutate(Group = can_type_e)

}

cox2 <-function(can_type_e, var){
  x<<-seer0to19 %>%
    filter(can_type == can_type_e)

model<-coxph(Surv(fumo, can_dead) ~ var + age_cat + race_ethnicity + Yost_ACS_2006_2010_quintile + cluster(state), data = x) # Model adjusted for race, age category, and Yost SES index

a<-as.data.frame(exp(cbind(HR = coef(model), confint(model))))
   a<-a[c(1,2,3),] %>%
    rownames_to_column() %>%
   mutate(model="Adjusted for age category, race_ethnicty, and Yost SES index") %>%
    mutate(Group = can_type_e) 
}

a<-cox("I. Leukemias, myeloproliferative diseases, and myelodysplastic diseases", x$In_Medicaid_Flag) 
b<-cox("II. Lymphomas and reticuloendothelial neoplasms",  x$In_Medicaid_Flag)
c<-cox("III. CNS and miscellaneous intracranial and intraspinal neoplasms", x$In_Medicaid_Flag)
d<-cox("IV. Neuroblastoma and other peripheral nervous cell tumors", x$In_Medicaid_Flag)
# e<-cox("V. Retinoblastoma", x$In_Medicaid_Flag) # convergence issues
f<-cox("VI. Renal tumors",x$In_Medicaid_Flag)
g<-cox("VII. Hepatic tumors", x$In_Medicaid_Flag) 
h<-cox("VIII. Malignant bone tumors",  x$In_Medicaid_Flag)
i<-cox("IX. Soft tissue and other extraosseous sarcomas",  x$In_Medicaid_Flag)
j<-cox("X. Germ cell tumors, trophoblastic tumors, and neoplasms of gonads",  x$In_Medicaid_Flag)
k<-cox("XI. Other malignant epithelial neoplasms and malignant melanomas", x$In_Medicaid_Flag)
# l<-cox("XII. Other and unspecified malignant neoplasms", x$In_Medicaid_Flag) # convergence issues

m<-cox2("I. Leukemias, myeloproliferative diseases, and myelodysplastic diseases", x$mcont66)
n<-cox2("II. Lymphomas and reticuloendothelial neoplasms", x$mcont66) 
o<-cox2("III. CNS and miscellaneous intracranial and intraspinal neoplasms", x$mcont66)
p<-cox2("IV. Neuroblastoma and other peripheral nervous cell tumors", x$mcont66)
# q<-cox2("V. Retinoblastoma", x$mcont66)# convergence issues
r<-cox2("VI. Renal tumors", x$mcont66)
s<-cox2("VII. Hepatic tumors", x$mcont66)
t<-cox2("VIII. Malignant bone tumors", x$mcont66)
u<-cox2("IX. Soft tissue and other extraosseous sarcomas", x$mcont66)
v<-cox2("X. Germ cell tumors, trophoblastic tumors, and neoplasms of gonads", x$mcont66)
w<-cox2("XI. Other malignant epithelial neoplasms and malignant melanomas", x$mcont66)
# y<-cox2("XII. Other and unspecified malignant neoplasms", x$mcont66) # convergence issues

SuppTable7c <- rbind(a,b,c,d,f,h,i,j,k, m,n,o,p,r,s,t,u,v,w)
```

# SUPPLEMENTARY TABLE 7: Effect modifcation table
```{r}
# combine age, race/ethnicity, and cancer type specific results
SuppTable7<- rbind(SuppTable7a, SuppTable7b, SuppTable7c)
table(SuppTable7$rowname)

SuppTable7 <- SuppTable7 %>%
  rename(Category = rowname) %>%
  mutate(Category = case_when(Category == "varContinuous" ~ "Continuous",
            Category == "varDiscontinuous" ~ "Discontinuous",
            Category == "varOther" ~ "Other",
            Category == "varEnrolled" ~ "Enrolled")) %>%
  mutate(Group2 = case_when(Group %in% c("0-4", "5-9", "10-14", "15-19") ~ "Age",
                            Group %in% c("I. Leukemias, myeloproliferative diseases, and myelodysplastic diseases", "II. Lymphomas and reticuloendothelial neoplasms", "III. CNS and miscellaneous intracranial and intraspinal neoplasms", "IV. Neuroblastoma and other peripheral nervous cell tumors", "V. Retinoblastoma", "VI. Renal tumors", "VII. Hepatic tumors", "VIII. Malignant bone tumors", "IX. Soft tissue and other extraosseous sarcomas",  "X. Germ cell tumors, trophoblastic tumors, and neoplasms of gonads", "XI. Other malignant epithelial neoplasms and malignant melanomas", "XII. Other and unspecified malignant neoplasms") ~ "CancerType",
                            Group %in% c("Non-Hispanic White", "Non-Hispanic Black", "Non-Hispanic API", "Non-Hispanic Other", "Hispanic") ~ "Race"))

# need abbreviations for plot fitting

SuppTable7 <- SuppTable7 %>%
  mutate(Abbrev = case_when(Group == "I. Leukemias, myeloproliferative diseases, and myelodysplastic diseases" ~ "LEUK",
                            Group =="II. Lymphomas and reticuloendothelial neoplasms" ~ "LYMPH",
                            Group == "III. CNS and miscellaneous intracranial and intraspinal neoplasms" ~ "CNS",
                            Group == "IV. Neuroblastoma and other peripheral nervous cell tumors" ~ "NB",
                            Group == "V. Retinoblastoma" ~ "RB",
                            Group == "VI. Renal tumors" ~ "RT",
                            Group == "VII. Hepatic tumors" ~ "HT",
                            Group == "VIII. Malignant bone tumors" ~ "MBT",
                            Group == "IX. Soft tissue and other extraosseous sarcomas" ~ "STS", 
                            Group == "X. Germ cell tumors, trophoblastic tumors, and neoplasms of gonads" ~ "GCT",
                            Group == "XI. Other malignant epithelial neoplasms and malignant melanomas" ~ "OTHEP",
                            Group == "XII. Other and unspecified malignant neoplasms" ~ "OTHUN",
                            Group == "Non-Hispanic White" ~ "NHW",
                            Group == "Non-Hispanic Black" ~ "NHB",
                            Group == "Non-Hispanic API" ~ "NHAPI",
                            Group == "Non-Hispanic Other"~ "NHOTH",
                            Group == "Hispanic" ~ "HISP",
                            Group == "0-4" ~ "0 to 4",
                            Group == "5-9" ~ "5 to 9",
                            Group == "10-14" ~ "10 to 14",
                            Group == "15-19" ~ "15 to 19"))

SuppTable7$Abbrev <- factor(SuppTable7$Abbrev, levels = c("LEUK", "LYMPH", "CNS","NB", "RB",
                                                  "RT", "HT", "MBT", "STS", "GCT",
                                                  "OTHEP", "OTHUN", "NHW", "NHB",  "NHAPI",
                                                  "NHOTH", "HISP", "0 to 4", "5 to 9", "10 to 14", "15 to 19"))
SuppTable7$Abbrev <- fct_rev(SuppTable7$Abbrev)
```

# Supplementary FIGURE 1: Effect modification by age, race, cancer type
- Medicaid enrollment variable
```{r}
plot1 <- SuppTable7 %>%
  filter(Category == "Enrolled", Group2 == "Age") %>%
ggplot() +
  geom_point(aes(y=HR, x = Abbrev, group=Category, color = Category)) + 
  geom_hline(yintercept =1, linetype=2) +
  theme(axis.text.y=element_blank()) +
  geom_errorbar(aes(y=HR, x=Abbrev, group = Category, color = Category, ymax = `2.5 %`, ymin = `97.5 %`), width = 0.1) +
  theme_bw() +
  theme(axis.line = element_line(colour = "black"),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_blank(),
    panel.background = element_blank(),
    axis.text.x=element_text(size=11, face = "bold", color = "black"),
    axis.text.y=element_text(size=11, face = "bold", color = "black"),
    axis.title.x=element_text(size=12, face = "bold", color = "black"),
    axis.title.y=element_text(size=12, face = "bold", color = "black"))+
  coord_flip() +
  labs(color = NULL, title = " ", x = "") +
  ylim(0.5,3) +
  theme(strip.placement = "inside", legend.position = "bottom",  strip.text = element_text(size = 9, face="bold", color="black"), strip.background = element_rect(fill="lightgray"), plot.title = element_text(face="bold"), legend.text=element_text(size=9, face="bold")) + 
          ggtitle("A") + theme(plot.title = element_text(face="bold")) +
  scale_color_manual(breaks = c("Enrolled"), values=c("black")) +
  theme(legend.title=element_blank())
plot1

plot2 <-SuppTable7 %>%
  filter(Category == "Enrolled", Group2 == "Race") %>% 
  ggplot() +
  geom_point(aes(y=HR, x = Abbrev, group = Category, color = Category)) + 
  geom_hline(yintercept =1, linetype=2) +
  theme(axis.text.y=element_blank()) +
  geom_errorbar(aes(y=HR, x=Abbrev, group = Category, color = Category, ymax = `2.5 %`, ymin = `97.5 %`), width = 0.1) +
  theme_bw() +
  theme(axis.line = element_line(colour = "black"),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_blank(),
    panel.background = element_blank(),
    axis.text.x=element_text(size=11, face = "bold", color = "black"),
    axis.text.y=element_text(size=11, face = "bold", color = "black"),
    axis.title.x=element_text(size=12, face = "bold", color = "black"),
    axis.title.y=element_text(size=12, face = "bold", color = "black"))+
  coord_flip() +
  labs(color = NULL, title = " ", x = "") +
  ylim(0.5,5) +
  theme(strip.placement = "inside", legend.position = "bottom",  strip.text = element_text(size = 9, face="bold", color="black"), strip.background = element_rect(fill="lightgray"), plot.title = element_text(face="bold"), legend.text=element_text(size=9, face="bold")) + ggtitle("C") +theme(plot.title = element_text(face="bold")) +
  scale_color_manual(breaks = c("Enrolled"),
                        values=c("black")) +
  theme(legend.title=element_blank())

plot2

plot3 <-SuppTable7 %>%
  filter(Category == "Enrolled", Group2 == "CancerType", Abbrev != "OTHUN") %>%
  ggplot() +
  geom_point(aes(y=HR, x = Abbrev, group = Category, color = Category)) + 
  geom_hline(yintercept =1, linetype=2) +
  theme(axis.text.y=element_blank()) +
  geom_errorbar(aes(y=HR, x=Abbrev, group = Category, color = Category, ymax = `2.5 %`, ymin = `97.5 %`), width = 0.1) +
  theme_bw() +
  theme(axis.line = element_line(colour = "black"),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_blank(),
    panel.background = element_blank(),
    axis.text.x=element_text(size=11, face = "bold", color = "black"),
    axis.text.y=element_text(size=11, face = "bold", color = "black"),
    axis.title.x=element_text(size=12, face = "bold", color = "black"),
    axis.title.y=element_text(size=12, face = "bold", color = "black"))+
  coord_flip() +
  labs(color = NULL, title = " ", x = "") +
  ylim(0.5,3.6) +
  theme(strip.placement = "inside", legend.position = "bottom",  strip.text = element_text(size = 9, face="bold", color="black"), strip.background = element_rect(fill="lightgray"), plot.title = element_text(face="bold"), legend.text=element_text(size=9, face="bold")) + ggtitle("E") +theme(plot.title = element_text(face="bold")) +
  scale_color_manual(breaks = c("Enrolled"),
                        values=c("black")) +
  theme(legend.title=element_blank())

plot3
```
# Supplementary FIGURE 1: Effect modification by age, race, cancer type
- mcont66
```{r}
SuppTable7$Category <- factor(SuppTable7$Category, levels = c("Enrolled", "Discontinuous", "Other",  "Continuous"))

plot4 <- SuppTable7 %>%
  filter(Category != "Enrolled", Group2 == "Age") %>%
ggplot() +
  geom_point(aes(y=HR, x = Abbrev, group=Category, color = Category), na.rm=TRUE, position=position_dodge(width = 0.9)) + 
  geom_hline(yintercept =1, linetype=2) +
   geom_vline(xintercept = c(1.5, 2.5,3.5), linetype=2, color = "lightgray") +
  theme(axis.text.y=element_blank()) +
  geom_errorbar(position=position_dodge(width = 0.9), aes(y=HR, x=Abbrev, group=Category, color = Category, ymax = `2.5 %`, ymin = `97.5 %`), width = 0.1) +
  theme_bw() +
  theme(axis.line = element_line(colour = "black"),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_blank(),
    panel.background = element_blank(),
    axis.text.x=element_text(size=11, face = "bold", color = "black"),
    axis.text.y=element_text(size=11, face = "bold", color = "black"),
    axis.title.x=element_text(size=12, face = "bold", color = "black"),
    axis.title.y=element_text(size=12, face = "bold", color = "black"))+
  coord_flip() +
  labs(color = NULL, title = " ", x = "") +
  ylim(0.5,3) +
  theme(strip.placement = "inside", legend.position = "bottom",  strip.text = element_text(size = 9, face="bold", color="black"), strip.background = element_rect(fill="lightgray"), plot.title = element_text(face="bold"), legend.text=element_text(size=9, face="bold")) + ggtitle("B") +theme(plot.title = element_text(face="bold")) +
  scale_color_manual(breaks = c("Discontinuous", "Continuous", "Other"),
                        values=c("red", "cyan3", "blue")) +
  theme(legend.title=element_blank())

plot4

plot5 <-SuppTable7 %>%
  filter(Category != "Enrolled", Group2 == "Race") %>% 
  ggplot() +
  geom_point(aes(y=HR, x = Abbrev, group=Category, color = Category), na.rm=TRUE, position=position_dodge(width = 0.9)) + 
  geom_hline(yintercept =1, linetype=2) +
  geom_vline(xintercept =c(1.5, 2.5,3.5, 4.5), linetype=2, color = "lightgray") +
  theme(axis.text.y=element_blank()) +
  geom_errorbar(position=position_dodge(width = 0.9), aes(y=HR, x=Abbrev, group=Category, color = Category, ymax = `2.5 %`, ymin = `97.5 %`), width = 0.1) +
  theme_bw() +
  theme(axis.line = element_line(colour = "black"),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_blank(),
    panel.background = element_blank(),
    axis.text.x=element_text(size=11, face = "bold", color = "black"),
    axis.text.y=element_text(size=11, face = "bold", color = "black"),
    axis.title.x=element_text(size=12, face = "bold", color = "black"),
    axis.title.y=element_text(size=12, face = "bold", color = "black"))+
  coord_flip() +
  labs(color = NULL,title = " ", x = "") +
  ylim(0.5,7) +
  theme(strip.placement = "inside", legend.position = "bottom",  strip.text = element_text(size = 9, face="bold", color="black"), strip.background = element_rect(fill="lightgray"), plot.title = element_text(face="bold"), legend.text=element_text(size=9, face="bold")) + ggtitle("D") +theme(plot.title = element_text(face="bold")) +
  scale_color_manual(breaks = c("Discontinuous", "Continuous", "Other"),
                        values=c("red", "cyan3", "blue")) +
  theme(legend.title=element_blank())

plot5

plot6 <-SuppTable7 %>%
  filter(Category != "Enrolled", Group2 == "CancerType", !Abbrev %in% c("RB", "OTHUN")) %>%
  ggplot() +
  geom_point(aes(y=HR, x = Abbrev, group=Category, color = Category), na.rm=TRUE, position=position_dodge(width = 0.9)) + 
  geom_hline(yintercept =1, linetype=2) +
  geom_vline(xintercept =c(1.5, 2.5,3.5,4.5,5.5,6.5,7.5,8.5,9.5), linetype=2, color = "lightgray") +
  theme(axis.text.y=element_blank()) +
  geom_errorbar(position=position_dodge(width = 0.9), aes(y=HR, x=Abbrev, group=Category, color = Category, ymax = `2.5 %`, ymin = `97.5 %`), width = 0.1) +
  theme_bw() +
  theme(axis.line = element_line(colour = "black"),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_blank(),
    panel.background = element_blank(),
    axis.text.x=element_text(size=11, face = "bold", color = "black"),
    axis.text.y=element_text(size=11, face = "bold", color = "black"),
    axis.title.x=element_text(size=12, face = "bold", color = "black"),
    axis.title.y=element_text(size=12, face = "bold", color = "black"))+
  coord_flip() +
  labs(color = NULL, title = " ", x = "") +
  ylim(0.5,4.5) +
  theme(strip.placement = "inside", legend.position = "bottom",  strip.text = element_text(size = 9, face="bold", color="black"), strip.background = element_rect(fill="lightgray"), plot.title = element_text(face="bold"), legend.text=element_text(size=9, face="bold")) + ggtitle("F") +theme(plot.title = element_text(face="bold"))+
  scale_color_manual(breaks = c("Discontinuous", "Continuous", "Other"),
                        values=c("red", "cyan3", "blue"))+
  theme(legend.title=element_blank())

plot6
```
# Supplementary FIGURE 1: Export
```{r}
pdf("Supplementary Figure 1.pdf", width=12, height=10, paper="USr")
plot1 + plot2 + plot3 + plot4 + plot5 + plot6 + plot_layout(ncol=3, nrow=2, guides='collect') &
  theme(legend.position='bottom')

dev.off()
```

# Clean up tables
```{r}
Table2 <- Table2%>%
  rename(`lower 95% CI` = cluster.lowCI,
         `upper 95% CI` = cluster.highCI) %>%
  mutate_if(is.numeric, round, 2)

Table2$`OR (95% CI)` <-paste0(Table2$OR, " (", Table2$`lower 95% CI`, " to ", Table2$`upper 95% CI`, ")")

Table3 <- Table3 %>% 
  rename(Category = rowname,
         "lower 95% CI" = "2.5 %",
         "upper 95% CI" = "97.5 %") %>%
  mutate(Category = case_when(Category == "In_Medicaid_FlagEnrolled" ~ "Enrolled",
            Category == "mcont66Discontinuous" ~ "Discontinuous",
            Category == "mcont66Other" ~ "Other",
            Category == "mcont66Continuous" ~ "Continuous"))  %>%
  mutate_if(is.numeric, round, 2)

Table3$`HR (95% CI)` <-paste0(Table3$HR, " (", Table3$`lower 95% CI`, " to ", Table3$`upper 95% CI`, ")") 
```

# Supplementary Tables
```{r}
# 1
SuppTable1 <- SuppTable1 %>%
  rename(`lower 95% CI` = cluster.lowCI,
         `upper 95% CI` = cluster.highCI,
          `Cancer Type` = CancerType) %>%
  mutate_if(is.numeric, round, 2)

SuppTable1$`OR (95% CI)` <-paste0(SuppTable1$OR, " (", SuppTable1$`lower 95% CI`, " to ", SuppTable1$`upper 95% CI`, ")") 
SuppTable1 <- SuppTable1 %>%
  dplyr::select(1,5,6,7)

# 2
SuppTable2 <- SuppTable2 %>%
  rename(`lower 95% CI` = `Lower CI`,
         `upper 95% CI` = `Upper CI`)

SuppTable2$`Survival Percentage (95% CI)` <-paste0(SuppTable2$`Survival Percentage`, " (", SuppTable2$`lower 95% CI`, " to ", SuppTable2$`upper 95% CI`, ")")
# drop uneeded columns
SuppTable2 <- SuppTable2 %>%
  dplyr::select(1,5)

# 3
SuppTable3 <- SuppTable3 %>%
  rename(`lower 95% CI` = LowerCI,
         `upper 95% CI` = UpperCI)

SuppTable3$`Survival Percentage (95% CI)` <-paste0(SuppTable3$`Survival Percentage`, " (", SuppTable3$`lower 95% CI`, " to ", SuppTable3$`upper 95% CI`, ")")
SuppTable3 <- SuppTable3 %>%
  dplyr::select(1,5,6)

# 4
SuppTable4 <- SuppTable4 %>%
  rename(`lower 95% CI` = LowerCI,
         `upper 95% CI` = UpperCI)

SuppTable4$`Survival Percentage (95% CI)` <-paste0(SuppTable4$`Survival Percentage`, " (", SuppTable4$`lower 95% CI`, " to ", SuppTable4$`upper 95% CI`, ")")
SuppTable4 <- SuppTable4 %>%
  dplyr::select(1,5,6)

# 5
SuppTable5 <- SuppTable5 %>%
  rename(`lower 95% CI` = LowerCI,
         `upper 95% CI` = UpperCI)

SuppTable5$`Survival Percentage (95% CI)` <-paste0(SuppTable5$`Survival Percentage`, " (", SuppTable5$`lower 95% CI`, " to ", SuppTable5$`upper 95% CI`, ")")

SuppTable5 <- SuppTable5 %>%
  dplyr::select(1,5,6)

# 6
SuppTable6 <- SuppTable6 %>%
  rename(`Time interval` = model,
         Category = rowname,
         `lower 95% CI` = `2.5 %`,
         `upper 95% CI` = `97.5 %`) %>%
  mutate(Category = case_when(Category == "In_Medicaid_FlagEnrolled" ~ "Enrolled",
            Category == "mcont66Discontinuous" ~ "Discontinuous",
            Category == "mcont66Other" ~ "Other",
            Category == "mcont66Continuous" ~ "Continuous")) %>%

  mutate_if(is.numeric, round, 2)

SuppTable6$`HR (95% CI)` <-paste0(SuppTable6$HR, " (", SuppTable6$`lower 95% CI`, " to ", SuppTable6$`upper 95% CI`, ")") 
SuppTable6 <- SuppTable6 %>%
  dplyr::select(1,5,6,7)

# 7
SuppTable7 <- SuppTable7 %>%
  rename(`lower 95% CI` = `2.5 %`,
         `upper 95% CI` = `97.5 %`) %>%
  dplyr::select(-7, -8)  %>%
  mutate_if(is.numeric, round, 2)

SuppTable7$`HR (95% CI)` <-paste0(SuppTable7$HR, " (", SuppTable7$`lower 95% CI`, " to ", SuppTable7$`upper 95% CI`, ")") 
SuppTable7 <- SuppTable7 %>%
  dplyr::select(1,6,7)
```

# Export data to excel (Tables 2-3)
```{r}
sheets<-list("Table 2 (unadj stagedx enro)" = df4,
             "Table 2 (unadj stagedx mcont66)" = df8,
             "Table 2 (stage dx analysis)" = Table2,
             "Table 3 (overall CPH)" = Table3)
write.xlsx(sheets, file="SEER0to19analysis_tables.xlsx", keepNA=TRUE, overwrite=TRUE)
```

# Export supplementary tables to excel
```{r}
sheets <- list("Supp Table 1 (stage cantype)" = SuppTable1,
               "Supp Table 2 (surv overall)" = SuppTable2,
               "Supp Table 3 (surv prob age )" = SuppTable3,
               "Supp Table 4 (surv prob race )" = SuppTable4,
               "Supp Table 5 (surv prob can )" = SuppTable5,
               "Supp Table 6 (ph analysis)" = SuppTable6, 
               "Supp Table 7 (EM)" = SuppTable7)
write.xlsx(sheets, file="Supplementary tables.xlsx", showNA=TRUE, overwrite = TRUE)
```

# SOME Sensitivity analyses
# This code validates the function above for clustered SEs using binomial logistic (nnet object is hard to work with for any standard functions that calculate robust or clustered SEs) 
```{r}
# regional vs. in situ/localized
model1_logit<-glm(stage ~ In_Medicaid_Flag, data = seer0to19_stage[which(seer0to19_stage$stage %in% c("In situ/localized", "Regional")),], family = binomial(link = "logit"))
odds.n.ends(model1_logit) # these are CIs without robust SEs

# distant vs. in situ/localized
model1_logit2<-glm(stage ~ In_Medicaid_Flag, data = seer0to19_stage[which(seer0to19_stage$stage %in% c("In situ/localized", "Distant")),], family = binomial(link = "logit"))
odds.n.ends(model1_logit2)  # these are CIs without robust SEs

model1_logit3 <-glm(stage ~ mcont66, data = seer0to19_stage[which(seer0to19_stage$stage %in% c("In situ/localized", "Distant")),], family = binomial(link = "logit"))
odds.n.ends(model1_logit3)  # these are CIs without robust SEs

model1_logit4 <-glm(stage ~ mcont66, data = seer0to19_stage[which(seer0to19_stage$stage %in% c("In situ/localized", "Distant")),], family = binomial(link = "logit"))
odds.n.ends(model1_logit4)  # these are CIs without robust SEs
```

# get results with clustered standard errors (solution 5  here: https://stackoverflow.com/questions/16498849/logistic-regression-with-robust-clustered-standard-errors-in-r)
```{r}
model1_logit_RSE <- miceadds::glm.cluster(data=seer0to19_stage[which(seer0to19_stage$stage %in% c("In situ/localized", "Regional")),], formula= stage ~ In_Medicaid_Flag + race_cat2 + age_cat + Yost_ACS_2006_2010_quintile, cluster="state", family="binomial")
summary(model1_logit_RSE)
exp(0.156116634)
exp(0.156116634 + (1.96*0.02744999)) 
exp(0.156116634 - (1.96*0.02744999))

model2_logit_RSE <- miceadds::glm.cluster(data=seer0to19_stage[which(seer0to19_stage$stage %in% c("In situ/localized", "Distant")),], formula= stage ~ In_Medicaid_Flag + race_cat2 + age_cat + Yost_ACS_2006_2010_quintile, cluster="state", family="binomial")
summary(model2_logit_RSE)

exp(0.354153776)
exp(0.354153776 + (1.96*0.02011365)) 
exp(0.354153776 - (1.96*0.02011365))

# The results are the same.
```

### 2023-09-06 Continuity coding protocol.Rmd
Initial commit of continuity coding protocol

---
title: "Protocol for Medicaid continuity variable coding"
author: "Kim Johnson"
date: "2023-09-06"
output: html_document
---

### **A. Import and create dataframes and variables needed for continuity variable classifications**

**1.	Import mcaid dataframe and reduce to patient ids and monthly enrollment variables.**    
a.	The medicaid.enroll.txt file that was used to create the mcaid dataframe by coauthor DB contained all the enrollment values for each observation id’s months of linkage (mcaid1: mcaid96).  This dataframe was created in Stata.

**2.	Import seer0to39 dataframe**       
a.	The medicaid.cancer.txt file was used to create a file for those diagnosed at 0 to 39 years by coauthor KJ. This was done to reduce the size of the dataframe used for all paper-associated data management and analysis.

**3.	Create seer0to39_formerge dataframe that includes observations with SEER variable values as:**    
a.	Firstmalignantprimary_indicator = 1,
b.	In_Medicaid_Flag = 1,
c.	Sequence_number = 00 or 01 (this was necessary to exclude non-unique cases).

This creates a file with unique ids for continuity coding containing the SEER variables needed for step 5.

**4.	Merge mcaid and seer0to39_formerge dataframes keeping only ids in seer0to39_formerge. These are the observation ids that will be coded for continuity.**

**5.	 Create variables needed for defining continuity in section B**        
a.	Scale diagnosis and death month using SEER diagnosis/death month and year variables (Year_of_diagnosis, Month_of_diagnosis_recode, SEER_DateofDeath_Year, SEER_DateofDeath_Month) to months 1-96 (the months corresponding to available Medicaid linkage data) -> dxmo and dthmo)    
b.	Define earlylate variable for observations diagnosed in the first 6 (1-6) or last 6 (91-96) months of the Medicaid observation window using dxmo since enrollment is not fully observable for these cases and they need to be handled as special cases for continuity coding. This is defined as dxmo < 7 or dxmo >90 -> earlylate    
c.	Define status to indicate whether they were enrolled in the dxmo, not enrolled, or enrollment data was NA during the diagnosis month (coded as 1, 0, 9 respectively)    
d.	Define vitalstat_post6 variable to indicate whether they died within the diagnosis month or months 1-5 following the dxmo

### **B. Continuity coding**

**Non-special cases**     
**1. Did not die within 6 months of diagnosis and not diagnosed in months 1 to 6 or 91 to 96) (dataframes = mcaid0to39_df1 and mcaid0to39_df2)**       
&nbsp;•	Define pre_mo6 to indicate the number of months enrolled prior to the dx month. Possible values are 0:6 or NA    
&nbsp;•	Define post_mo6 to indicate the number of months enrolled after the diagnosis month. Possible values are 0:6 or NA.   
&nbsp;•	Define any_zero_pre and any_zero_post as binary variables if enrollment in any of the 6 months prior to diagnosis has a value of 0 or any of the 6 months following diagnosis = 0    
&nbsp;•	mcontb (continuity variable) =     
&nbsp;&nbsp;&nbsp;o	0 (continuous) if pre_mo6 = 6 and post_mo6 = 6 and status = 1,     
&nbsp;&nbsp;&nbsp;o	1 (discontinuous) if status = 0 or any_zero_pre = 0 or any_zero_post = 0,     
&nbsp;&nbsp;&nbsp;o	2 (other) is everyone else    
&nbsp;•	Validate with random samples of 10 observations each for mcontb = 0,1,2. Manually looked at patterns in each of the six months around diagnosis month for each observation to validate. If any observation’s classification did not validate, corrected code, and validated again until 100% validated in each sample.

**Special cases**    
**2. Died within 6 months of diagnosis, not diagnosed in months 1:6 or 91:96 (dataframe = mcaid0to39_df3)**    
&nbsp;•	pre_mo6 and any_zero_pre defined as for non-special cases    
&nbsp;•	sum_T1_post_todth defined as the sum of the enrollment values from the month after diagnosis to dthmo        
&nbsp;•	mcontb (continuity variable) =     
&nbsp;&nbsp;&nbsp;o	0 (continuous) if pre_mo6 = 6 and status = 1 and the sum_T1_post_todth ≥ dthmo-dxmo or dxmo = dthmo    
&nbsp;&nbsp;&nbsp;o	1 (discontinuous) if any_zero_pre = 1 or status = 0 or sum_T1_post_todth < dthmo – dxmo     
&nbsp;&nbsp;&nbsp;o	2 (other) is everyone else    
&nbsp;•	Validate as above

**3.  Those diagnosed in months 1:6 and did not die within 6 months following diagnosis (dataframe = mcaid0to39_df4)**    
&nbsp;•	post_mo6 defined same as in non-special cases    
&nbsp;•	can only define this group as discontinuous or other because not enough pre data    
&nbsp;•	mcontb (continuity variable) =     
&nbsp;&nbsp;&nbsp;o	1 (discontinuous) if status = 0 or post_mo <6 or enrollment in any months prior to the dxmo has a value of 0    
&nbsp;&nbsp;&nbsp;o	2 (other) is everyone else       
&nbsp;•	Validate as above

|**4.  Those diagnosed in months 1:6 and died within 6 months following diagnosis (dataframe = mcaid0to39_df4b)**    
&nbsp;• sum_T1_post_todth defined same as above    
&nbsp;•	can only define this group as discontinuous or other because not enough pre data    
&nbsp;mcontb (continuity variable) =            
&nbsp;&nbsp;&nbsp;o 1 (discontinuous) if status = 0 or sum_T1_post_todth < dthmo-dxmo or any months prior to the dxmo = 0     
&nbsp;&nbsp;&nbsp;o	2 (other) is everyone else    
&nbsp;•	Validate as above

**5.   Those diagnosed in months 91:96 and who did not die within 6 months following dx or died after month 96 (dataframe = mcaid0to39_df5)**    
&nbsp;•	pre_mo6 defined as for non-special cases    
&nbsp;•	can only define discontinuous or other because not enough post data    
&nbsp;•	mcontb (continuity variable) =     
&nbsp;&nbsp;&nbsp;o	1 (discontinuous) if any months prior to dxmo have a value of 0 or status = 0 or pre_mo6 < 6    
&nbsp;&nbsp;&nbsp;o	2 (other) is everyone else    
&nbsp;•	Validate as above

**6. Those who died within 6 months of diagnosis and who were diagnosed in months 91 to 96 (dataframe = mcaid0to39_df6)**    
&nbsp;•	pre_mo6 defined as for non-special cases    
&nbsp;•	can define as continuous, discontinuous, and other    
&nbsp;•	mcontb (continuity variable) =     
&nbsp;&nbsp;&nbsp;o	0 (continuous) = if pre_mo6 = 6 and status = 1 and enrollment data in all months from diagnosis month up to and including dthmo has a value of 1     
&nbsp;&nbsp;&nbsp;o	1 (discontinuous) if enrollment data in any month prior to dxmo has a value of 0 or pre_mo6 <6 or status = 0    
&nbsp;&nbsp;&nbsp;o	2 (other) is everyone else    
&nbsp;•	Validate as above

### **C. Merging**    
mcaid0to39 dataframe: Merge mcaid0to39_df1:mcaid0to39_df6 to get medicaid continuity classifications. This dataframe containing mcontb gets merged with the analytic dataset in Figure 1 for analysis.

### 2023-09-07 Manuscript Analysis code 0 to 19.Rmd
added code to calculate p-values for stage analysis discontinuous vs. continuous comparison

@@
 ```
+
 # Import files sent by SEER Medicaid (IMS), get data into right columns, and assign column names
@@
 ```
+
 # Relevel to determine if there is a difference between continuous and discontinuous
@@
 se <- sqrt(diag(var))
+z <- b/se
+p <- (1 - pnorm(abs(z), 0, 1))*2 # two-tailed z test, reported in manuscript
+p
 

### 2023-09-15 Manuscript Analysis code 0 to 19.Rmd
Minor code edits

@@
 
-#
 table(seer0to39$SEER_DateofDeath_Year, useNA = "always")
+
+table(seer0to39$can_type)
 
@@
 
-table1(~age_cat + sex  + race_ethnicity + Yost_ACS_2006_2010_quintile + In_Medicaid_Flag  + stage + fumo  + can_type + can_dead2|mcont66, droplevels = TRUE, seer0to19)
-
+table1(~age_cat + sex  + race_ethnicity + Yost_ACS_2006_2010_quintile + In_Medicaid_Flag  + stage + fumo  + can_type + can_dead2|mcont66, droplevels = TRUE, seer0to19 %>% filter(Agerecodewithsingleages_and_100 !=18 ))
+
+table1(~mcont66|age_cat, droplevels = TRUE, seer0to19%>% filter(Agerecodewithsingleages_and_100 <18 ))
 ```
@@
     pval.coord = c(2,55),
-    conf.int = TRUE)
-
+    conf.int = TRUE,
+    hline = 90)
+
+# plot1$plot + geom_hline(yintercept = 90)
 plot1
@@
 # mcont66
- plot2 <- ggsurvplot(fit = survfit(Surv(fumo, can_dead) ~ mcont66, data = seer0to19), 
+ plot2 <- ggsurvplot(fit = survfit(Surv(fumo, can_dead) ~ mcont66, data = seer0to19 ), 
     xlab = "Months from diagnosis", 
@@
 ```
-# SUPPLEMENTARY TABLE 2: Estimate survival time 0 to 19
-```{r}
+# SUPPLEMENTARY TABLE 2: Estimate survival time 0 to 19 at 160 mo
+```{r}
+# Overall
+summary(survfit(Surv(fumo, can_dead) ~ 1, data = seer0to19), times=60) # estimates 60 month survival time
+
 # In_Medicaid_Flag
-SuppTable2a <- summary(survfit(Surv(fumo, can_dead) ~ In_Medicaid_Flag, data = seer0to19), times = 60) # estimates 60 month survival time
-SuppTable2a <- as.data.frame(cbind(SuppTable2a$strata, round(SuppTable2a$surv*100, 2), round(SuppTable2a$lower*100, 2), round(SuppTable2a$upper*100,2))) # makes dataframe
-
-SuppTable2a <- setNames(SuppTable2a, c("Category", "Survival Percentage", "Lower CI", "Upper CI")) # gives column names
+SuppTable2a <- summary(survfit(Surv(fumo, can_dead) ~ In_Medicaid_Flag, data = seer0to19), times=60) # estimates 60 month survival time
+
+SuppTable2a <- as.data.frame(cbind(SuppTable2a$strata, SuppTable2a$time, round(SuppTable2a$surv*100, 2), round(SuppTable2a$lower*100, 2), round(SuppTable2a$upper*100,2))) # makes dataframe
+
+SuppTable2a <- setNames(SuppTable2a, c("Category", "Time (months)", "Survival Percentage", "Lower CI", "Upper CI")) # gives column names
 SuppTable2a <- SuppTable2a %>%
@@
 SuppTable2b <- summary(survfit(Surv(fumo, can_dead) ~ mcont66, data = seer0to19), times = 60)
-SuppTable2b <- as.data.frame(cbind(SuppTable2b$strata, round(SuppTable2b$surv*100, 2), round(SuppTable2b$lower*100, 2), round(SuppTable2b$upper*100,2)))
-
-SuppTable2b <- setNames(SuppTable2b, c("Category", "Survival Percentage", "Lower CI", "Upper CI"))
+SuppTable2b <- as.data.frame(cbind(SuppTable2b$strata, SuppTable2b$time, round(SuppTable2b$surv*100, 2), round(SuppTable2b$lower*100, 2), round(SuppTable2b$upper*100,2)))
+
+SuppTable2b <- setNames(SuppTable2b, c("Category", "Time (months)", "Survival Percentage", "Lower CI", "Upper CI"))
 SuppTable2b <- SuppTable2b %>%
@@
 
-# SUPPLEMENTARY TABLE 3: Populate a table with survival by age group
-```{r}
+# SUPPLEMENTARY TABLE 3: Populate a table with survival by age group at 60 mo
+```{r}
+age <- unique(seer0to19$age_cat)
 # In_Medicaid_Flag
-# calculate survival at 60 mo
-results <- data.frame(matrix(ncol=5, nrow=0)) #create the table
-columns<-c("Medicaid_Category", "Survival Percentage",  "LowerCI", "UpperCI", "Age_cat") #label the columns with names
+results <- data.frame(matrix(ncol=6, nrow=0)) #create the table
+columns<-c("Medicaid_Category", "Time (months)", "Survival Percentage",  "LowerCI", "UpperCI", "Age_cat") # label the columns with names
 colnames(results)<-columns
@@
 # calculate survival at 60 mo
-results <- data.frame(matrix(ncol=5, nrow=0)) #create the table
-columns<-c("Medicaid_Category", "Survival Percentage",  "LowerCI", "UpperCI", "Age_cat") #label the columns with names
+results <- data.frame(matrix(ncol=6, nrow=0)) #create the table
+columns<-c("Medicaid_Category", "Time (months)","Survival Percentage",  "LowerCI", "UpperCI", "Age_cat") #label the columns with names
 colnames(results)<-columns
@@
 
-# SUPPLEMENTARY TABLE 4: make a function to populate a table with survival by race
+# SUPPLEMENTARY TABLE 4: Populate a table with survival by race at 60 mo
 ```{r}
@@
 
-# SUPPLEMENTARY TABLE 5: make a function to populate a table with survival by tumor type
+# SUPPLEMENTARY TABLE 5: Populate a table with survival by tumor type at 12, 24, 36, 48, and 60 mo
 ```{r}
@@
 # adjust for stage
-res.cox2a <- coxph(Surv(fumo, can_dead) ~ In_Medicaid_Flag + race_ethnicity + age_cat + Yost_ACS_2006_2010_quintile + stage + cluster(state), data = seer0to19) # Model adjusted for race age and a measure of poverty and stagee
+res.cox2a <- coxph(Surv(fumo, can_dead) ~ In_Medicaid_Flag + race_ethnicity + age_cat + Yost_ACS_2006_2010_quintile + stage + cluster(state), data = seer0to19) # Model adjusted for race age and a measure of poverty and stage
 res.cox2a

### 2023-09-27 Manuscript Analysis code 0 to 19.Rmd
Deleted some unecessary code from 60 month survival percentage calculations

@@
 
+# test code
+```{r}
+
+# create categories for AYA codes
+class(seer0to39$AYA_site_recode_WHO_2008) # numeric
+
+seer0to39 <- seer0to39 %>%
+  mutate(can_typeAYA = case_when(AYA_site_recode_WHO_2008 %in% c(1,2,3,4) ~ 0,
+                                 AYA_site_recode_WHO_2008 %in% c(5,6) ~ 1,
+                                 AYA_site_recode_WHO_2008 %in% c(7,8,9,10,11,12,13,14,15,16) ~ 2,
+                                 AYA_site_recode_WHO_2008 %in% c(17,18,19,20) ~ 3,
+                                 AYA_site_recode_WHO_2008 %in% c(21,22,23,24,25) ~ 4,
+                                 AYA_site_recode_WHO_2008 %in% c(26,27,28) ~ 5,
+                                 AYA_site_recode_WHO_2008 %in% c(29,30) ~ 6,
+                                 AYA_site_recode_WHO_2008 %in% c(31,32,33,34,35,36,37,38,39,40,41,42,43,44,45,46,47,48) ~ 7,
+                                 AYA_site_recode_WHO_2008 %in% c(49,50,51,52,53,54,55) ~ 8,
+                                 AYA_site_recode_WHO_2008 %in% c(55,56) ~ 9,
+                                 AYA_site_recode_WHO_2008 == 99 ~ 10))
+
+seer0to39$can_typeAYA <- factor(seer0to39$can_typeAYA, levels = c(0:10), labels = c(
+  "1. Leukemias",
+  "2. Lymphomas",
+  "3. CNS and Other Intracranial and Intraspinal Neoplasms (all behaviors)",
+  "4. Osseous & Chondromatous Neoplasms",
+  "5. Soft Tissue Sarcomas",
+  "6. Germ Cell and Trophoblastic Neoplasms",
+  "7. Melanoma and Skin Carcinomas",
+  "8. Carcinomas",
+  "9. Miscellaneous specified neoplasms, NOS",
+  "10. Unspecified Malignant Neoplasms",
+  "Unclassified"))
+table(seer0to39$can_typeAYA, useNA = "always")
+
+# categorize brain tumors according to AYA definition
+
+seer0to39 <- seer0to39 %>%
+  mutate(brain = case_when (can_typeAYA == "3. CNS and Other Intracranial and Intraspinal Neoplasms (all behaviors)"  & AYA_site_recode_WHO_2008 %in% c(7,8,9) ~ 0, # 3.1 Astrocytoma
+                            can_typeAYA == "3. CNS and Other Intracranial and Intraspinal Neoplasms (all behaviors)" & AYA_site_recode_WHO_2008 %in% c(10) ~ 1, # 3.2 Other glioma
+                            can_typeAYA == "3. CNS and Other Intracranial and Intraspinal Neoplasms (all behaviors)" & AYA_site_recode_WHO_2008 %in% c(11) ~ 2, # 3.3 Ependymoma
+                            can_typeAYA == "3. CNS and Other Intracranial and Intraspinal Neoplasms (all behaviors)" & AYA_site_recode_WHO_2008 %in% c(12,13) ~ 3, # 3.4 Medulloblastoma and other PNET
+                            can_typeAYA == "3.
+CNS and Other Intracranial and Intraspinal Neoplasms (all behaviors)" & AYA_site_recode_WHO_2008 %in% c(14) ~ 4, # 3.5 Other specified intracranial and instraspinal neoplasms
+                            can_typeAYA == "3.
+CNS and Other Intracranial and Intraspinal Neoplasms (all behaviors)" & AYA_site_recode_WHO_2008 %in% c(15,16) ~ 5)) # 3.6 Unspecified intracranial and intraspinal neoplasms
+seer0to39$brain
+table(seer0to39$brain, useNA = "always")
+
+seer0to39$brain <- factor(seer0to39$brain, levels = c(0:5), labels = c(
+  "3.1 Astrocytoma",
+  "3.2 Other glioma",
+  "3.3 Ependymoma",
+  "3.4 Medulloblastoma and other PNET",
+  "3.5 Other specified intracranial and instraspinal neoplasms",
+  "3.6 Unspecified intracranial and intraspinal neoplasms"))
+table(seer0to39$brain, useNA = "always")
+
+df <- seer0to39 %>%
+  filter(can_type == "III. CNS and miscellaneous intracranial and intraspinal neoplasms"|can_typeAYA == "3. CNS and Other Intracranial and Intraspinal Neoplasms (all behaviors)") %>%
+  group_by(can_type, can_typeAYA)%>%
+  tally()
+```
+
 # Exclusions
@@
   theme(legend.position='bottom')
-
 dev.off()
@@
 ```
-# SUPPLEMENTARY TABLE 2: Estimate survival time 0 to 19 at 160 mo
+# SUPPLEMENTARY TABLE 2: Estimate survival time 0 to 19 at 60 mo
 ```{r}
@@
 age <- unique(seer0to19$age_cat)
+
 # In_Medicaid_Flag
-results <- data.frame(matrix(ncol=6, nrow=0)) #create the table
-columns<-c("Medicaid_Category", "Time (months)", "Survival Percentage",  "LowerCI", "UpperCI", "Age_cat") # label the columns with names
+results <- data.frame(matrix(ncol=4, nrow=0)) #create the table
+columns<-c("Medicaid Category", "Survival Percentage",  "LowerCI", "UpperCI") # label the columns with names
 colnames(results)<-columns
@@
 # calculate survival at 60 mo
-results <- data.frame(matrix(ncol=6, nrow=0)) #create the table
-columns<-c("Medicaid_Category", "Time (months)","Survival Percentage",  "LowerCI", "UpperCI", "Age_cat") #label the columns with names
+results <- data.frame(matrix(ncol=4, nrow=0)) #create the table
+columns<-c("Medicaid_Category", "Survival Percentage",  "LowerCI", "UpperCI") # label the columns with names
 colnames(results)<-columns
@@
   results2 <- data.frame(matrix(ncol=4, nrow=0)) # create the table
-  columns<-c("Medicaid_Category", "Survival Percentage",  "LowerCI", "UpperCI") #label the columns with names
+  columns<-c("Medicaid_Category", "Survival Percentage",  "LowerCI", "UpperCI") # label the columns with names
   colnames(results2)<-columns
@@
 # calculate survival at 60 mo
-results <- data.frame(matrix(ncol=5, nrow=0)) # create the table
-columns<-c("Medicaid_Category", "Survival Percentage",  "LowerCI", "UpperCI", "Race_ethnicity") #label the columns with names
+results <- data.frame(matrix(ncol=4, nrow=0)) # create the table
+columns<-c("Medicaid_Category", "Survival Percentage",  "LowerCI", "UpperCI") #label the columns with names
 colnames(results)<-columns
@@
 results <- data.frame(matrix(ncol=5, nrow=0)) #create the table
-columns<-c("Medicaid_Category", "Survival Percentage",  "LowerCI", "UpperCI", "Race_ethnicity") #label the columns with names
+columns<-c("Medicaid_Category", "Survival Percentage",  "LowerCI", "UpperCI") #label the columns with names
 colnames(results)<-columns
@@
 # calculate survival at 60 mo
-results <- data.frame(matrix(ncol=5, nrow=0)) # create the table
-columns<-c("Medicaid_Category", "Survival Percentage",  "LowerCI", "UpperCI", "Cancer_type") #label the columns with names
+results <- data.frame(matrix(ncol=4, nrow=0)) # create the table
+columns<-c("Medicaid_Category", "Survival Percentage",  "LowerCI", "UpperCI") #label the columns with names
 colnames(results)<-columns
@@
 # calculate survival at 60 mo
-results <- data.frame(matrix(ncol=5, nrow=0)) # create the table
-columns<-c("Medicaid_Category", "Survival Percentage",  "LowerCI", "UpperCI", "Cancer_type") # label the columns with names
+results <- data.frame(matrix(ncol=4, nrow=0)) # create the table
+columns<-c("Medicaid_Category", "Survival Percentage",  "LowerCI", "UpperCI") # label the columns with names
 colnames(results)<-columns
@@
 SuppTable7c <- rbind(a,b,c,d,f,h,i,j,k, m,n,o,p,r,s,t,u,v,w)
+
 ```

### 2023-10-03 Manuscript Analysis code 0 to 19.Rmd
Updated using ghlt contrast furnction for time interval calculations

@@
 
-# Run models based on intervals for Medicaid_flag
-```{r}
-# first interval
-seer0to19first<-seer0to19.cp.format[which(seer0to19.cp.format$interval=="First"),]
-# adjusted models 0 to 19
-res.cox12 <- coxph(Surv(fumo, can_dead) ~ In_Medicaid_Flag + race_ethnicity +age_cat + 
-                     Yost_ACS_2006_2010_quintile + cluster(state), data = seer0to19first) 
-res.cox12
-summary(res.cox12)
-
-# second interval
-seer0to19second<-seer0to19.cp.format[which(seer0to19.cp.format$interval=="Second"),]
-# adjusted models 0 to 19
-res.cox24 <- coxph(Surv(fumo, can_dead) ~ In_Medicaid_Flag + race_ethnicity +age_cat + 
-                     Yost_ACS_2006_2010_quintile + cluster(state), data = seer0to19second) 
-res.cox24
-summary(res.cox24)
-
-# third interval
-seer0to19third<-seer0to19.cp.format[which(seer0to19.cp.format$interval=="Third"),]
-
-# adjusted models 0 to 19
-res.cox36 <- coxph(Surv(fumo, can_dead) ~ In_Medicaid_Flag + race_ethnicity +age_cat + 
-                     Yost_ACS_2006_2010_quintile + cluster(state), data = seer0to19third) 
-res.cox36
-summary(res.cox36)
-
-# fourth interval
-seer0to19fourth<-seer0to19.cp.format[which(seer0to19.cp.format$interval=="Fourth"),]
-# adjusted models 0 to 19
-res.cox36plus <- coxph(Surv(fumo, can_dead) ~ In_Medicaid_Flag + race_ethnicity +age_cat + 
-                     Yost_ACS_2006_2010_quintile + cluster(state), data = seer0to19fourth) 
-res.cox36plus
-summary(res.cox36plus)
+# Supplementary Table 6a: Run models based on intervals for Medicaid_flag
+```{r}
+age <- unique(seer0to19.cp.format$age_cat)
+
+# testing here
+m <-coxph(SurvObj ~ In_Medicaid_Flag + In_Medicaid_Flag*strata(interval) + race_ethnicity + age_cat + Yost_ACS_2006_2010_quintile + cluster(state), seer0to19.cp.format)
+
+m
+library(multcomp)
+X<- matrix(c(1,0,0,0,0,0,0,0,0,0,0,0,0,0,0), 1)
+
+t <- glht(m, linfct = X) 
+summary(t) # here are the results
+
+CI <- confint(t)
+CI <- as.data.frame(exp(CI$confint))
+# end testing
+
+W <- matrix(c(1,0,0,0,0,0,0,0,0,0,0,0,0,0,0), 1) 
+X <- matrix(c(1,0,0,0,0,0,0,0,0,0,0,0,1,0,0), 1)
+Y <- matrix(c(1,0,0,0,0,0,0,0,0,0,0,0,0,1,0), 1)
+Z <- matrix(c(1,0,0,0,0,0,0,0,0,0,0,0,0,0,1), 1)
+
+x <- coxph(SurvObj ~ In_Medicaid_Flag + In_Medicaid_Flag*strata(interval) + race_ethnicity + age_cat + Yost_ACS_2006_2010_quintile + cluster(state), data = seer0to19.cp.format) 
+  t0<- glht(x, linfct = W) 
+  df1<-confint(t0)
+  df1 <- as.data.frame(exp(df1$confint))
+
+  df1$interval = "0-<12" # enrolled first interval
+
+  t12<- glht(x, linfct = X) 
+  df2<-confint(t12)
+  df2 <- as.data.frame(exp(df2$confint))
+  df2$interval = "12-<24" # enrolled second interval
+
+  t24<- glht(x, linfct = Y) 
+  df3<-confint(t24)
+  df3 <- as.data.frame(exp(df3$confint))
+  df3$interval = "24-<36" # enrolled third interval
+
+  t36 <- glht(x, linfct = Z) 
+  df4 <-confint(t36)
+  df4 <- as.data.frame(exp(df4$confint))
+  df4$interval = "36+" # enrolled fourth interval
+
+  results <-rbind(df1, df2, df3, df4)  # this puts the results from each loop round in the results data file
+
+  results$Group <- "Overall"
+  results$category <- "Enrolled"
+SuppTable6a <- results
 
@@
 
-# SUPPLEMENTARY TABLE 6a and Graph hazard ratios and confidence intervals across time for Medicaid_flag variable
-```{r}
-# need to get HRs in a dataframe
-a<-as.data.frame(exp(cbind(HR = coef(res.cox12), confint(res.cox12))))
-   a<-a[c(1),] %>%
-    rownames_to_column() %>%
-    mutate(model= 0) %>%
-     mutate(Group = "Overall")
-
-b<-as.data.frame(exp(cbind(HR = coef(res.cox24), confint(res.cox24))))
-   b<-b[c(1),] %>%
-    rownames_to_column() %>%
-    mutate(model= 1) %>%
-     mutate(Group = "Overall")
-
-c<-as.data.frame(exp(cbind(HR = coef(res.cox36), confint(res.cox36))))
-   c<-c[c(1),] %>%
-    rownames_to_column() %>%
-    mutate(model= 2) %>%
-     mutate(Group = "Overall")
-
-d <-as.data.frame(exp(cbind(HR = coef(res.cox36plus), confint(res.cox36plus))))
-   d<-d[c(1),] %>%
-    rownames_to_column() %>%
-    mutate(model=3) %>%
-     mutate(Group = "Overall")
-
-SuppTable6a <- rbind(a,b,c,d)
-
-SuppTable6a$model <- factor(SuppTable6a$model, levels = c(0:3), labels = c("0 to 12 months", "12 to 24 months","24 to 36 months", ">36 months"))
-
-# graph HRs and 95% CIs
-SuppTable6a %>%
+# Graph hazard ratios and confidence intervals across time for enrolled variable
+```{r}
+p<- results %>%
   ggplot() +
-  geom_point(aes(y=HR, x=model)) + 
-  #scale_color_manual(values=c("black", "blue", "green", "red"), name="") + 
+  geom_point(aes(y=Estimate, x= interval), na.rm=TRUE, position=position_dodge(width=0.9), size =4) + 
   geom_hline(yintercept =1, linetype=2) +
   theme(axis.text.y=element_blank()) +
-  geom_errorbar(aes(y=HR, x=model, ymax = `2.5 %`, ymin = `97.5 %`), width = 0.1) +
+  geom_errorbar(aes(y=Estimate, x=interval, ymax = lwr, ymin = upr), na.rm=TRUE, position=position_dodge(width=0.9), width = 0.1) +
   theme_bw() +
@@
   coord_flip() 
-```
-
+p
+```
 # SUPPLEMENTARY TABLE 6b: Which cancer is driving this? 
 ```{r}
-# Split a survival data set at specified times to form a counting process format
-# For those with survival time of 0, add 0.5 mo because otherwise splitting doesn't work and it is more accurate
-
-cantype <- function(cantype, int, time){
-
-df <- survSplit(data  = seer0to19 %>% 
-                  filter(can_type == cantype),
-                            cut   = c(12, 24, 36),       # cut at time 12,24,36
-                            end   = "fumo_half",    # original survival time
-                            event = "can_dead",   # event indicator
-                            start = "start")    # will be created. zero by default
-
-# Recoding
-df <- within(df, {
-    # Create new survival object
-    SurvObj <- Surv(start, fumo_half, can_dead) 
-    # Create interval indicator
-    interval <- factor(start, levels = c(0,12, 24, 36), labels = c("First","Second", "Third", "Fourth"))
-})
-
-# Reordering
-df <- df[with(df, order(patient_id, fumo_half)),]  # sort by patient_id and then by fumo
-
-# run models based on intervals for Medicaid_flag
-# first interval
-df<-df[which(df$interval==int),]
-mod<-coxph(Surv(fumo, can_dead) ~ In_Medicaid_Flag + race_ethnicity +age_cat + Yost_ACS_2006_2010_quintile + cluster(state), data = df) 
-summary(mod)
-
-# need to get HRs in a dataframe
-a<-as.data.frame(exp(cbind(HR = coef(mod), confint(mod))))
-  a<-a[c(1),] %>%
-   rownames_to_column() %>%
-    mutate(model= time,
-          Group = cantype) 
+# In_Medicaid_Flag
+results <- data.frame(matrix(ncol=5, nrow=0)) # create empty table for results that will be populated by for loop
+columns<-c("estimate", "lwr", "upr", "age", "interval") # label the columns with names
+colnames(results) <- columns
+
+cantype <- c("I. Leukemias, myeloproliferative diseases, and myelodysplastic diseases", "III. CNS and miscellaneous intracranial and intraspinal neoplasms")
+colnames(seer0to19.cp.format)
+
+for (i in cantype) {
+  df <- seer0to19.cp.format %>% filter(can_type == i)
+
+  x <- coxph(SurvObj ~ In_Medicaid_Flag + In_Medicaid_Flag*strata(interval) + age_cat + race_ethnicity + Yost_ACS_2006_2010_quintile + cluster(state), data = df) 
+  t0<- glht(x, linfct = W) 
+  df1<-confint(t0)
+  df1 <- as.data.frame(exp(df1$confint))
+  df1$Group <- i
+  df1$interval = "0-<12" # enrolled first interval
+
+  t12<- glht(x, linfct = X) 
+  df2<-confint(t12)
+  df2 <- as.data.frame(exp(df2$confint))
+  df2$Group <- i
+  df2$interval = "12-<24" # enrolled second interval
+
+  t24<- glht(x, linfct = Y) 
+  df3<-confint(t24)
+  df3 <- as.data.frame(exp(df3$confint))
+  df3$Group <- i
+  df3$interval = "24-<36" # enrolled third interval
+
+  t36 <- glht(x, linfct = Z) 
+  df4 <-confint(t36)
+  df4 <- as.data.frame(exp(df4$confint))
+  df4$Group <- i
+  df4$interval = "36+" # enrolled fourth interval
+
+  results <-rbind(results, df1, df2, df3, df4)  # this puts the results from each loop round in the results data file
 }
 
-# Run top three cancer types
-a <-cantype("I. Leukemias, myeloproliferative diseases, and myelodysplastic diseases", "First", 0)
-b <-cantype("I. Leukemias, myeloproliferative diseases, and myelodysplastic diseases", "Second", 1)
-c <-cantype("I. Leukemias, myeloproliferative diseases, and myelodysplastic diseases", "Third", 2)
-d <-cantype("I. Leukemias, myeloproliferative diseases, and myelodysplastic diseases", "Fourth", 3)
-
-# Validate function for one cancer type
-test <- survSplit(data  = seer0to19 %>% 
-                  filter(can_type == "I. Leukemias, myeloproliferative diseases, and myelodysplastic diseases"),
-                            cut   = c(12, 24, 36),       # cut at time 12,24,36
-                            end   = "fumo_half",    # original survival time
-                            event = "can_dead",   # event indicator
-                            start = "start")    # will be created. zero by default
-
-# Recoding
-test<- within(test, {
-    # Create new survival object
-    SurvObj <- Surv(start, fumo_half, can_dead) 
-    # Create interval indicator
-    interval <- factor(start, levels = c(0,12, 24, 36), labels = c("First","Second", "Third", "Fourth"))
-})
-
-# Reordering
-test <- test[with(test, order(patient_id, fumo_half)),]  # sort by patient_id and then by fumo
-
-# run models based on intervals for Medicaid_flag
-# first interval
-test <- test[which(test$interval=="First"),]
-mod<-coxph(Surv(fumo, can_dead) ~ In_Medicaid_Flag + race_ethnicity +age_cat + Yost_ACS_2006_2010_quintile + cluster(state), data = test) 
-summary(mod)
-
-# need to get HRs in a dataframe
-a.2<-as.data.frame(exp(cbind(HR = coef(mod), confint(mod))))
-  a.2<-a.2[c(1),] %>%
-   rownames_to_column() %>%
-    mutate(model= 0,
-          Group = "I. Leukemias, myeloproliferative diseases, and myelodysplastic diseases")
-
-# Not presented
-# e <-cantype("II. Lymphomas and reticuloendothelial neoplasms", "First", 0) # converg. issues
-# f <-cantype("II. Lymphomas and reticuloendothelial neoplasms", "Second", 1) # converg. issues
-# g <-cantype("II. Lymphomas and reticuloendothelial neoplasms", "Third", 2) # converg. issues
-# h <-cantype("II. Lymphomas and reticuloendothelial neoplasms", "Fourth", 3)
-
-i <-cantype("III. CNS and miscellaneous intracranial and intraspinal neoplasms", "First", 0) 
-j <-cantype("III. CNS and miscellaneous intracranial and intraspinal neoplasms", "Second", 1)
-k <-cantype("III. CNS and miscellaneous intracranial and intraspinal neoplasms", "Third", 2) 
-l <-cantype("III. CNS and miscellaneous intracranial and intraspinal neoplasms", "Fourth", 3)
-
-SuppTable6b<-rbind(a,b,c,d,i,j,k,l)
-
-SuppTable6b$model <- factor(SuppTable6b$model, levels = c(0:3), labels = c("0 to 12 months", "12 to 24 months","24 to 36 months", ">36 months"))
-
-table(seer0to19$can_type)
+results$category = "Enrolled"
+SuppTable6b <- results
 ```
@@
   ggplot() +
-  geom_point(aes(y=HR, x=model)) + 
+  geom_point(aes(y=Estimate, x=interval)) + 
   geom_hline(yintercept =1, linetype=2) +
   theme(axis.text.y=element_blank()) +
-  geom_errorbar(aes(y=HR, x=model, ymax = `2.5 %`, ymin = `97.5 %`), width = 0.1) +
+  geom_errorbar(aes(y=Estimate, x=interval, ymax = lwr, ymin = upr), width = 0.1) +
   theme_bw() +
@@
 ```
-
 # run models based on intervals for mcont66
 ```{r}
-# first interval
-res.cox12m3 <- coxph(Surv(fumo, can_dead) ~ mcont66 + race_ethnicity +age_cat + 
-                     Yost_ACS_2006_2010_quintile + cluster(state), data = seer0to19first) 
-summary(res.cox12m3)
-
-# second interval
-seer0to19second<-seer0to19.cp.format[which(seer0to19.cp.format$interval=="Second"),]
-# adjusted models 0 to 19
-res.cox24m3 <- coxph(Surv(fumo, can_dead) ~ mcont66 + race_ethnicity +age_cat + 
-                     Yost_ACS_2006_2010_quintile + cluster(state), data = seer0to19second) 
-summary(res.cox24m3)
-
-# third interval
-seer0to19third<-seer0to19.cp.format[which(seer0to19.cp.format$interval=="Third"),]
-# adjusted models 0 to 19
-res.cox36m3 <- coxph(Surv(fumo, can_dead) ~ mcont66 + race_ethnicity +age_cat + 
-                     Yost_ACS_2006_2010_quintile + cluster(state), data = seer0to19third) 
-summary(res.cox36m3)
-
-# fourth interval
-seer0to19fourth<-seer0to19.cp.format[which(seer0to19.cp.format$interval=="Fourth"),]
-# adjusted models 0 to 19
-res.cox36plusm3 <- coxph(Surv(fumo, can_dead) ~ mcont66 + race_ethnicity +age_cat + 
-                     Yost_ACS_2006_2010_quintile + cluster(state), data = seer0to19fourth) 
-summary(res.cox36plusm3)
+# testing here
+m <-coxph(SurvObj ~ mcont66 + mcont66*strata(interval) + age_cat + race_ethnicity +  Yost_ACS_2006_2010_quintile + cluster(state), data = seer0to19.cp.format) 
+m
+
+X <- matrix(c(0,1,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0), 1)
+
+t <- glht(m, linfct = X) 
+summary(t) # here are the results
+
+CI<-confint(t)
+CI <- as.data.frame(exp(CI$confint))
+# end testing
+
+A <- matrix(c(1,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0), 1) # discontinuous 1st interval
+B <- matrix(c(0,1,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0), 1) # other 1st interval
+C <- matrix(c(0,0,1,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0), 1) # Not enrolled 1st interval
+
+D <- matrix(c(1,0,0,0,0,0,0,0,0,0,0,0,0,0,1,0,0,0,0,0,0,0,0), 1) # discontinuous 2nd interval
+E <- matrix(c(0,1,0,0,0,0,0,0,0,0,0,0,0,0,0,1,0,0,0,0,0,0,0), 1) # other 2nd interval
+F <- matrix(c(0,0,1,0,0,0,0,0,0,0,0,0,0,0,0,0,1,0,0,0,0,0,0), 1) # Not enrolled 2nd interval
+
+G <- matrix(c(1,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,1,0,0,0,0,0), 1) # discontinuous 3rd interval
+H <- matrix(c(0,1,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,1,0,0,0,0), 1) # other 3rd interval
+I <- matrix(c(0,0,1,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,1,0,0,0), 1) # Not enrolled 3rd interval
+
+J <- matrix(c(1,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,1,0,0), 1) # discontinuous 3rd interval
+K <- matrix(c(0,1,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,1,0), 1) # other 3rd interval
+L <- matrix(c(0,0,1,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,1), 1) # Not enrolled 3rd interval
+
+  x <- coxph(SurvObj ~ mcont66 + mcont66*strata(interval) + race_ethnicity + age_cat + Yost_ACS_2006_2010_quintile + cluster(state), data = seer0to19.cp.format) 
+  t0<- glht(x, linfct = A) 
+  df1<-confint(t0)
+  df1 <- as.data.frame(exp(df1$confint))
+  df1$Group <- "Overall"
+  df1$interval = "0-<12"
+  df1$category ="continuous" # first interval
+
+  t0<- glht(x, linfct = B) 
+  df2<-confint(t0)
+  df2 <- as.data.frame(exp(df2$confint))
+  df2$Group <- "Overall"
+  df2$interval = "0-<12"
+  df2$category ="discontinuous" # first interval
+
+  t0<- glht(x, linfct = C) 
+  df3<-confint(t0)
+  df3 <- as.data.frame(exp(df3$confint))
+  df3$Group <- "Overall"
+  df3$interval = "0-<12"
+  df3$category ="other" # first interval
+
+  t12<- glht(x, linfct = D) 
+  df4<-confint(t12)
+  df4 <- as.data.frame(exp(df4$confint))
+  df4$Group <- "Overall"
+  df4$interval = "12-<24"
+  df4$category ="continuous" # second interval
+
+  t12<- glht(x, linfct = E) 
+  df5<-confint(t12)
+  df5 <- as.data.frame(exp(df5$confint))
+  df5$Group <- "Overall"
+  df5$interval = "12-<24"
+  df5$category ="discontinuous" # second interval
+
+  t12<- glht(x, linfct = F) 
+  df6<-confint(t12)
+  df6 <- as.data.frame(exp(df6$confint))
+  df6$Group <- "Overall"
+  df6$interval = "12-<24"
+  df6$category ="other" # second interval
+
+  t24<- glht(x, linfct = G) 
+  df7<-confint(t24)
+  df7 <- as.data.frame(exp(df7$confint))
+  df7$Group <- "Overall"
+  df7$interval = "24-<36"
+  df7$category ="continuous" # third interval
+
+  t24<- glht(x, linfct = H) 
+  df8<-confint(t24)
+  df8<- as.data.frame(exp(df8$confint))
+  df8$Group <- "Overall"
+  df8$interval = "24-<36"
+  df8$category ="discontinuous" # third interval
+
+  t24<- glht(x, linfct = I) 
+  df9<-confint(t24)
+  df9 <- as.data.frame(exp(df9$confint))
+  df9$Group <- "Overall"
+  df9$interval = "24-<36"
+  df9$category ="other" # third interval
+
+  t36<- glht(x, linfct = J) 
+  df10<-confint(t36)
+  df10 <- as.data.frame(exp(df10$confint))
+  df10$Group <- "Overall"
+  df10$interval = "36+"
+  df10$category ="continuous" # fourth interval
+
+  t36<- glht(x, linfct = K) 
+  df11<-confint(t36)
+  df11<- as.data.frame(exp(df11$confint))
+  df11$Group <- "Overall"
+  df11$interval = "36+"
+  df11$category ="discontinuous" # fourth interval
+
+  t36<- glht(x, linfct = L) 
+  df12<-confint(t36)
+  df12 <- as.data.frame(exp(df12$confint))
+  df12$Group <- "Overall"
+  df12$interval = "36+"
+  df12$category ="other" # fourth interval
+
+  results <-rbind(df1, df2, df3, df4, df5, df6, df7, df8, df9, df10, df11, df12)  # this puts the results from each loop round in the results data file
+
+SuppTable6c <- results
 ```
@@
 ```{r}
-# need to get HRs in a dataframe
-a<-as.data.frame(exp(cbind(HR = coef(res.cox12m3), confint(res.cox12m3))))
-   a<-a[c(1,2,3),] %>%
-    rownames_to_column() %>%
-    mutate(model= 0) %>%
-    mutate(Group = "Overall")
-
-b<-as.data.frame(exp(cbind(HR = coef(res.cox24m3), confint(res.cox24m3))))
-   b<-b[c(1,2,3),] %>%
-    rownames_to_column() %>%
-    mutate(model= 1)  %>%
-    mutate(Group ="Overall")
-
-c<-as.data.frame(exp(cbind(HR = coef(res.cox36m3), confint(res.cox36m3))))
-   c<-c[c(1,2,3),] %>%
-    rownames_to_column() %>%
-    mutate(model= 2)  %>%
-    mutate(Group ="Overall")
-
-d <-as.data.frame(exp(cbind(HR = coef(res.cox36plusm3), confint(res.cox36plusm3))))
-   d<-d[c(1,2,3),] %>%
-    rownames_to_column() %>%
-    mutate(model=3)  %>%
-    mutate(Group ="Overall")
-
-SuppTable6c<- rbind(a,b,c,d)
-
-SuppTable6c$model <- factor(SuppTable6c$model, levels = c(0:3), labels = c("0 to 12 months", "12 to 24 months","24 to 36 months", ">36 months"))
-
 # graph HRs and 95% CIs
@@
   ggplot() +
-  geom_point(aes(y=HR, x=model)) + 
+  geom_point(aes(y=Estimate, x=interval, group = category, color = category),  position=position_dodge(width=0.9)) + 
   geom_hline(yintercept =1, linetype=2) +
   theme(axis.text.y=element_blank()) +
-  geom_errorbar(aes(y=HR, x=model, ymax = `2.5 %`, ymin = `97.5 %`), width = 0.1) +
+  geom_errorbar(aes(y=Estimate, x=interval, ymax = lwr, ymin = upr,  group = category, color = category), position=position_dodge(width=0.9), width = 0.1) +
+  geom_vline(xintercept =c(1.5, 2.5,3.5,4.5,5.5,6.5,7.5,8.5,9.5), linetype=2, color = "lightgray") +
   theme_bw() +
@@
   coord_flip() +
- facet_wrap(~rowname, nrow=1, ncol=3) +
-  labs(title = "Category", x = "time") + 
+  labs(y = "HR", x = "Interval") + 
   theme(strip.placement = "inside", legend.position = "bottom",  strip.text = element_text(size = 9, face="bold", color="black"), strip.background = element_rect(fill="lightgray"), plot.title = element_text(face="bold"), legend.text=element_text(size=9, face="bold"))
 ```
+
 # SUPPLEMENTARY TABLE 6: HRs by time interval
@@
 SuppTable6 <- rbind(SuppTable6a, SuppTable6b, SuppTable6c)
+
 ```
@@
 SuppTable6 <- SuppTable6 %>%
-  rename(`Time interval` = model,
-         Category = rowname,
-         `lower 95% CI` = `2.5 %`,
-         `upper 95% CI` = `97.5 %`) %>%
-  mutate(Category = case_when(Category == "In_Medicaid_FlagEnrolled" ~ "Enrolled",
-            Category == "mcont66Discontinuous" ~ "Discontinuous",
-            Category == "mcont66Other" ~ "Other",
-            Category == "mcont66Continuous" ~ "Continuous")) %>%
-
+  rename(`Time interval` = interval,
+         Category = category,
+         `lower 95% CI` = lwr,
+         `upper 95% CI` = upr) %>%
+  mutate(Category = case_when(Category == "Enrolled"~ "Enrolled",
+            Category == "discontinuous" ~ "Discontinuous",
+            Category == "other" ~ "Other",
+            Category == "continuous" ~ "Continuous")) %>%
   mutate_if(is.numeric, round, 2)
 
-SuppTable6$`HR (95% CI)` <-paste0(SuppTable6$HR, " (", SuppTable6$`lower 95% CI`, " to ", SuppTable6$`upper 95% CI`, ")") 
+SuppTable6$`HR (95% CI)` <-paste0(SuppTable6$Estimate, " (", SuppTable6$`lower 95% CI`, " to ", SuppTable6$`upper 95% CI`, ")") 
 SuppTable6 <- SuppTable6 %>%
-  dplyr::select(1,5,6,7)
+  dplyr::select(4,5,6,7)
 
@@
                "Supp Table 7 (EM)" = SuppTable7)
-write.xlsx(sheets, file="Supplementary tables.xlsx", showNA=TRUE, overwrite = TRUE)
+write.xlsx(sheets, file="Supplementary tables updated.xlsx", showNA=TRUE, overwrite = TRUE)
 ```

### 2023-10-04 Manuscript Analysis code 0 to 19.Rmd
edited some code that resulted in incorrect Supplementary Table 2 output to excel

@@
 SuppTable2 <- SuppTable2 %>%
-  dplyr::select(1,5)
+  dplyr::select(1,6)
 

### 2023-12-04 Manuscript Analysis code 0 to 19.Rmd
adding revision changes

@@
 "Behavior_code_ICD_O_2", "Histologic_Type_ICD_O_3", "Behavior_code_ICD_O_3",
-"Grade", "Diagnostic_Confirmation", "Type_of_Reporting_Source", 
-"Derived_AJCC_T_6th_ed_2004_2015", 
-"Derived_AJCC_N_6th_ed_2004_2015",
+"Grade", "Diagnostic_Confirmation", "Type_of_Reporting_Source", "Derived_AJCC_T_6th_ed_2004_2015", "Derived_AJCC_N_6th_ed_2004_2015",
 "Derived_AJCC_M_6th_ed_2004_2015", 
@@
 seer0to39$race_cat<-factor(seer0to39$race_cat, levels=c(0:4), labels=c("White", "Black", "AIAN", "API", "Unknown"))
+seer0to39$race_cat2<-factor(seer0to39$race_cat2, levels=c(0:2), labels=c("Non-Hispanic White", "Non-Hispanic Black", "Non-Hispanic Other"))
 seer0to39$hispanic<-factor(seer0to39$hispanic, levels=c(0:1), labels=c("Non-Hispanic", "Hispanic"))
@@
 
-seer0to39$stage <- factor(seer0to39$stage, levels = c(0:2), labels = c("In situ/localized", "Regional", "Distant"))
+seer0to39$stage <- factor(seer0to39$stage, levels = c(0:2), labels = c("in situ/localized", "Regional", "Distant"))
 
@@
 
-```
-
-# test code
-```{r}
-
-# create categories for AYA codes
-class(seer0to39$AYA_site_recode_WHO_2008) # numeric
-
-seer0to39 <- seer0to39 %>%
-  mutate(can_typeAYA = case_when(AYA_site_recode_WHO_2008 %in% c(1,2,3,4) ~ 0,
-                                 AYA_site_recode_WHO_2008 %in% c(5,6) ~ 1,
-                                 AYA_site_recode_WHO_2008 %in% c(7,8,9,10,11,12,13,14,15,16) ~ 2,
-                                 AYA_site_recode_WHO_2008 %in% c(17,18,19,20) ~ 3,
-                                 AYA_site_recode_WHO_2008 %in% c(21,22,23,24,25) ~ 4,
-                                 AYA_site_recode_WHO_2008 %in% c(26,27,28) ~ 5,
-                                 AYA_site_recode_WHO_2008 %in% c(29,30) ~ 6,
-                                 AYA_site_recode_WHO_2008 %in% c(31,32,33,34,35,36,37,38,39,40,41,42,43,44,45,46,47,48) ~ 7,
-                                 AYA_site_recode_WHO_2008 %in% c(49,50,51,52,53,54,55) ~ 8,
-                                 AYA_site_recode_WHO_2008 %in% c(55,56) ~ 9,
-                                 AYA_site_recode_WHO_2008 == 99 ~ 10))
-
-seer0to39$can_typeAYA <- factor(seer0to39$can_typeAYA, levels = c(0:10), labels = c(
-  "1. Leukemias",
-  "2. Lymphomas",
-  "3. CNS and Other Intracranial and Intraspinal Neoplasms (all behaviors)",
-  "4. Osseous & Chondromatous Neoplasms",
-  "5. Soft Tissue Sarcomas",
-  "6. Germ Cell and Trophoblastic Neoplasms",
-  "7. Melanoma and Skin Carcinomas",
-  "8. Carcinomas",
-  "9. Miscellaneous specified neoplasms, NOS",
-  "10. Unspecified Malignant Neoplasms",
-  "Unclassified"))
-table(seer0to39$can_typeAYA, useNA = "always")
-
-# categorize brain tumors according to AYA definition
-
-seer0to39 <- seer0to39 %>%
-  mutate(brain = case_when (can_typeAYA == "3. CNS and Other Intracranial and Intraspinal Neoplasms (all behaviors)"  & AYA_site_recode_WHO_2008 %in% c(7,8,9) ~ 0, # 3.1 Astrocytoma
-                            can_typeAYA == "3. CNS and Other Intracranial and Intraspinal Neoplasms (all behaviors)" & AYA_site_recode_WHO_2008 %in% c(10) ~ 1, # 3.2 Other glioma
-                            can_typeAYA == "3. CNS and Other Intracranial and Intraspinal Neoplasms (all behaviors)" & AYA_site_recode_WHO_2008 %in% c(11) ~ 2, # 3.3 Ependymoma
-                            can_typeAYA == "3. CNS and Other Intracranial and Intraspinal Neoplasms (all behaviors)" & AYA_site_recode_WHO_2008 %in% c(12,13) ~ 3, # 3.4 Medulloblastoma and other PNET
-                            can_typeAYA == "3.
-CNS and Other Intracranial and Intraspinal Neoplasms (all behaviors)" & AYA_site_recode_WHO_2008 %in% c(14) ~ 4, # 3.5 Other specified intracranial and instraspinal neoplasms
-                            can_typeAYA == "3.
-CNS and Other Intracranial and Intraspinal Neoplasms (all behaviors)" & AYA_site_recode_WHO_2008 %in% c(15,16) ~ 5)) # 3.6 Unspecified intracranial and intraspinal neoplasms
-seer0to39$brain
-table(seer0to39$brain, useNA = "always")
-
-seer0to39$brain <- factor(seer0to39$brain, levels = c(0:5), labels = c(
-  "3.1 Astrocytoma",
-  "3.2 Other glioma",
-  "3.3 Ependymoma",
-  "3.4 Medulloblastoma and other PNET",
-  "3.5 Other specified intracranial and instraspinal neoplasms",
-  "3.6 Unspecified intracranial and intraspinal neoplasms"))
-table(seer0to39$brain, useNA = "always")
-
-df <- seer0to39 %>%
-  filter(can_type == "III. CNS and miscellaneous intracranial and intraspinal neoplasms"|can_typeAYA == "3. CNS and Other Intracranial and Intraspinal Neoplasms (all behaviors)") %>%
-  group_by(can_type, can_typeAYA)%>%
-  tally()
 ```
@@
 
-# Exclude those diagnosed before 2006 (a handful)
+# Exclude those diagnosed before 2006 
 ```{r}
@@
                              mcontb == 1 ~ 2, # discontinuous
-                             mcontb == 2 ~ 3)) # Other
+                             mcontb == 2 ~ 3)) %>% # Other
+  mutate(mcont66b = case_when(In_Medicaid_Flag == 0 ~ 0, # not linked
+                             mcontb2 == 0 ~ 1, # continuous
+                             mcontb2 == 1 ~ 2, # discontinuous
+                             mcontb2 == 2 ~ 3, # Other
+                             mcontb2 == 3 ~ 4)) # at or after diagnosis
+
 table(seer0to19$mcont66, useNA = "always")
-
 table(seer0to19$mcont66, seer0to19$In_Medicaid_Flag, useNA = "always")
+
+table(seer0to19$mcont66b, useNA = "always")
+table(seer0to19$mcont66b, seer0to19$In_Medicaid_Flag, useNA = "always")
 
@@
 seer0to19$mcont66 <- factor(seer0to19$mcont66, levels = c(0:3), labels = c("Not enrolled", "Continuous", "Discontinuous", "Other"))
+seer0to19$mcont66b <- factor(seer0to19$mcont66b, levels = c(0:4), labels = c("Not enrolled", "Continuous", "Discontinuous", "Other", "At diagnosis"))
+
+table(seer0to19$mcont66b)
 ```
@@
 
-table1(~age_cat + sex  + race_ethnicity + Yost_ACS_2006_2010_quintile + In_Medicaid_Flag  + stage + fumo  + can_type + can_dead2|mcont66, droplevels = TRUE, seer0to19 %>% filter(Agerecodewithsingleages_and_100 !=18 ))
-
-table1(~mcont66|age_cat, droplevels = TRUE, seer0to19%>% filter(Agerecodewithsingleages_and_100 <18 ))
+table1(~can_type|In_Medicaid_Flag, seer0to19)
+
+table1(~age_cat + sex  + race_ethnicity + Yost_ACS_2006_2010_quintile + In_Medicaid_Flag  + stage + fumo  + can_type + can_dead2|mcont66b, droplevels = TRUE, seer0to19)
+```
+
+# FOR REVISION (insurance distribution)
+```{r}
+# classify primary payor variable
+class(seer0to19$Primary_Payer_at_DX)
+seer0to19$Primary_Payer_at_DX <- as.numeric(seer0to19$Primary_Payer_at_DX)
+
+seer0to19 <- seer0to19 %>%
+  mutate(payer = case_when(Primary_Payer_at_DX %in% c(1,2) ~ 0,
+                           Primary_Payer_at_DX %in% c(10,20,21) ~ 1,
+                           Primary_Payer_at_DX %in% c(31,35) ~ 2,
+                           Primary_Payer_at_DX %in% c(60,64) ~ 3,
+                           Primary_Payer_at_DX %in% c(65,66,67,68) ~ 4,
+                           Primary_Payer_at_DX == 99 ~ 5))
+
+seer0to19$payer <- factor(seer0to19$payer, levels = c(0:5), labels = c("Uninsured", "Private and Insurance NOS", "Medicaid", "Medicare", "Other", "Unknown"))
+
+seer0to19$Insurance_Recode_2007 <- as.numeric(seer0to19$Insurance_Recode_2007)
+table(seer0to19$Insurance_Recode_2007, useNA = "always")
+seer0to19 <- seer0to19 %>%
+  mutate(insurance = case_when(Insurance_Recode_2007 == 1 ~ 0,
+                               Insurance_Recode_2007 == 2 ~ 1,
+                               Insurance_Recode_2007 == 3 ~ 2,
+                               Insurance_Recode_2007 == 4 ~ 3,
+                               Insurance_Recode_2007 == 5 ~ 4))
+
+label(seer0to19$insurance) <- "Insurance at diagnosis or initial treatment"
+
+seer0to19$insurance <- factor(seer0to19$insurance, levels = c(0:4), labels = c("Uninsured", "Any Medicaid", "Insured", "Insured/No specifics", "Insurance status unknown"))
+
+# Not shown in paper
+table1(~age_cat + sex  + race_ethnicity + Yost_ACS_2006_2010_quintile + In_Medicaid_Flag  + stage + fumo  + can_type + can_dead2 + insurance|mcont66, droplevels = TRUE, seer0to19)
+
+table1(~insurance|mcont66, droplevels = TRUE, seer0to19)
 ```
@@
   remove_labels()
-
+```
+
+# FOR REVISION QUESTION (why not use AJCC staging vs. summary stage)--check missing data by cancer type for both
+```{r}
+colnames(seer0to19_stage)
+
+table1(~ as.factor(SEERCombinedSummaryStage2000200) + as.factor(DerivedAJCCStageGroup6thed20042), seer0to19_stage)
+```
+
+```{r}
 # exclude those with missing stage and redefine some variables so the function doesn't throw an error.
@@
   select(stage, In_Medicaid_Flag, SEER_registry, mcont66, race_cat2, age_cat, Yost_ACS_2006_2010_quintile, state, can_type) %>%
-  mutate(stage2 = case_when(stage == "In situ/localized" ~ "0: In situ/localized",
+  mutate(stage2 = case_when(stage == "in situ/localized" ~ "0: In situ/localized",
                             stage == "Regional" ~ "1: Regional",
@@
 
-# relabel columns and dropt intercept rows
+# relabel columns and drop intercept rows
 names(df4)[1:3]<-c("OR", "lowCI", "highCI")
@@
 df4<-df4[-c(1,3),]
-df4 <- tibble::rownames_to_column(df4, "Category") # Apply rownames_to_column
+df4 <- tibble::rownames_to_column(df4, "Exposure comparison") # Apply rownames_to_column
 
 # Replace multiple strings at a time
-rep_str = c('1: Regional:In_Medicaid_FlagEnrolled'='Regional','2: Distant:In_Medicaid_FlagEnrolled'='Distant')
-rep_str2 = c('1: Regional:In_Medicaid_FlagEnrolled'='Enrolled','2: Distant:In_Medicaid_FlagEnrolled'='Enrolled')
-df4$Stage<- str_replace_all(df4$Category, rep_str)
-df4$Category<- str_replace_all(df4$Category, rep_str2) # this is the dataframe for summary results
+rep_str = c('1: Regional:In_Medicaid_FlagEnrolled'='Regional vs. in situ/localized','2: Distant:In_Medicaid_FlagEnrolled'='Distant vs. in situ/localized')
+rep_str2 = c('1: Regional:In_Medicaid_FlagEnrolled'='Enrolled vs. Not enrolled','2: Distant:In_Medicaid_FlagEnrolled'='Enrolled vs. Not enrolled')
+df4$Stage<- str_replace_all(df4$`Exposure comparison`, rep_str)
+df4$`Exposure comparison`<- str_replace_all(df4$`Exposure comparison`, rep_str2) # this is the dataframe for summary results
 
@@
 df8<-df8[-c(1,5),]
-df8 <- tibble::rownames_to_column(df8, "Category") # Apply rownames_to_column
+df8 <- tibble::rownames_to_column(df8, "Exposure comparison") # Apply rownames_to_column
 df8
@@
 # Replace multiple strings at a time
-rep_str = c('1: Regional:mcont66Continuous'='Regional','1: Regional:mcont66Discontinuous'='Regional','1: Regional:mcont66Other'='Regional', '2: Distant:mcont66Continuous'='Distant','2: Distant:mcont66Discontinuous'='Distant','2: Distant:mcont66Other'='Distant')
-df8$Stage<- str_replace_all(df8$Category, rep_str)
+rep_str = c('1: Regional:mcont66Continuous'='Regional vs.
+in situ/localized','1: Regional:mcont66Discontinuous'='Regional.
+vs.
+in situ/localized','1: Regional:mcont66Other'='Regional vs.
+in situ/localized', '2: Distant:mcont66Continuous'='Distant vs.
+in situ/localized','2: Distant:mcont66Discontinuous'='Distant vs.
+in situ/localized','2: Distant:mcont66Other'='Distant vs.
+in situ/localized')
+df8$Stage<- str_replace_all(df8$`Exposure comparison`, rep_str)
 df8
-rep_str2 = c('1: Regional:mcont66Continuous'='Continuous','1: Regional:mcont66Discontinuous'='Discontinuous','1: Regional:mcont66Other'='Other', '2: Distant:mcont66Continuous'='Continuous','2: Distant:mcont66Discontinuous'='Discontinuous','2: Distant:mcont66Other'='Other')
-df8$Category<- str_replace_all(df8$Category, rep_str2)
+
+rep_str2 = c('1: Regional:mcont66Continuous'='Continuous vs.
+Not enrolled','1: Regional:mcont66Discontinuous'='Discontinuous vs.
+Not enrolled','1: Regional:mcont66Other'='Other vs.
+Not enrolled', '2: Distant:mcont66Continuous'='Continuous vs.
+Not enrolled','2: Distant:mcont66Discontinuous'='Discontinuous vs.
+Not enrolled','2: Distant:mcont66Other'='Other vs.
+Not enrolled')
+
+df8$`Exposure comparison`<- str_replace_all(df8$`Exposure comparison`, rep_str2)
 df8 # this is the dataframe for summary results
@@
 ```{r}
-model1a <- multinom(stage2 ~ In_Medicaid_Flag + race_cat2 + age_cat + Yost_ACS_2006_2010_quintile , data=seer0to19_stage)
+model1a <- multinom(stage2 ~ In_Medicaid_Flag + race_cat2 + age_cat + Yost_ACS_2006_2010_quintile + can_type, data=seer0to19_stage)
 summary(model1a)
-nrow(fitted(model1a)) # 21,507 is number of obs
+nrow(fitted(model1a)) # 21,502 is number of obs
 
@@
 
-df12<-df12[-c(1,3:11, 13:20),]
+df12<-df12[-c(1, 3:23, 25:44),]
 df12 <- tibble::rownames_to_column(df12, "Category") # Apply rownames_to_column
@@
 # adjusted model
-model2a <- multinom(stage2 ~mcont66 + race_cat2 + age_cat + Yost_ACS_2006_2010_quintile, data=seer0to19_stage)
+model2a <- multinom(stage2 ~mcont66 + race_cat2 + age_cat + Yost_ACS_2006_2010_quintile + can_type, data=seer0to19_stage)
 summary(model2a)
@@
 
-df16<-df16[-c(1,5:13, 17:24),]
+df16<-df16[-c(1, 5:25, 29:48),]
 df16 <- tibble::rownames_to_column(df16, "Category") # Apply rownames_to_column
@@
 rep_str2 = c('1: Regional:mcont66Continuous'='Continuous','1: Regional:mcont66Discontinuous'='Discontinuous','1: Regional:mcont66Other'='Other', '2: Distant:mcont66Continuous'='Continuous','2: Distant:mcont66Discontinuous'='Discontinuous','2: Distant:mcont66Other'='Other')
+
 df16$Category<- str_replace_all(df16$Category, rep_str2)
@@
 # adjusted model
-model2a <- multinom(stage2 ~ mcont66.2 + race_cat2 + age_cat + Yost_ACS_2006_2010_quintile, data=seer0to19_stage)
+model2a <- multinom(stage2 ~ mcont66.2 + race_cat2 + age_cat + Yost_ACS_2006_2010_quintile + can_type, data=seer0to19_stage)
 summary(model2a)
@@
 
-df20<-df20[-c(1,5:13, 17:24),]
+df20<-df20[-c(1, 5:25, 29:48),]
 df20 <- tibble::rownames_to_column(df20, "Category") # Apply rownames_to_column
@@
 names(df2a)[1:3]<-c("OR", "cluster.lowCI", "cluster.highCI")
-df2a<-df2a[-c(1,3:11, 13:20),]
-df2a <- tibble::rownames_to_column(df2a, "Category") # Apply rownames_to_column
+
+df2a<-df2a[-c(1, 3:12, 14:22),]
+df2a <- tibble::rownames_to_column(df2a, "Category comparison") # Apply rownames_to_column
 df2a  
@@
 names(df4a)[1:3]<-c("OR", "cluster.lowCI", "cluster.highCI")
-df4a <-df4a[-c(1,5:13, 17:24),]
-df4a <- tibble::rownames_to_column(df4a, "Category") # Apply rownames_to_column
+
+df4a <-df4a[-c(1, 5:14, 18:26),]
+df4a <- tibble::rownames_to_column(df4a, "Category comparison") # Apply rownames_to_column
 df4a  
@@
 # Replace multiple strings at a time
-rep_str = c('1: Regional:mcont66Continuous'='Regional','1: Regional:mcont66Discontinuous'='Regional','1: Regional:mcont66Other'='Regional', '2: Distant:mcont66Continuous'='Distant','2: Distant:mcont66Discontinuous'='Distant','2: Distant:mcont66Other'='Distant', '1: Regional:In_Medicaid_FlagEnrolled'='Regional','2: Distant:In_Medicaid_FlagEnrolled'='Distant')
-SuppTable1$Stage<- str_replace_all(SuppTable1$Category, rep_str)
-
-rep_str2 = c('1: Regional:mcont66Continuous'='Continuous','1: Regional:mcont66Discontinuous'='Discontinuous','1: Regional:mcont66Other'='Other', '2: Distant:mcont66Continuous'='Continuous','2: Distant:mcont66Discontinuous'='Discontinuous','2: Distant:mcont66Other'='Other', '1: Regional:In_Medicaid_FlagEnrolled'='Enrolled','2: Distant:In_Medicaid_FlagEnrolled'='Enrolled')
-SuppTable1$Category<- str_replace_all(SuppTable1$Category, rep_str2)
-
-SuppTable1
+rep_str = c('1: Regional:mcont66Continuous'='Regional vs.
+in situ/localized','1: Regional:mcont66Discontinuous'='Regional vs.
+in situ/localized','1: Regional:mcont66Other'='Regional vs.
+in situ/localized', '2: Distant:mcont66Continuous'='Distant vs.
+in situ/localized','2: Distant:mcont66Discontinuous'='Distant vs.
+in situ/localized','2: Distant:mcont66Other'='Distant vs.
+in situ/localized', '1: Regional:In_Medicaid_FlagEnrolled'='Regional vs.
+in situ/localized','2: Distant:In_Medicaid_FlagEnrolled'='Distant vs.
+in situ/localized')
+SuppTable1$`Stage outcome comparison`<- str_replace_all(SuppTable1$`Category comparison`, rep_str)
+
+rep_str2 = c('1: Regional:mcont66Continuous'='Continuous vs.
+Not enrolled','1: Regional:mcont66Discontinuous'='Discontinuous vs.
+Not enrolled','1: Regional:mcont66Other'='Other vs.
+Not enrolled', '2: Distant:mcont66Continuous'='Continuous vs.
+Not enrolled','2: Distant:mcont66Discontinuous'='Discontinuous vs.
+Not enrolled','2: Distant:mcont66Other'='Other vs.
+Not enrolled', '1: Regional:In_Medicaid_FlagEnrolled'='Enrolled vs.
+Not enrolled','2: Distant:In_Medicaid_FlagEnrolled'='Enrolled vs.
+Not enrolled')
+SuppTable1$`Category comparison`<- str_replace_all(SuppTable1$`Category comparison`, rep_str2)
+
+names(SuppTable1)[1]<-c("Exposure comparison")
+
 ```
@@
 Table2$CancerType<-"Overall"
+
+# Replace multiple strings at a time
+rep_str = c('Regional'='Regional vs. in situ/localized','Distant'='Distant vs. in situ/localized')
+Table2$Stage<- str_replace_all(Table2$Stage, rep_str)
+
+rep_str2 = c('Enrolled'='Enrolled vs. Not enrolled','Continuous'='Continuous vs. Not enrolled', 'Discontinuous'='Discontinuous vs. Not enrolled', 'Other'='Other vs. Not enrolled')
+Table2$Category<- str_replace_all(Table2$Category, rep_str2)
+
+names(Table2)[1]<-c("Exposure comparison")
+names(Table2)[5]<-c("Stage outcome comparison")
+
+# Reorder columns using the pipe operator
+Table2 <- Table2 %>% select("Exposure comparison" ,  "OR", "cluster.lowCI","cluster.highCI", "CancerType", "Stage outcome comparison")
 
@@
 SuppTable1a <- SuppTable1 %>%
-  filter(CancerType != "V. Retinoblastoma")
+  filter(`CancerType` != "V. Retinoblastoma")
 
@@
 
-SuppTable1a$Stage <- factor(SuppTable1a$Stage, levels = c("Regional", "Distant"))
+SuppTable1a$`Stage outcome comparison` <- factor(SuppTable1a$`Stage outcome comparison`, levels = c("Regional vs. in situ/localized", "Distant vs. in situ/localized"))
 
@@
 Plot1 <- SuppTable1a %>%
-  filter(Category == "Enrolled", Stage == "Regional") %>%
+  filter(`Exposure comparison` == "Enrolled vs. Not enrolled", `Stage outcome comparison` == "Regional vs. in situ/localized") %>%
   ggplot() +
-  geom_point(aes(y=OR, x = Abbrev, group = Category, color = Category)) + 
+  geom_point(aes(y=OR, x = Abbrev, group = `Exposure comparison`, color = `Exposure comparison`), color = "black") + 
   geom_hline(yintercept =1, linetype=2) +
   theme(axis.text.y=element_blank()) +
-  geom_errorbar(aes(y=OR, x=Abbrev, group = Category, color = Category, ymax = cluster.highCI, ymin = cluster.lowCI), width = 0.1) +
+  geom_errorbar(aes(y=OR, x=Abbrev, group = `Exposure comparison`, color = `Exposure comparison`, ymax = cluster.highCI, ymin = cluster.lowCI), width = 0.1,  color = "black") +
   theme_bw() +
@@
   coord_flip() +
-  labs(color = NULL, title = "A", x = "Cancer type") +
-  ylim(0,2) +
+  labs(color = NULL, title = "A. Regional vs. in situ/localized", x = "Cancer type") +
+  ylim(0,3.5) +
   theme(strip.placement = "inside", legend.position = "bottom",  strip.text = element_text(size = 9, face="bold", color="black"), strip.background = element_rect(fill="lightgray"), plot.title = element_text(face="bold"), legend.text=element_text(size=9, face="bold")) +
@@
 Plot2 <- SuppTable1a %>%
-  filter(Category == "Enrolled", Stage == "Distant") %>%
+  filter(`Exposure comparison` == "Enrolled vs. Not enrolled", `Stage outcome comparison` == "Distant vs. in situ/localized") %>%
   ggplot() +
-  geom_point(aes(y=OR, x = Abbrev, group = Category, color = Category)) + 
+  geom_point(aes(y=OR, x = Abbrev, group = `Exposure comparison`, color = `Exposure comparison`), color = "black") + 
   geom_hline(yintercept =1, linetype=2) +
   theme(axis.text.y=element_blank()) +
-  geom_errorbar(aes(y=OR, x=Abbrev, group = Category, color = Category, ymax = cluster.highCI, ymin = cluster.lowCI), width = 0.1) +
+  geom_errorbar(aes(y=OR, x=Abbrev, group = `Exposure comparison`, color = `Exposure comparison`, ymax = cluster.highCI, ymin = cluster.lowCI), width = 0.1, color = "black") +
   theme_bw() +
@@
   coord_flip() +
-  labs(color = NULL, title = "B", x = "Cancer type") +
+  labs(color = NULL, title = "B. Distant vs. in situ/localized", x = "Cancer type") +
   ylim(0,3.5) +
@@
 # graph ORs and 95% CIs
-SuppTable1a$Category <- factor(SuppTable1a$Category, levels = c("Enrolled", "Discontinuous", "Other", "Continuous"))
+SuppTable1a$`Exposure comparison` <- factor(SuppTable1a$`Exposure comparison`, levels = c("Enrolled vs.
+Not enrolled", "Discontinuous vs.
+Not enrolled", "Other vs.
+Not enrolled", "Continuous vs.
+Not enrolled"))
 
 Plot3 <- SuppTable1a %>%
-  filter(Category != "Enrolled", Stage == "Regional") %>%
+  filter(`Exposure comparison` != "Enrolled vs. Not enrolled", `Stage outcome comparison` == "Regional vs. in situ/localized") %>%
   ggplot() +
-  geom_point(aes(y=OR, x = Abbrev, group = Category, color = Category), na.rm=TRUE, position=position_dodge(width=0.9)) + 
+  geom_point(aes(y=OR, x = Abbrev, group = `Exposure comparison`, color = `Exposure comparison`), na.rm=TRUE, position=position_dodge(width=0.9)) + 
   geom_hline(yintercept =1, linetype=2) +
@@
   theme(axis.text.y=element_blank()) +
-  geom_errorbar(position = position_dodge(width=0.9), aes(y=OR, x=Abbrev, group = Category, color = Category,  ymax = cluster.highCI, ymin = cluster.lowCI),  width = 0.1) +
+  geom_errorbar(position = position_dodge(width=0.9), aes(y=OR, x=Abbrev, group = `Exposure comparison`, color = `Exposure comparison`,  ymax = cluster.highCI, ymin = cluster.lowCI),  width = 0.1) +
+  theme_bw() +
+  theme(axis.line = element_line(colour = "black"),
+    panel.grid.major = element_blank(),
+    panel.grid.minor = element_blank(),
+    panel.border = element_blank(),
+    panel.background = element_blank(),
+    axis.text.x=element_text(size=11, face = "bold", color = "black"),
+    axis.text.y=element_text(size=11, face = "bold", color = "black"),
+    axis.title.x=element_text(size=12, face = "bold", color = "black"),
+    axis.title.y=element_text(size=12, face = "bold", color = "black"))+
+  coord_flip() +
+  labs(color = NULL, title = "C. Regional vs. in situ/localized", x = "Cancer type") +
+  ylim(0,4.0) +
+  theme(strip.placement = "inside", legend.position = "bottom",  strip.text = element_text(size = 9, face="bold", color="black"), strip.background = element_rect(fill="lightgray"), plot.title = element_text(face="bold"), legend.text=element_text(size=9, face="bold")) +
+  scale_color_manual(breaks = c("Discontinuous vs. Not enrolled", "Continuous vs. Not enrolled", "Other vs. Not enrolled"),
+                        values=c("red", "cyan3", "blue"))
+
+Plot3
+table(SuppTable1a$`Exposure comparison`)
+
+Plot4 <- SuppTable1a %>%
+  filter(`Exposure comparison` != "Enrolled vs. Not enrolled", `Stage outcome comparison` == "Distant vs. in situ/localized") %>%
+  ggplot() +
+  geom_point(aes(y=OR, x = Abbrev, group = `Exposure comparison`, color = `Exposure comparison`), na.rm=TRUE, position=position_dodge(width=0.9)) + 
+  geom_hline(yintercept =1, linetype=2) +
+  geom_vline(xintercept =c(1.5, 2.5,3.5,4.5,5.5,6.5,7.5,8.5,9.5), linetype=2, color = "lightgray") +
+  theme(axis.text.y=element_blank()) +
+  geom_errorbar(position = position_dodge(width=0.9), aes(y=OR, x=Abbrev, group = `Exposure comparison`, color = `Exposure comparison`,  max = cluster.highCI, ymin = cluster.lowCI),  width = 0.1) +
   theme_bw() +
@@
  # facet_grid(cols = vars(y.level)) +
-  labs(color = NULL, title = "C", x = "Cancer type") +
-  ylim(0,4.0) +
-  theme(strip.placement = "inside", legend.position = "bottom",  strip.text = element_text(size = 9, face="bold", color="black"), strip.background = element_rect(fill="lightgray"), plot.title = element_text(face="bold"), legend.text=element_text(size=9, face="bold")) +
-  scale_color_manual(breaks = c("Discontinuous", "Continuous", "Other"),
-                        values=c("red", "cyan3", "blue"))
-
-Plot3
-
-Plot4 <- SuppTable1a %>%
-  filter(Category != "Enrolled", Stage == "Distant") %>%
-  ggplot() +
-  geom_point(aes(y=OR, x = Abbrev, group = Category, color = Category), na.rm=TRUE, position=position_dodge(width=0.9)) + 
-  geom_hline(yintercept =1, linetype=2) +
-  geom_vline(xintercept =c(1.5, 2.5,3.5,4.5,5.5,6.5,7.5,8.5,9.5), linetype=2, color = "lightgray") +
-  theme(axis.text.y=element_blank()) +
-  geom_errorbar(position = position_dodge(width=0.9), aes(y=OR, x=Abbrev, group = Category, color = Category,  max = cluster.highCI, ymin = cluster.lowCI),  width = 0.1) +
-  theme_bw() +
-  theme(axis.line = element_line(colour = "black"),
-    panel.grid.major = element_blank(),
-    panel.grid.minor = element_blank(),
-    panel.border = element_blank(),
-    panel.background = element_blank(),
-    axis.text.x=element_text(size=11, face = "bold", color = "black"),
-    axis.text.y=element_text(size=11, face = "bold", color = "black"),
-    axis.title.x=element_text(size=12, face = "bold", color = "black"),
-    axis.title.y=element_text(size=12, face = "bold", color = "black"))+
-  coord_flip() +
- # facet_grid(cols = vars(y.level)) +
-  labs(color = NULL, title = "D", x = "Cancer type") +
+  labs(color = NULL, title = "D. Distant vs. in situ/localized", x = "Cancer type") +
   ylim(0,5.5) +
   theme(strip.placement = "inside", legend.position = "bottom",  strip.text = element_text(size = 9, face="bold", color="black"), strip.background = element_rect(fill="lightgray"), plot.title = element_text(face="bold"), legend.text=element_text(size=9, face="bold")) +
-  scale_color_manual(breaks = c("Discontinuous", "Continuous", "Other"),
+  scale_color_manual(breaks = c("Discontinuous vs. Not enrolled", "Continuous vs. Not enrolled", "Other vs. Not enrolled"),
                         values=c("red", "cyan3", "blue"))
@@
 ```
+# Survival plots with discontinuous divided to at diagnosis and the rest
+```{r}
+# mcont66b
+ plot3 <- ggsurvplot(fit = survfit(Surv(fumo, can_dead) ~ mcont66b, data = seer0to19 ), 
+    xlab = "Months from diagnosis", 
+    ylab = "Survival probability", 
+    fun="pct", 
+    title = "B",
+    font.title = c(12, "bold"),
+    legend.title = "",
+    legend = "bottom",
+    font.legend=c(12, "bold"),
+    legend.labs = c("Not enrolled", "Continuous", "Discontinuous", "Other", "At diagnosis"),
+    palette = c("red", "cyan3", "blue", "black", "purple"),
+    font.x = c(12, "bold"),
+    font.y = c(12, "bold"),
+    xlim=c(0, 60),
+    break.x.by = 12,
+    ylim=c(50, 100),
+    censor.size=0.5,
+    font.tickslab = c(12, "bold"),
+    pval=TRUE,
+    pval.size = 5,
+    pval.coord = c(2,55),
+    conf.int = FALSE)
+
+plot3
+
+pdf("Figure 3.pdf", width=8.5, height=11, paper="USr")
+plot1$plot + plot2$plot + plot_layout(ncol=1, nrow=2, guides='keep') &
+  theme(legend.position='bottom')
+
+dev.off()
+```
+
 # SUPPLEMENTARY TABLE 2: Estimate survival time 0 to 19 at 60 mo
@@
 # Overall
-summary(survfit(Surv(fumo, can_dead) ~ 1, data = seer0to19), times=60) # estimates 60 month survival time
+overall <- summary(survfit(Surv(fumo, can_dead) ~ 1, data = seer0to19), times=60) # estimates 60 month survival time
+
+overall <- as.data.frame(cbind(overall$time, round(overall$surv*100, 2), round(overall$lower*100, 2), round(overall$upper*100,2))) # makes dataframe
+overall$Category <- "Overall"
+overall <- setNames(overall, c("Time (months)", "Survival Percentage", "Lower CI", "Upper CI", "Category")) # gives column names
+overall <- overall %>% select("Category","Time (months)", "Survival Percentage", "Lower CI", "Upper CI")
 
@@
 # Make Supplementary table 2
-SuppTable2 <- rbind(SuppTable2a, SuppTable2b)
+SuppTable2 <- rbind(overall, SuppTable2a, SuppTable2b)
 SuppTable2 <- unique(SuppTable2)
@@
 
-# SUPPLEMENTARY TABLE 5: Populate a table with survival by tumor type at 12, 24, 36, 48, and 60 mo
+# SUPPLEMENTARY TABLE 5: Populate a table with survival by tumor type at 60 mo
 ```{r}
@@
 # mcont66 adjusted models 0 to 19
-res.cox4 <- coxph(Surv(fumo, can_dead) ~ mcont66 + race_ethnicity +age_cat + Yost_ACS_2006_2010_quintile + cluster(state), data = seer0to19) # Model adjusted for race age and a measure of poverty
+res.cox4 <- coxph(Surv(fumo, can_dead) ~ mcont66 + race_ethnicity + age_cat + Yost_ACS_2006_2010_quintile + cluster(state), data = seer0to19) # Model adjusted for race age and a measure of poverty
 res.cox4
@@
     mutate(age_group = "0 to 19 years")
+
 Table3 <- rbind(a,b,c,d)  
@@
 res.cox4.2
+
 summary(res.cox4.2)
@@
 
-Medflag.surv <- survfit(Surv(fumo, can_dead) ~ In_Medicaid_Flag + race_ethnicity +age_cat + Yost_ACS_2006_2010_quintile + cluster(state), data = seer0to19) # get survival probabilities
+Medflag.surv <- survfit(Surv(fumo, can_dead) ~ In_Medicaid_Flag + race_ethnicity + age_cat + Yost_ACS_2006_2010_quintile + cluster(state), data = seer0to19) # get survival probabilities
 
@@
 # For those with survival time of 0, add 0.5 mo because otherwise splitting doesn't work and it is more accurate
+
+# the if_elsee function does not like labels
+seer0to19 <- seer0to19 %>% 
+  remove_labels()
 
@@
     # Create interval indicator
-    interval <- factor(start, levels = c(0,12, 24, 36), labels = c("First","Second", "Third", "Fourth"))
+    interval <- factor(start, levels = c(0, 12, 24, 36), labels = c("First","Second", "Third", "Fourth"))
+    SurvObj2 <- Surv((fumo_half - start), can_dead)
 })
@@
 m <-coxph(SurvObj ~ In_Medicaid_Flag + In_Medicaid_Flag*strata(interval) + race_ethnicity + age_cat + Yost_ACS_2006_2010_quintile + cluster(state), seer0to19.cp.format)
-
 m
+
+# try another way
+m2 <- coxph(SurvObj2 ~ In_Medicaid_Flag + In_Medicaid_Flag*strata(interval) + race_ethnicity + age_cat + Yost_ACS_2006_2010_quintile + cluster(state), seer0to19.cp.format)
+
+m
+m2
+
 library(multcomp)
@@
 
-  df1$interval = "0-<12" # enrolled first interval
+  df1$interval = "0 to ≥12" # enrolled first interval
 
@@
   df2 <- as.data.frame(exp(df2$confint))
-  df2$interval = "12-<24" # enrolled second interval
+  df2$interval = ">12 to ≥24" # enrolled second interval
 
@@
   df3 <- as.data.frame(exp(df3$confint))
-  df3$interval = "24-<36" # enrolled third interval
+  df3$interval = ">24 to ≥36" # enrolled third interval
 
   t36 <- glht(x, linfct = Z) 
-  df4 <-confint(t36)
-  df4 <- as.data.frame(exp(df4$confint))
-  df4$interval = "36+" # enrolled fourth interval
-
-  results <-rbind(df1, df2, df3, df4)  # this puts the results from each loop round in the results data file
+  df4.0 <-confint(t36)
+  df4.0 <- as.data.frame(exp(df4.0$confint))
+  df4.0$interval = "≥36" # enrolled fourth interval
+
+  results <-rbind(df1, df2, df3, df4.0)  # this puts the results from each loop round in the results data file
 
@@
   df1$Group <- i
-  df1$interval = "0-<12" # enrolled first interval
+  df1$interval = "0-≥12" # enrolled first interval
 
@@
   df2$Group <- i
-  df2$interval = "12-<24" # enrolled second interval
+  df2$interval = ">12-≥24" # enrolled second interval
 
@@
   df3$Group <- i
-  df3$interval = "24-<36" # enrolled third interval
+  df3$interval = ">24-≥36" # enrolled third interval
 
   t36 <- glht(x, linfct = Z) 
-  df4 <-confint(t36)
-  df4 <- as.data.frame(exp(df4$confint))
-  df4$Group <- i
-  df4$interval = "36+" # enrolled fourth interval
-
-  results <-rbind(results, df1, df2, df3, df4)  # this puts the results from each loop round in the results data file
+  df4.1 <-confint(t36)
+  df4.1 <- as.data.frame(exp(df4.1$confint))
+  df4.1$Group <- i
+  df4.1$interval = "≥36" # enrolled fourth interval
+
+  results <-rbind(results, df1, df2, df3, df4.1)  # this puts the results from each loop round in the results data file
 }
@@
   df1$Group <- "Overall"
-  df1$interval = "0-<12"
+  df1$interval = "0 to ≥12"
   df1$category ="continuous" # first interval
@@
   df2$Group <- "Overall"
-  df2$interval = "0-<12"
+  df2$interval = "0 to ≥12"
   df2$category ="discontinuous" # first interval
@@
   df3$Group <- "Overall"
-  df3$interval = "0-<12"
+  df3$interval = "0 to ≥12"
   df3$category ="other" # first interval
@@
   t12<- glht(x, linfct = D) 
-  df4<-confint(t12)
-  df4 <- as.data.frame(exp(df4$confint))
-  df4$Group <- "Overall"
-  df4$interval = "12-<24"
-  df4$category ="continuous" # second interval
+  df4.2<-confint(t12)
+  df4.2 <- as.data.frame(exp(df4.2$confint))
+  df4.2$Group <- "Overall"
+  df4.2$interval = ">12 to ≥24"
+  df4.2$category ="continuous" # second interval
 
@@
   df5$Group <- "Overall"
-  df5$interval = "12-<24"
+  df5$interval = ">12 to ≥24"
   df5$category ="discontinuous" # second interval
@@
   df6$Group <- "Overall"
-  df6$interval = "12-<24"
+  df6$interval = ">12 to ≥24"
   df6$category ="other" # second interval
@@
   df7$Group <- "Overall"
-  df7$interval = "24-<36"
+  df7$interval = ">24 to ≥36"
   df7$category ="continuous" # third interval
@@
   t24<- glht(x, linfct = H) 
-  df8<-confint(t24)
-  df8<- as.data.frame(exp(df8$confint))
-  df8$Group <- "Overall"
-  df8$interval = "24-<36"
-  df8$category ="discontinuous" # third interval
+  df8.2<-confint(t24)
+  df8.2<- as.data.frame(exp(df8.2$confint))
+  df8.2$Group <- "Overall"
+  df8.2$interval = ">24 to ≥36"
+  df8.2$category ="discontinuous" # third interval
 
@@
   df9$Group <- "Overall"
-  df9$interval = "24-<36"
+  df9$interval = ">24 to ≥36"
   df9$category ="other" # third interval
@@
   df10$Group <- "Overall"
-  df10$interval = "36+"
+  df10$interval = "≥36"
   df10$category ="continuous" # fourth interval
@@
   df11$Group <- "Overall"
-  df11$interval = "36+"
+  df11$interval = "≥36"
   df11$category ="discontinuous" # fourth interval
@@
   df12$Group <- "Overall"
-  df12$interval = "36+"
+  df12$interval = "≥36"
   df12$category ="other" # fourth interval
 
-  results <-rbind(df1, df2, df3, df4, df5, df6, df7, df8, df9, df10, df11, df12)  # this puts the results from each loop round in the results data file
+  results <-rbind(df1, df2, df3, df4.2, df5, df6, df7, df8.2, df9, df10, df11, df12)  # this puts the results from each loop round in the results data file
 
@@
 SuppTable6 <- rbind(SuppTable6a, SuppTable6b, SuppTable6c)
-
 ```
@@
 SuppTable7<- rbind(SuppTable7a, SuppTable7b, SuppTable7c)
+
 table(SuppTable7$rowname)
@@
   rename(Category = rowname) %>%
-  mutate(Category = case_when(Category == "varContinuous" ~ "Continuous",
-            Category == "varDiscontinuous" ~ "Discontinuous",
-            Category == "varOther" ~ "Other",
-            Category == "varEnrolled" ~ "Enrolled")) %>%
+  mutate(Category = case_when(Category == "varContinuous" ~ "Continuous vs. Not enrolled",
+            Category == "varDiscontinuous" ~ "Discontinuous vs. Not enrolled",
+            Category == "varOther" ~ "Other vs. Not enrolled",
+            Category == "varEnrolled" ~ "Enrolled vs. Not enrolled")) %>%
   mutate(Group2 = case_when(Group %in% c("0-4", "5-9", "10-14", "15-19") ~ "Age",
@@
 plot1 <- SuppTable7 %>%
-  filter(Category == "Enrolled", Group2 == "Age") %>%
+  filter(Category == "Enrolled vs. Not enrolled", Group2 == "Age") %>%
 ggplot() +
-  geom_point(aes(y=HR, x = Abbrev, group=Category, color = Category)) + 
+  geom_point(aes(y=HR, x = Abbrev, group=Category, color = Category), colour = "black") + 
   geom_hline(yintercept =1, linetype=2) +
   theme(axis.text.y=element_blank()) +
-  geom_errorbar(aes(y=HR, x=Abbrev, group = Category, color = Category, ymax = `2.5 %`, ymin = `97.5 %`), width = 0.1) +
+  geom_errorbar(aes(y=HR, x=Abbrev, group = Category, color = Category, ymax = `2.5 %`, ymin = `97.5 %`), width = 0.1, colour = "black") +
   theme_bw() +
@@
   theme(strip.placement = "inside", legend.position = "bottom",  strip.text = element_text(size = 9, face="bold", color="black"), strip.background = element_rect(fill="lightgray"), plot.title = element_text(face="bold"), legend.text=element_text(size=9, face="bold")) +
-          ggtitle("A") + theme(plot.title = element_text(face="bold")) +
-  scale_color_manual(breaks = c("Enrolled"), values=c("black")) +
+          ggtitle("A. Age group") + theme(plot.title = element_text(face="bold")) +
+  scale_color_manual(breaks = c("Enrolled vs. Not enrolled"), values=c("black")) +
   theme(legend.title=element_blank())
@@
 plot2 <-SuppTable7 %>%
-  filter(Category == "Enrolled", Group2 == "Race") %>% 
+  filter(Category == "Enrolled vs. Not enrolled", Group2 == "Race") %>% 
   ggplot() +
-  geom_point(aes(y=HR, x = Abbrev, group = Category, color = Category)) + 
+  geom_point(aes(y=HR, x = Abbrev, group = Category, color = Category), colour = "black") + 
   geom_hline(yintercept =1, linetype=2) +
   theme(axis.text.y=element_blank()) +
-  geom_errorbar(aes(y=HR, x=Abbrev, group = Category, color = Category, ymax = `2.5 %`, ymin = `97.5 %`), width = 0.1) +
+  geom_errorbar(aes(y=HR, x=Abbrev, group = Category, color = Category, ymax = `2.5 %`, ymin = `97.5 %`), width = 0.1, colour = "black") +
   theme_bw() +
@@
   ylim(0.5,5) +
-  theme(strip.placement = "inside", legend.position = "bottom",  strip.text = element_text(size = 9, face="bold", color="black"), strip.background = element_rect(fill="lightgray"), plot.title = element_text(face="bold"), legend.text=element_text(size=9, face="bold")) + ggtitle("C") +theme(plot.title = element_text(face="bold")) +
-  scale_color_manual(breaks = c("Enrolled"),
+  theme(strip.placement = "inside", legend.position = "bottom",  strip.text = element_text(size = 9, face="bold", color="black"), strip.background = element_rect(fill="lightgray"), plot.title = element_text(face="bold"), legend.text=element_text(size=9, face="bold")) + ggtitle("C.
+Race/Ethnicity") +theme(plot.title = element_text(face="bold")) +
+  scale_color_manual(breaks = c("Enrolled vs. Not enrolled"),
                         values=c("black")) +
@@
 plot3 <-SuppTable7 %>%
-  filter(Category == "Enrolled", Group2 == "CancerType", Abbrev != "OTHUN") %>%
+  filter(Category == "Enrolled vs. Not enrolled", Group2 == "CancerType", Abbrev != "OTHUN") %>%
   ggplot() +
-  geom_point(aes(y=HR, x = Abbrev, group = Category, color = Category)) + 
+  geom_point(aes(y=HR, x = Abbrev, group = Category, color = Category), colour = "black") + 
   geom_hline(yintercept =1, linetype=2) +
   theme(axis.text.y=element_blank()) +
-  geom_errorbar(aes(y=HR, x=Abbrev, group = Category, color = Category, ymax = `2.5 %`, ymin = `97.5 %`), width = 0.1) +
+  geom_errorbar(aes(y=HR, x=Abbrev, group = Category, color = Category, ymax = `2.5 %`, ymin = `97.5 %`), width = 0.1, colour = "black") +
   theme_bw() +
@@
   ylim(0.5,3.6) +
-  theme(strip.placement = "inside", legend.position = "bottom",  strip.text = element_text(size = 9, face="bold", color="black"), strip.background = element_rect(fill="lightgray"), plot.title = element_text(face="bold"), legend.text=element_text(size=9, face="bold")) + ggtitle("E") +theme(plot.title = element_text(face="bold")) +
-  scale_color_manual(breaks = c("Enrolled"),
+  theme(strip.placement = "inside", legend.position = "bottom",  strip.text = element_text(size = 9, face="bold", color="black"), strip.background = element_rect(fill="lightgray"), plot.title = element_text(face="bold"), legend.text=element_text(size=9, face="bold")) + ggtitle("E.
+Cancer type") +theme(plot.title = element_text(face="bold")) +
+  scale_color_manual(breaks = c("Enrolled vs. Not enrolled"),
                         values=c("black")) +
@@
 ```{r}
-SuppTable7$Category <- factor(SuppTable7$Category, levels = c("Enrolled", "Discontinuous", "Other",  "Continuous"))
+SuppTable7$Category <- factor(SuppTable7$Category, levels = c("Enrolled vs. Not enrolled", "Discontinuous vs. Not enrolled", "Other vs. Not enrolled",  "Continuous vs. Not enrolled"))
 
 plot4 <- SuppTable7 %>%
-  filter(Category != "Enrolled", Group2 == "Age") %>%
+  filter(Category != "Enrolled vs. Not enrolled", Group2 == "Age") %>%
 ggplot() +
@@
   ylim(0.5,3) +
-  theme(strip.placement = "inside", legend.position = "bottom",  strip.text = element_text(size = 9, face="bold", color="black"), strip.background = element_rect(fill="lightgray"), plot.title = element_text(face="bold"), legend.text=element_text(size=9, face="bold")) + ggtitle("B") +theme(plot.title = element_text(face="bold")) +
-  scale_color_manual(breaks = c("Discontinuous", "Continuous", "Other"),
-                        values=c("red", "cyan3", "blue")) +
+  theme(strip.placement = "inside", legend.position = "bottom",  strip.text = element_text(size = 9, face="bold", color="black"), strip.background = element_rect(fill="lightgray"), plot.title = element_text(face="bold"), legend.text=element_text(size=9, face="bold")) + ggtitle("B.
+Age group") +theme(plot.title = element_text(face="bold")) +
+  scale_color_manual(breaks = c("Discontinuous vs. Not enrolled", "Continuous vs. Not enrolled", "Other vs. Not enrolled"), values=c("red", "cyan3", "blue")) +
   theme(legend.title=element_blank())
@@
 plot5 <-SuppTable7 %>%
-  filter(Category != "Enrolled", Group2 == "Race") %>% 
+  filter(Category != "Enrolled vs. Not enrolled", Group2 == "Race") %>% 
   ggplot() +
@@
   ylim(0.5,7) +
-  theme(strip.placement = "inside", legend.position = "bottom",  strip.text = element_text(size = 9, face="bold", color="black"), strip.background = element_rect(fill="lightgray"), plot.title = element_text(face="bold"), legend.text=element_text(size=9, face="bold")) + ggtitle("D") +theme(plot.title = element_text(face="bold")) +
-  scale_color_manual(breaks = c("Discontinuous", "Continuous", "Other"),
+  theme(strip.placement = "inside", legend.position = "bottom",  strip.text = element_text(size = 9, face="bold", color="black"), strip.background = element_rect(fill="lightgray"), plot.title = element_text(face="bold"), legend.text=element_text(size=9, face="bold")) + ggtitle("D.
+Race/Ethnicity") +theme(plot.title = element_text(face="bold")) +
+  scale_color_manual(breaks = c("Discontinuous vs. Not enrolled", "Continuous vs. Not enrolled", "Other vs. Not enrolled"),
                         values=c("red", "cyan3", "blue")) +
@@
 plot6 <-SuppTable7 %>%
-  filter(Category != "Enrolled", Group2 == "CancerType", !Abbrev %in% c("RB", "OTHUN")) %>%
+  filter(Category != "Enrolled vs. Not enrolled", Group2 == "CancerType", !Abbrev %in% c("RB", "OTHUN")) %>%
   ggplot() +
@@
   ylim(0.5,4.5) +
-  theme(strip.placement = "inside", legend.position = "bottom",  strip.text = element_text(size = 9, face="bold", color="black"), strip.background = element_rect(fill="lightgray"), plot.title = element_text(face="bold"), legend.text=element_text(size=9, face="bold")) + ggtitle("F") +theme(plot.title = element_text(face="bold"))+
-  scale_color_manual(breaks = c("Discontinuous", "Continuous", "Other"),
+  theme(strip.placement = "inside", legend.position = "bottom",  strip.text = element_text(size = 9, face="bold", color="black"), strip.background = element_rect(fill="lightgray"), plot.title = element_text(face="bold"), legend.text=element_text(size=9, face="bold")) + ggtitle("F.
+Cancer type") +theme(plot.title = element_text(face="bold"))+
+  scale_color_manual(breaks = c("Discontinuous vs. Not enrolled", "Continuous vs. Not enrolled", "Other vs. Not enrolled"),
                         values=c("red", "cyan3", "blue"))+
@@
 ```{r}
-Table2 <- Table2%>%
+Table2 <- Table2 %>%
   rename(`lower 95% CI` = cluster.lowCI,
@@
          "upper 95% CI" = "97.5 %") %>%
-  mutate(Category = case_when(Category == "In_Medicaid_FlagEnrolled" ~ "Enrolled",
-            Category == "mcont66Discontinuous" ~ "Discontinuous",
-            Category == "mcont66Other" ~ "Other",
-            Category == "mcont66Continuous" ~ "Continuous"))  %>%
+  mutate(`Exposure comparison` = case_when(Category == "In_Medicaid_FlagEnrolled" ~ "Enrolled vs. Not enrolled",
+            Category == "mcont66Discontinuous" ~ "Discontinuous vs. Not enrolled",
+            Category == "mcont66Other" ~ "Other vs. Not enrolled",
+            Category == "mcont66Continuous" ~ "Continuous vs. Not enrolled"))  %>%
   mutate_if(is.numeric, round, 2)
@@
          `upper 95% CI` = upr) %>%
-  mutate(Category = case_when(Category == "Enrolled"~ "Enrolled",
-            Category == "discontinuous" ~ "Discontinuous",
-            Category == "other" ~ "Other",
-            Category == "continuous" ~ "Continuous")) %>%
+  mutate(`Exposure comparison`= case_when(Category == "Enrolled"~ "Enrolled vs. Not enrolled",
+            Category == "discontinuous" ~ "Discontinuous vs. Not enrolled",
+            Category == "other" ~ "Other vs. Not enrolled",
+            Category == "continuous" ~ "Continuous vs. Not enrolled")) %>%
   mutate_if(is.numeric, round, 2)
@@
 SuppTable6 <- SuppTable6 %>%
-  dplyr::select(4,5,6,7)
+  dplyr::select(4,5,7,8)
 
@@
   dplyr::select(1,6,7)
+
+SuppTable7 <- SuppTable7%>%
+  rename(`Exposure comparison` = Category)
 ```
@@
 sheets <- list("Supp Table 1 (stage cantype)" = SuppTable1,
-               "Supp Table 2 (surv overall)" = SuppTable2,
+               "Supp Table 2 (surv prob)" = SuppTable2,
                "Supp Table 3 (surv prob age )" = SuppTable3,
@@
 # regional vs. in situ/localized
-model1_logit<-glm(stage ~ In_Medicaid_Flag, data = seer0to19_stage[which(seer0to19_stage$stage %in% c("In situ/localized", "Regional")),], family = binomial(link = "logit"))
+model1_logit<-glm(stage ~ In_Medicaid_Flag, data = seer0to19_stage[which(seer0to19_stage$stage %in% c("in situ/localized", "Regional")),], family = binomial(link = "logit"))
 odds.n.ends(model1_logit) # these are CIs without robust SEs
@@
 # distant vs. in situ/localized
-model1_logit2<-glm(stage ~ In_Medicaid_Flag, data = seer0to19_stage[which(seer0to19_stage$stage %in% c("In situ/localized", "Distant")),], family = binomial(link = "logit"))
+model1_logit2<-glm(stage ~ In_Medicaid_Flag, data = seer0to19_stage[which(seer0to19_stage$stage %in% c("in situ/localized", "Distant")),], family = binomial(link = "logit"))
 odds.n.ends(model1_logit2)  # these are CIs without robust SEs
 
-model1_logit3 <-glm(stage ~ mcont66, data = seer0to19_stage[which(seer0to19_stage$stage %in% c("In situ/localized", "Distant")),], family = binomial(link = "logit"))
+model1_logit3 <-glm(stage ~ mcont66, data = seer0to19_stage[which(seer0to19_stage$stage %in% c("in situ/localized", "Distant")),], family = binomial(link = "logit"))
 odds.n.ends(model1_logit3)  # these are CIs without robust SEs
 
-model1_logit4 <-glm(stage ~ mcont66, data = seer0to19_stage[which(seer0to19_stage$stage %in% c("In situ/localized", "Distant")),], family = binomial(link = "logit"))
+model1_logit4 <-glm(stage ~ mcont66, data = seer0to19_stage[which(seer0to19_stage$stage %in% c("in situ/localized", "Distant")),], family = binomial(link = "logit"))
 odds.n.ends(model1_logit4)  # these are CIs without robust SEs
@@
 
-model2_logit_RSE <- miceadds::glm.cluster(data=seer0to19_stage[which(seer0to19_stage$stage %in% c("In situ/localized", "Distant")),], formula= stage ~ In_Medicaid_Flag + race_cat2 + age_cat + Yost_ACS_2006_2010_quintile, cluster="state", family="binomial")
+model2_logit_RSE <- miceadds::glm.cluster(data=seer0to19_stage[which(seer0to19_stage$stage %in% c("in situ/localized", "Distant")),], formula= stage ~ In_Medicaid_Flag + race_cat2 + age_cat + Yost_ACS_2006_2010_quintile, cluster="state", family="binomial")
 summary(model2_logit_RSE)

### 2023-12-15 Manuscript Analysis code 0 to 19.Rmd
adding revision changes

@@
                                      race_cat == 1 ~ 1, # Non-Hispanic Black
-                                     race_cat %in% c(2,3,4,5) ~ 2)) %>% # Non-Hispanic Other
+                                     race_cat %in% c(2,3,4) ~ 2)) %>% # Non-Hispanic Other
 
@@
                                 SEERcausespecificdeathclassific == 1 ~1)) # Dead from cancer
-# note 8 an 9 are missing or N/A
+# note 8 an 9 are dead missing COD and N/A
 
@@
 seer0to39$race_ethnicity<-factor(seer0to39$race_ethnicity, levels=c(0:4), labels=c("Non-Hispanic White", "Non-Hispanic Black", "Non-Hispanic API", "Non-Hispanic Other",  "Hispanic")) 
-seer0to39$sex <- factor(seer0to39$sex, levels = c(0:2), labels = c("Male", "Female", "Not Stated"))
+seer0to39$sex <- factor(seer0to39$sex, levels = c(0:2), labels = c("Male", "Female", "Unknown"))
 
@@
 # calculate follow-up time from month of dx, recode year of dx TO Month_of_last_follow_up_recode, Year_of_last_follow_up_recode.
-The difference between these variable dates would provide months of follow-up
+The difference between these variable dates provides months of follow-up
 summary(seer0to39$Month_of_diagnosis_recode)
@@
 summary(as.numeric(seer0to39$Month_of_last_follow_up_recode))
-seer0to39$Month_of_last_follow_up_recode<-as.numeric(seer0to39$Month_of_last_follow_up_recode) # make variable numeric
-seer0to39$SEER_DateofDeath_Month<-as.numeric(seer0to39$SEER_DateofDeath_Month) # make variable numeric
+seer0to39$Month_of_last_follow_up_recode <- as.numeric(seer0to39$Month_of_last_follow_up_recode) # make variable numeric
+seer0to39$SEER_DateofDeath_Month <- as.numeric(seer0to39$SEER_DateofDeath_Month) # make variable numeric
 summary(seer0to39$Month_of_last_follow_up_recode)
@@
 # assign a date of diagnosis using 15 as the day because there is no day of dx so assume mean is 15
-seer0to39$proxy_day_of_dx <-15
-seer0to39$proxy_day_of_fu <-15
-seer0to39$proxy_day_of_dth <-15
+seer0to39$proxy_day_of_dx <- 15
+seer0to39$proxy_day_of_fu <- 15
+seer0to39$proxy_day_of_dth <- 15
 
@@
 
-seer0to19b <- seer0to19
+seer0to19b <- seer0to19 # using to not write over seer0to19
 
@@
                              mcontb == 2 ~ 3)) %>% # Other
+  # not used for this paper
   mutate(mcont66b = case_when(In_Medicaid_Flag == 0 ~ 0, # not linked
@@
 seer0to19$mcont66 <- factor(seer0to19$mcont66, levels = c(0:3), labels = c("Not enrolled", "Continuous", "Discontinuous", "Other"))
+
 seer0to19$mcont66b <- factor(seer0to19$mcont66b, levels = c(0:4), labels = c("Not enrolled", "Continuous", "Discontinuous", "Other", "At diagnosis"))
@@
 
+# used in paper
 table1(~age_cat + sex  + race_ethnicity + Yost_ACS_2006_2010_quintile + In_Medicaid_Flag  + stage + fumo  + can_type + can_dead2|mcont66, droplevels = TRUE, seer0to19)
 
-table1(~can_type|In_Medicaid_Flag, seer0to19)
-
-table1(~age_cat + sex  + race_ethnicity + Yost_ACS_2006_2010_quintile + In_Medicaid_Flag  + stage + fumo  + can_type + can_dead2|mcont66b, droplevels = TRUE, seer0to19)
-```
-
-# FOR REVISION (insurance distribution)
+# table1(~can_type|In_Medicaid_Flag, seer0to19)
+
+# table1(~age_cat + sex  + race_ethnicity + Yost_ACS_2006_2010_quintile + In_Medicaid_Flag  + stage + fumo  + can_type + can_dead2|mcont66b, droplevels = TRUE, seer0to19)
+```
+
+# FOR REVISION response (insurance distribution)
 ```{r}
@@
   remove_labels()
+
+table(seer0to19_stage$mcont66, seer0to19_stage$stage)
+
 ```
@@
 df20
+
+p
 ```
@@
 
-# Mcont66
+# mcont66
 df3a <- data.frame(matrix(ncol=6, nrow=0)) #create the table
@@
   theme(strip.placement = "inside", legend.position = "bottom",  strip.text = element_text(size = 9, face="bold", color="black"), strip.background = element_rect(fill="lightgray"), plot.title = element_text(face="bold"), legend.text=element_text(size=9, face="bold")) +
-  scale_color_manual(breaks = c("Discontinuous vs. Not enrolled", "Continuous vs. Not enrolled", "Other vs. Not enrolled"),
+  scale_color_manual(breaks = c("Discontinuous vs. Not enrolled","Continuous vs. Not enrolled", "Other vs. Not enrolled"),
                         values=c("red", "cyan3", "blue"))
@@
 Plot3
+
 table(SuppTable1a$`Exposure comparison`)
@@
   theme(strip.placement = "inside", legend.position = "bottom",  strip.text = element_text(size = 9, face="bold", color="black"), strip.background = element_rect(fill="lightgray"), plot.title = element_text(face="bold"), legend.text=element_text(size=9, face="bold")) +
-  scale_color_manual(breaks = c("Discontinuous vs. Not enrolled", "Continuous vs. Not enrolled", "Other vs. Not enrolled"),
+  scale_color_manual(breaks = c( "Discontinuous vs. Not enrolled","Continuous vs. Not enrolled", "Other vs. Not enrolled"),
                         values=c("red", "cyan3", "blue"))
@@
 ```
-# Survival plots with discontinuous divided to at diagnosis and the rest
+# Survival plots with discontinuous divided to at diagnosis and the rest (NOT IN PAPER)
 ```{r}
@@
 
-pdf("Figure 3.pdf", width=8.5, height=11, paper="USr")
-plot1$plot + plot2$plot + plot_layout(ncol=1, nrow=2, guides='keep') &
-  theme(legend.position='bottom')
-
-dev.off()
 ```
@@
 # Overall
-overall <- summary(survfit(Surv(fumo, can_dead) ~ 1, data = seer0to19), times=60) # estimates 60 month survival time
+overall <- summary(survfit(Surv(fumo, can_dead) ~ 1, data = seer0to19), times=60) # estimates 60 month survival time overall
 
 overall <- as.data.frame(cbind(overall$time, round(overall$surv*100, 2), round(overall$lower*100, 2), round(overall$upper*100,2))) # makes dataframe
+
 overall$Category <- "Overall"
+
 overall <- setNames(overall, c("Time (months)", "Survival Percentage", "Lower CI", "Upper CI", "Category")) # gives column names
+
 overall <- overall %>% select("Category","Time (months)", "Survival Percentage", "Lower CI", "Upper CI")
@@
 # adjusted models 0 to 19
-res.cox2 <- coxph(Surv(fumo, can_dead) ~ In_Medicaid_Flag + race_ethnicity + age_cat + Yost_ACS_2006_2010_quintile + cluster(state), data = seer0to19) # Model adjusted for race age and a measure of poverty
+res.cox2 <- coxph(Surv(fumo, can_dead) ~ In_Medicaid_Flag + race_ethnicity + age_cat + Yost_ACS_2006_2010_quintile + cluster(state), data = seer0to19) # Model adjusted for race, age and a measure of poverty
 res.cox2
@@
 # adjusted models 0 to 19
-Medflag <- cox.zph(res.cox2, terms = FALSE)
+Medflag <- cox.zph(res.cox2, terms = TRUE)
 Medflag
@@
 # mcont66 adjusted models 0 to 19
-Med4cat <- cox.zph(res.cox4, terms = FALSE)
+Med4cat <- cox.zph(res.cox4, terms = TRUE)
 Med4cat
@@
 seer0to19.cp.format <- seer0to19.cp.format[with(seer0to19.cp.format, order(patient_id, fumo_half)),]  # sort by patient_id and then by fumo
+
 summary(seer0to19$fumo_half)
@@
 library(multcomp)
-X<- matrix(c(1,0,0,0,0,0,0,0,0,0,0,0,0,0,0), 1)
+X<- matrix(c(1,0,0,0,0,0,0,0,0,0,0,0,0,0,0), 1) # number of 0's and 1's in vector is equal to number of coefficients in model
 
@@
 
+# set up matrices to calculate HRs for different intervals
 W <- matrix(c(1,0,0,0,0,0,0,0,0,0,0,0,0,0,0), 1) 
@@
 x <- coxph(SurvObj ~ In_Medicaid_Flag + In_Medicaid_Flag*strata(interval) + race_ethnicity + age_cat + Yost_ACS_2006_2010_quintile + cluster(state), data = seer0to19.cp.format)
+
   t0<- glht(x, linfct = W) 
@@
 ```{r}
+# factor for order
+results$interval <- factor(results$interval, levels = c("0 to ≥12", ">12 to ≥24", ">24 to ≥36", "≥36"))
+
 p<- results %>%
@@
 ```{r}
-SuppTable6b %>%
+
+plot <- SuppTable6b %>%
   group_by(Group) %>%
@@
   theme(strip.placement = "inside", legend.position = "bottom",  strip.text = element_text(size = 9, face="bold", color="black"), strip.background = element_rect(fill="lightgray"), plot.title = element_text(face="bold"), legend.text=element_text(size=12, face="bold"))
+
+plot
 ```
@@
 
-X <- matrix(c(0,1,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0), 1)
+X <- matrix(c(0,1,0,0,0,
+              0,0,0,0,0,
+              0,0,0,0,0,
+              0,0,0,0,0,
+              0,0,0), 1) # 23 coefficients
 
@@
 
-A <- matrix(c(1,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0), 1) # discontinuous 1st interval
-B <- matrix(c(0,1,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0), 1) # other 1st interval
-C <- matrix(c(0,0,1,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0), 1) # Not enrolled 1st interval
-
-D <- matrix(c(1,0,0,0,0,0,0,0,0,0,0,0,0,0,1,0,0,0,0,0,0,0,0), 1) # discontinuous 2nd interval
-E <- matrix(c(0,1,0,0,0,0,0,0,0,0,0,0,0,0,0,1,0,0,0,0,0,0,0), 1) # other 2nd interval
-F <- matrix(c(0,0,1,0,0,0,0,0,0,0,0,0,0,0,0,0,1,0,0,0,0,0,0), 1) # Not enrolled 2nd interval
-
-G <- matrix(c(1,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,1,0,0,0,0,0), 1) # discontinuous 3rd interval
-H <- matrix(c(0,1,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,1,0,0,0,0), 1) # other 3rd interval
-I <- matrix(c(0,0,1,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,1,0,0,0), 1) # Not enrolled 3rd interval
-
-J <- matrix(c(1,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,1,0,0), 1) # discontinuous 3rd interval
-K <- matrix(c(0,1,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,1,0), 1) # other 3rd interval
-L <- matrix(c(0,0,1,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,1), 1) # Not enrolled 3rd interval
+A <- matrix(c(1,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0), 1) # Continuous 1st interval
+B <- matrix(c(0,1,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0), 1) # Discontinuous 1st interval
+C <- matrix(c(0,0,1,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0), 1) # Other 1st interval
+
+D <- matrix(c(1,0,0,0,0,0,0,0,0,0,0,0,0,0,1,0,0,0,0,0,0,0,0), 1) # Continuous 2nd interval
+E <- matrix(c(0,1,0,0,0,0,0,0,0,0,0,0,0,0,0,1,0,0,0,0,0,0,0), 1) # Discontinuous 2nd interval
+F <- matrix(c(0,0,1,0,0,0,0,0,0,0,0,0,0,0,0,0,1,0,0,0,0,0,0), 1) # Other 2nd interval
+
+G <- matrix(c(1,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,1,0,0,0,0,0), 1) # Continuous 3rd interval
+H <- matrix(c(0,1,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,1,0,0,0,0), 1) # Discontinuous 3rd interval
+I <- matrix(c(0,0,1,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,1,0,0,0), 1) # Other 3rd interval
+
+J <- matrix(c(1,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,1,0,0), 1) # Continuous 4th interval
+K <- matrix(c(0,1,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,1,0), 1) # Discontinuous 4th interval
+L <- matrix(c(0,0,1,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,1), 1) # Other 4th interval
 
@@
 SuppTable7c <- rbind(a,b,c,d,f,h,i,j,k, m,n,o,p,r,s,t,u,v,w)
-
 ```

### 2024-01-11 Manuscript Analysis code 0 to 19.Rmd
latest version after minor clean-up

@@
 ```{r}
+# install.packages("pacman")
 # install.packages("pacman")
@@
 
-x <- unique(df$patient_id) # 841 vector of unique ids who had enrollment in multiple states to exclude in exclusion section
+x <- unique(df$patient_id) # 3588 vector of unique ids who had enrollment in multiple states to exclude in exclusion section
 ```
@@
 ```{r}
-Figure1<-grViz("digraph flowchart {
+Figure1 <- grViz("digraph flowchart {
       # node definitions with substituted label text
@@
       ")
+
 Figure1
+
+library(DiagrammeRsvg)  # for conversion to svg
+library(rsvg)  # for saving svg
+
+Figure1 %>%
+   DiagrammeRsvg::export_svg() %>% 
+  charToRaw() %>% 
+  rsvg::rsvg_pdf("Figure1.pdf")
 ```
@@
 
-# SUPPLEMENTAEY TABLE 1:
+# SUPPLEMENTAEY TABLE 1: Stage at diagnosis cancer type analysis
 ```{r}
@@
 # graph ORs and 95% CIs
-SuppTable1a$`Exposure comparison` <- factor(SuppTable1a$`Exposure comparison`, levels = c("Enrolled vs.
+SuppTable1$`Exposure comparison` <- factor(SuppTable1$`Exposure comparison`, levels = c("Enrolled vs.
 Not enrolled", "Discontinuous vs.
@@
 
-# the if_elsee function does not like labels
+# the if_else function does not like labels
 seer0to19 <- seer0to19 %>% 
